---
title: "Processing images"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Introduction

All scripts were run from the directory

```
cd /project2/gilad/joycehsiao/fucci-seq/code/
```

and under the `conda` environment [fucci-seq](/project2/gilad/joycehsiao/fucci-seq/environment.yaml). To activate the `conda` environment:

```
cd /project2/gilad/joycehsiao/fucci-seq
source activate fucci-seq
```

The raw image files are located at:

```
cd /project2/gilad/fucci-seq/code/images_curated
```

There are three majar steps in image processing. In each step, we use R script to perform the analysis and then use bash script to submit jobs to the cluster.

Step 1: Detect nucleus outline

Step 2: Combine images from all channels 

Step 3: Compute summary metrics for image intensities


---

## Detect nucleus outline

[fucciMask.R](/project2/gilad/joycehsiao/fucci-seq/code/fucciMask.R) processes data for one plate at a time, from reading in images from all four channels (`/project2/gilad/fucci-seq/images_curated/plateID`), detecting nucleus edge using DAPI channel, outputing cropped images fro each channel (`/scratch/midway2/joycehsiao/fucci-seq/images_processed`), to saving matrices identifying areas inside nucleus (`/project2/gilad/fucci-seq/intensities/plateID/*.ints.rds`), matrices containing pixel intensities for each channel (`/project2/gilad/fucci-seq/intensities/plateID/*.ints.rds`) and matrices containg the number of nuclei detected for each sample (`project2/gilad/fucci-seq/intensities/*.rds`).  

[submit-fucciMask.sh](/project2/gilad/joycehsiao/fucci-seq/code/submit-fucciMask.sh) submits batch jobs running [fucciMask.R](/project2/gilad/joycehsiao/fucci-seq/code/fucciMask.R) for one plate a time.

```
bash submit-fucciMask.sh
```


## Combine images from all channels 

[makeWellIdImages.R](/project2/gilad/joycehsiao/fucci-seq/code/makeWellIdImages.R) creates labels (00001, ..., 00096) for the images files and saves the labels in `/project2/gilad/fucci-seq/images_wellid`.

[fucciMask.combineImages.R](/project2/gilad/joycehsiao/fucci-seq/code/fucciMask.combineImages.R) combines the image labels (`/project2/gilad/fucci-seq/images_wellid`) with the sample cropped images (`/scratch/midway2/joycehsiao/fucci-seq/images_processed`) for one plate at a time, making a combined image for each single cell sample for visualization (`/project2/gilad/fucci-seq/images_combined`).


[submit-fucciMask.combineImages.sh](/project2/gilad/joycehsiao/fucci-seq/code/submit-fucciMask.combineImages.sh) submits batch jobs running [fucciMask.combineImages.R](/project2/gilad/joycehsiao/fucci-seq/code/fucciMask.combineImages.R) for one plate at a time.

```
bash submit-fucciMask.combineImages.sh
```

## Compute summary metrics for image intensities

[intensity.summarize.R](/project2/gilad/joycehsiao/fucci-seq/code/intensity.summarize.R) uses R packakge [EBImage](https://bioconductor.org/packages/release/bioc/html/EBImage.html) to compute intensity summary statistics for the single cell samples detected to have single nucleus. The computation is done for one plate a time. Results are stored in `/project2/gilad/joycehsiao/fucci-seq/intensities_stats`.

[submit-intensity.summarize.sh](/project2/gilad/joycehsiao/fucci-seq/code/submit-intensity.summarize.sh) submits batch jobs running [intensity.summarize.R](/project2/gilad/joycehsiao/fucci-seq/code/intensity.summarize.R) for one plate at a time.

```
bash submit-intensity.summarize.sh
```


