---
title: "Correlation between intensities and expression"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Overview/Results


I computed correlation between intensity measurements and gene expression (log2cpm). Then apply CorShrink to account for varying sample size between correlations. 

Results show that there are some indication of cell cycle signal in teh data. Of the genes with s-vaule < .01, 33 correlate with DAPI are annotated as cell cycle gene, 74 correlation with RFP and annotated as cell cycle gene, and 110 correlate with GFP are annotated as cell cycle gene. However, these numbers are not large consider that there are a total of 11429 genes in the data, and among these only 1055 are annotated as cell cycle gene. 

In addition, there are sources of variation that may drive sample differences other than cell cycle, and crucially our design induces a strong sample variation between batches, as a trade-off for avoiding confounding of individual and plate effect. Hence analysis of sample variation in our data should consider technical sources of variation along with biological variation.  

---

## Data and packages

Packages

```{r}
library(CorShrink)
library(mygene)
```

Load data

```{r}
df <- readRDS(file="../data/eset-filtered.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

# import corrected intensities
pdata.adj <- readRDS("../output/images-normalize-anova.Rmd/pdata.adj.rds")

log2cpm <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.rds")
log2cpm.adjust <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.adjust.rds")

log2cpm <- log2cpm[grep("ENSG", rownames(log2cpm)), 
                   colnames(log2cpm) %in% rownames(pdata.adj)]

log2cpm.adjust <- log2cpm.adjust[grep("ENSG", rownames(log2cpm)),
                   colnames(log2cpm.adjust) %in% rownames(pdata.adj)]

all.equal(rownames(pdata.adj), colnames(log2cpm))

genes.variable <- readRDS("../output/seqdata-select-cellcyclegenes.Rmd/genes.variable.rds")
genes.cycle <- readRDS(file = "../output/seqdata-select-cellcyclegenes.Rmd/genes.cycle.rds")
```



## Correlations

compute correlation between adjusted intensities and log2 expression data.

```{r, fig.width=6, fig.height=5}
corr.rfp <- do.call(rbind, lapply(1:nrow(log2cpm), function(i) {
  vec <- cbind(pdata.adj$rfp.median.log10sum.adjust.ash,
               log2cpm[i,])
  filt <- counts[i,] > 1
  nsamp <- sum(filt)
  if (nsamp > 100) {
#    cnt <- counts[i,filt]
    vec <- vec[filt,]
    corr <- cor(vec[,1], vec[,2])
    nsam <- nrow(vec)
    data.frame(corr=corr, nsam=nsam)
  } else {
    data.frame(corr=NA, nsam=nrow(vec))
  }
  }) )


corr.gfp <- do.call(rbind, lapply(1:nrow(log2cpm), function(i) {
  vec <- cbind(pdata.adj$gfp.median.log10sum.adjust.ash,
               log2cpm[i,])
  filt <- counts[i,] > 1
  nsamp <- sum(filt)
  if (nsamp > 100) {
    vec <- vec[filt,]
    corr <- cor(vec[,1], vec[,2])
    nsam <- nrow(vec)
    data.frame(corr=corr, nsam=nsam)
  } else {
    data.frame(corr=NA, nsam=nrow(vec))
  }
  }) )


corr.dapi <- do.call(rbind, lapply(1:nrow(log2cpm), function(i) {
  vec <- cbind(pdata.adj$dapi.median.log10sum.adjust.ash,
               log2cpm[i,])
  filt <- counts[i,] > 1
  nsamp <- sum(filt)
  if (nsamp > 100) {
    vec <- vec[filt,]
    corr <- cor(vec[,1], vec[,2])
    nsam <- nrow(vec)
    data.frame(corr=corr, nsam=nsam)
  } else {
    data.frame(corr=NA, nsam=nrow(vec))
  }
  }) )
rownames(corr.rfp) <- rownames(log2cpm)
rownames(corr.gfp) <- rownames(log2cpm)
rownames(corr.dapi) <- rownames(log2cpm)


corr.rfp.val <- corr.rfp[!is.na(corr.rfp$corr),]
corr.gfp.val <- corr.gfp[!is.na(corr.gfp$corr),]
corr.dapi.val <- corr.dapi[!is.na(corr.dapi$corr),]


par(mfrow=c(2,2))
hist(corr.rfp.val$corr, main = "correlation between RFP and expression")
hist(corr.gfp.val$corr, main = "correlation between GFP and expression")
hist(corr.dapi.val$corr, main = "correlation between DAPI and expression")
```

Apply CorShrink.

```{r}
corr.rfp.shrink <- CorShrinkVector(corr.rfp.val$corr, nsamp_vec = corr.rfp.val$nsam,
                                   optmethod = "mixEM", report_model = TRUE)
names(corr.rfp.shrink$estimate) <- rownames(corr.rfp.val)

corr.gfp.shrink <- CorShrinkVector(corr.gfp.val$corr, nsamp_vec = corr.gfp.val$nsam,
                                   optmethod = "mixEM", report_model = TRUE)
names(corr.gfp.shrink$estimate) <- rownames(corr.gfp.val)

corr.dapi.shrink <- CorShrinkVector(corr.dapi.val$corr, nsamp_vec = corr.dapi.val$nsam,
                                   optmethod = "mixEM", report_model = TRUE)
names(corr.dapi.shrink$estimate) <- rownames(corr.dapi.val)


par(mfcol=c(2,2))
plot(corr.rfp.val$corr, corr.rfp.shrink$estimate,
     col=1+as.numeric(corr.rfp.shrink$model$result$svalue < .01),
     xlim=c(-.5,.5), ylim=c(-.5,.5), pch=16, cex=.8,
     xlab = "Correlation", ylab = "Shrunken estimate",
     main = "RFP")
abline(0,1)

plot(corr.gfp.val$corr, corr.gfp.shrink$estimate,
     col=1+as.numeric(corr.gfp.shrink$model$result$svalue < .01),
     xlim=c(-.5,.5), ylim=c(-.5,.5), pch=16, cex=.8,
     xlab = "Correlation", ylab = "Shrunken estimate",
     main = "GFP")
abline(0,1)

plot(corr.dapi.val$corr, corr.dapi.shrink$estimate,
     col=1+as.numeric(corr.dapi.shrink$model$result$svalue < .01),
     xlim=c(-.5,.5), ylim=c(-.5,.5), pch=16, cex=.8,
     xlab = "Correlation", ylab = "Shrunken estimate",
     main = "DAPI")
abline(0,1)
title("s-value < .01", outer = TRUE, line = -1)
```

Of the ones with s-value < .05, consider the proportion of genes in our annotated set.

```{r}
ii.dapi <- names(corr.dapi.shrink$estimate)[which(corr.dapi.shrink$model$result$svalue < .01)]
print(sum(ii.dapi %in% genes.cycle))
print(sum(ii.dapi %in% genes.variable))

ii.rfp <- names(corr.rfp.shrink$estimate)[which(corr.rfp.shrink$model$result$svalue < .01)]
print(sum(ii.rfp %in% genes.cycle))
print(sum(ii.rfp %in% genes.variable))

ii.gfp <- names(corr.gfp.shrink$estimate)[which(corr.gfp.shrink$model$result$svalue < .01)]
print(sum(ii.gfp %in% genes.cycle))
print(sum(ii.gfp %in% genes.variable))
```



---

## Session information

```{r, echo = FALSE}
sessionInfo()
```



