---
title: "Evaluate projected cell time"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```


---

## Estimate cell time

```{r}
library(Biobase)
# load gene expression
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

log2cpm.all <- t(log2(1+(10^6)*(t(exprs(df))/pdata$molecules)))

macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")

pc.fucci <- prcomp(subset(pdata, 
                          select=c("rfp.median.log10sum.adjust",
                                   "gfp.median.log10sum.adjust")),
                   center = T, scale. = T)
Theta.cart <- pc.fucci$x
library(circular)
Theta.fucci <- coord2rad(Theta.cart)
Theta.fucci <- 2*pi - Theta.fucci
```

Cluster cell times to move the origin of the cell times

```{r, eval = F}
# cluster cell time
library(movMF)
clust.res <- lapply(2:5, function(k) {
  movMF(Theta.cart, k=k, nruns = 100, kappa = list(common = TRUE))
})
k.list <- sapply(clust.res, function(x) length(x$theta) + length(x$alpha) + 1)
bic <- sapply(1:length(clust.res), function(i) {
  x <- clust.res[[i]]
  k <- k.list[i]
  n <- nrow(Theta.cart)
  -2*x$L + k *(log(n) - log(2*pi)) })
plot(bic)
labs <- predict(clust.res[[2]])

saveRDS(labs, file = "../output/images-time-eval.Rmd/labs.rds")
```


```{r}
labs <- readRDS(file = "../output/images-time-eval.Rmd/labs.rds")

summary(as.numeric(Theta.fucci)[labs==1])
summary(as.numeric(Theta.fucci)[labs==2])
summary(as.numeric(Theta.fucci)[labs==3])

# move the origin to 1.17
Theta.fucci.new <- vector("numeric", length(Theta.fucci))
cutoff <- min(Theta.fucci[labs==2])
Theta.fucci.new[Theta.fucci>=cutoff] <- Theta.fucci[Theta.fucci>=cutoff] - cutoff
Theta.fucci.new[Theta.fucci<cutoff] <- Theta.fucci[Theta.fucci<cutoff] - cutoff + 2*pi
```

Try plotting for one gene

```{r}
macosko[macosko$hgnc == "CDK1",]
cdk1 <- log2cpm.all[rownames(log2cpm.all)=="ENSG00000170312",]
plot(x=Theta.fucci.new, y = cdk1)
points(y=cdk1[labs==1], x=as.numeric(Theta.fucci.new)[labs==1], pch=16, cex=.7, col = "red")
points(y=cdk1[labs==2], x=as.numeric(Theta.fucci.new)[labs==2], pch=16, cex=.7, col = "forestgreen")
points(y=cdk1[labs==3], x=as.numeric(Theta.fucci.new)[labs==3], pch=16, cex=.7, col = "blue")
```



---

## Results

Check data points with outlier DAPI values

```{r}
ii.min.dapi <- order(pData(df)$dapi.median.log10sum.adjust)[1:2]

pData(df)[ii.min.dapi,]
```


```{r, fig.width=7, fig.height=7}
par(mfrow=c(2,2))
ylims <- with(pdata, range(c(dapi.median.log10sum.adjust,
                             gfp.median.log10sum.adjust,
                             rfp.median.log10sum.adjust)))
plot(as.numeric(Theta.fucci.new), pdata$dapi.median.log10sum.adjust, col = "blue", 
     ylab= "adjusted intensity values",
     ylim = ylims, main = "DAPI intensity values",
     xlab ="Estimated cell time")
points(as.numeric(Theta.fucci.new)[ii.min.dapi],
       pdata$dapi.median.log10sum.adjust[ii.min.dapi], pch=4, lwd=2, cex=1)
plot(as.numeric(Theta.fucci.new), pdata$gfp.median.log10sum.adjust, col = "forestgreen", 
     ylab= "adjusted intensity values",
     ylim = ylims, main = "GFP and RFP intensity values",
     xlab ="Estimated cell time")
points(as.numeric(Theta.fucci.new)[ii.min.dapi],
       pdata$gfp.median.log10sum.adjust[ii.min.dapi], pch=4, lwd=2, cex=1)
points(as.numeric(Theta.fucci.new), pdata$rfp.median.log10sum.adjust, col = "red")
points(as.numeric(Theta.fucci.new)[ii.min.dapi],
       pdata$rfp.median.log10sum.adjust[ii.min.dapi], pch=4, lwd=2, cex=1)

plot(as.numeric(Theta.fucci.new), pdata$molecules, main = "Total molecule count",
     xlab ="Estimated cell time", ylab = "Total molecule count", col = "gray50")
points(as.numeric(Theta.fucci.new)[ii.min.dapi],
       pdata$molecules[ii.min.dapi], pch=4, lwd=2, cex=1)

plot(pdata$dapi.median.log10sum.adjust, pdata$molecules, main = "DAPI vs. molecule count",
     xlab = "DAPI intensity adjusted for C1 batch", ylab = "Total molecule count", 
     col = "gray50")
points(pdata$dapi.median.log10sum.adjust[ii.min.dapi],
       pdata$molecules[ii.min.dapi], pch=4, lwd=2, cex=1)
```

Test the association between total sample molecule count and DAPI.

* After excluding outliers, pearson correlation is .2

* Consider lm(molecules ~ dapi). The adjusted R-squared is .04

* Consider lm(log10(molecules) ~ dapi). The adjusted R-squared is .04

* Weak linear trend between molecule count and DAPI...


```{r}
xy <- data.frame(dapi=pdata$dapi.median.log10sum.adjust,
                 molecules=pdata$molecules,
                 chip_id=pdata$chip_id)
xy <- xy[xy$dapi > -1,]

fit <- lm(molecules~dapi+factor(chip_id), data=xy)
summary(fit)
fit <- lm(log10(molecules)~dapi+factor(chip_id), data=xy)
summary(fit)

fit <- lm(molecules~dapi, data=xy)
summary(fit)
fit <- lm(log10(molecules)~dapi, data=xy)
summary(fit)

cor(xy$dapi, xy$molecules, method = "pearson")
cor(xy$dapi, log10(xy$molecules), method = "pearson")


par(mfrow=c(1,2))
plot(x=xy$dapi, y = xy$molecules,
     xlab = "DAPI intensity", ylab = "Total molecule count",
     main = "DAPI vs. molecule count \n adj-Rsq = .03; cor=.2")
abline(lm(molecules~dapi, data=xy), col = "red")
plot(x=xy$dapi, y = log10(xy$molecules),
     xlab = "DAPI intensity", ylab = "log10 total molecule count",
     main = "DAPI vs. log10 molecule count \n adj-Rsq = .03, cor=.2")
abline(lm(log10(molecules)~dapi, data=xy), col = "red")

```

---

## Save re-ordered cell times

```{r, eval = F}
theta <- as.numeric(Theta.fucci.new)
names(theta) <- colnames(log2cpm.all)

saveRDS(theta,
        file = "../output/images-time-eval.Rmd/theta.rds")
```

---


```{r}
sessionInfo()
```

