---
title: "Inspect samples detected as absent or not having a nucleus"
author: "Joyce Hsiao"
output: html_document
---


<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load packages and data 

```{r}
library(knitr)
```

Load nuclei detection results directly from image analysis output `/project2/gilad/fucci-seq/intensities`. The results are made into a summary table.

`plate`: labeled by the mix of individual cell lines  
`image`: labels assigned by the JULI imaging system. These will be converted to correspond to C1 plate labels (rows and columns).  
`nnuclei`: number of nuclei detected in a single cell sample  

```{r, eval = FALSE}
fls <- list.files("/project2/gilad/fucci-seq/intensities", pattern = "rds", full.names = TRUE)

nuclei_table <- do.call(rbind, lapply(1:length(fls), function(index) {
  tmp <- readRDS(fls[[index]])
  plate <- strsplit(strsplit(fls[index], split="/")[[1]][[6]], 
                    split=".", fixed = TRUE)[[1]][[1]]
  data.frame(plate=plate, nnuclei=tmp, image=names(tmp))
}))

saveRDS(nuclei_table, "/project2/gilad/joycehsiao/fucci-seq/output/image-qc.rds")
```

Inspect samples

```{r}
nuclei_table <- readRDS("/project2/gilad/joycehsiao/fucci-seq/output/image-qc.rds")
kable(nuclei_table[which(nuclei_table$nnuclei==0|is.na(nuclei_table$nnuclei)),])
```


---

## Session information

```{r}
sessionInfo()
```

