---
title: "Investigate intensities against transgenes"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```


---

## Introduction/Summary

Here I combine image intensity data with sequencing data on the FUCCI transgenes, focusing on the subset of samples that are identified as single nucleus (1,264 samples). 

**Results**

1. There are more reads mapped to EGFP than to mCherry. The regions in the transgenes statring from 5' end: promoter region -> EGFP -> 20 bp linker -> mCherry. We don't yet have an explanation on the low number of reads mapped to mCherry... We decided to use EGFP to determine if the transgene is silenced at all. 

2. Consistent with image intensity data, there are more variation between batches of C1 plates than between individuals. Fortunatley, given our balanced design, technical variation between batches are spread out across individuals. The main concern here would be the inflated within individual variation due to differences between the C1 plates.

3. Considering the relationship between intensity and transgene molecule count: samples with no EGFP molecules also have lower than average (z-score < 0) GFP and RFP intensity. 

---

## Packages

```{r}
library(ggplot2)
library(cellcycleR)
library(data.table)
library(dplyr)
library(cowplot)
library(wesanderson)
library(RColorBrewer)
library(Biobase)
```

---

## Data

Import sequencing data.

```{r}
fname <- Sys.glob("../data/eset/*.rds")
eset <- Reduce(combine, Map(readRDS, fname))

pdata <- pData(eset)
pdata$unique <- paste(pdata$image_individual, sprintf("%05d", pdata$image_label), sep="_")

transgenes <- data.frame(t(exprs(eset)[c("EGFP", "mCherry"),]))
```


Imort intensity data for samples with single nucleus. For source of the intensity data, see [combine-intensity-data.R][/project2/gilad/joycehsiao/fucci-seq/code/combine-intensity-data.R].


```{r, echo = TRUE}
ints <- readRDS(file="/project2/gilad/joycehsiao/fucci-seq/data/intensity.rds")
ints$rfp.z <- scale(ints$rfp.median.log10sum, center=TRUE, scale=TRUE)
ints$gfp.z <- scale(ints$gfp.median.log10sum, center=TRUE, scale=TRUE)
ints$dapi.z <- scale(ints$dapi.median.log10sum, center=TRUE, scale=TRUE)

ints_tmp <- data.frame(unique=ints$unique,
                       rfp.z=ints$rfp.z,
                       gfp.z=ints$gfp.z,
                       dapi.z=ints$dapi.z, 
                       stringsAsFactors = F)
```

Combine intensities with sequencing data. This includes 1,264 samples with single nucleus.

```{r}
combo <- ints_tmp
combo$EGFP <- transgenes$EGFP[match(combo$unique, pdata$unique)]
combo$mCherry <- transgenes$mCherry[match(combo$unique, pdata$unique)]
combo$chip_id <- pdata$chip_id[match(combo$unique, pdata$unique)]
combo$experiment <- pdata$experiment[match(combo$unique, pdata$unique)]
```

$~$

---

## Transgene count

**Results**: there are more reads mapped to EGFP than to mCherry. The regions in the transgenes statring from 5' end: promoter region -> EGFP -> 20 bp linker -> mCherry. We don't yet have an explanation on the low number of reads mapped to mCherry... We decided to use EGFP to determine if the transgene is silenced at all. 

```{r}
summary(combo$EGFP)
summary(combo$mCherry)
```

```{r, fig.width=7, fig.height=6, echo = FALSE}
par(mfrow=c(2,2))
plot(combo$EGFP, combo$gfp.z,
     xlab = "GFP intensity", ylab = "EGFP molecule count")
plot(combo$EGFP, combo$rfp.z,
     xlab = "RFP intensity", ylab = "EGFP molecule count")
plot(combo$mCherry, combo$gfp.z,
     xlab = "GFP intensity", ylab = "mCherry molecule count")
plot(combo$mCherry, combo$rfp.z,
     xlab = "RFP intensity", ylab = "mCherry molecule count")
```

$~$

---

## EGFP

### By individual

```{r, fig.width=7, fig.height=7, echo = FALSE}
ggplot(combo, aes(x=EGFP)) +
  geom_histogram() + facet_wrap(~chip_id)
```

$~$

### By C1 plate

```{r, fig.width=11, fig.height=11, echo = FALSE}
ggplot(combo, aes(x=EGFP)) +
  geom_histogram() + facet_wrap(~experiment, ncol=4) 
```

$~$

### By individual and C1 plate

```{r, fig.width=12, fig.height=20, echo = FALSE}
ggplot(combo, aes(x=EGFP)) +
  geom_histogram() + facet_grid(experiment ~ chip_id)
```

$~$

---

## mCherry

### By individual

```{r, fig.width=7, fig.height=7, echo = FALSE}
ggplot(combo, aes(x=mCherry)) +
  geom_histogram() + facet_wrap(~chip_id)
```

$~$

### By C1 plate

```{r, fig.width=11, fig.height=11, echo = FALSE}
ggplot(combo, aes(x=mCherry)) +
  geom_histogram() + facet_wrap(~experiment, ncol=4) 
```

$~$

### By individual and C1 plate

```{r, fig.width=12, fig.height=20, echo = FALSE}
ggplot(combo, aes(x=mCherry)) +
  geom_histogram() + facet_grid(experiment ~ chip_id)
```


$~$

---

## Session information

```{r, echo = FALSE}
sessionInfo()
```


