---
title: "Inspect samples detected having more than 1 nucleus"
author: "Joyce Hsiao"
output: html_document
---


<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load data 

Load image QC table - for now it has only number of nuclei detected in each sample now. This table does not include all of the single cell samples, and the ones that are not included are the samples classifed as no DNA content in DAPI images. 

The summary table contains three columns:

`plate`: labeled by the mix of individual cell lines  
`well`: labels assigned by the JULI imaging system. These will be converted to correspond to C1 plate labels (rows and columns).  
`nnuclei`: number of nuclei detected in a single cell sample  

```{r}
nuclei_table <- readRDS("/project2/gilad/joycehsiao/fucci-seq/output/image-qc.rds")

with(nuclei_table, table(nnuclei))
```

Inspect samples with 12 nuclei.

```{r}
nuclei_table[which(nuclei_table$nnuclei == 12),]
```

Inspect samples with 20 nuclei.

```{r}
nuclei_table[which(nuclei_table$nnuclei == 20),]
```

---

## Printing images for inspection

**This section is for record keeping. All images for samples with more than 1 nucleus detected have been converted to PNG format and stored in /project2/gilad/fucci-seq/images_inspect.**

First select samples for inspection. The code below selects from plate `18511_18855` the samples that have two nuclei detected.

```{r}
cases <- nuclei_table[which(nuclei_table$nnuclei == 2 & nuclei_table$plate == "18511_18855"),]
```

Extract the wells labels and convert to a character variable.

```{r}
wells <- as.character(cases$well)

wells <- c("00012","00032","00044","00047","00051","00052","00067","00076","00085","00088","00096")

print(wells)
```

The code below will save images in png format at `/project2/gilad/fucci-seq/images_inspect`.

```{r, eval = FALSE}
#' @param wells vector of wells ID (use the format specified above).
#' @param plate plate ID

print_png <- function(plate, wells) {
  for (index in 1:length(wells)) {
    id <- wells[index]
    dir_images_data_pl <- paste0("/project2/gilad/fucci-seq/images_curated/",plate,"/")
    dir_output <- "/project2/gilad/fucci-seq/images_inspect/"
    bright <- readImage(paste0(dir_images_data_pl, "BRIGHT/", id, ".TIFF"))
    dapi <- readImage(paste0(dir_images_data_pl, "DAPI/", id, ".TIFF"))
    gfp <- readImage(paste0(dir_images_data_pl, "GFP/", id, ".TIFF"))
    rfp <- readImage(paste0(dir_images_data_pl, "RFP/", id, ".TIFF"))
    
    writeImage(bright, paste0(dir_output, plate,".",id,".bright.png"))
    writeImage(dapi, paste0(dir_output, plate,".",id,".dapi.png"))
    writeImage(gfp, paste0(dir_output, plate,".",id,".gfp.png"))
    writeImage(rfp, paste0(dir_output, plate,".",id,".rfp.png"))
    # combo <- combine(dapi, bright, gfp, rfp)
    # writeImage(combo, paste0(dir_output, plate,".",id,".combo.png"))
  }
}


plates <- c("18511_18855","18855_19101","18855_19160","18870_18511",
            "18870_18855","18870_19101","18870_19160","19098_18511",
            "19098_18870","19098_19160","19101_18511","19101_19098",
            "19160_18870","18855_19098","19101_19160","19160_18511")

for (index in 1:length(plates)) {
  plate <- plates[index]
  cases <- nuclei_table[which(nuclei_table$nnuclei > 1 & nuclei_table$plate == plate),]
  wells <- as.character(cases$well)
  print_png(plate, wells) 
}


```


### Single cell using manual inspection

The distortion during the sorting makes the shape not as circular, resulting in mis-identification.

18511_18855.00012
18511_18855.00052
18511_18855.00067
18855_19098.00005
18855_19098.00009
18855_19098.00013
18855_19098.00039
18855_19101.00021
18855_19101.00038
18855_19101.00063
18855_19160.00079
18870_18855.00060
18870_19101.00092
18870_19160.00039
19098_18511.00050
19098_18870.00051
19098_18870.00063
19098_19160.00046
19101_19098.00006
19101_19098.00054
19101_19098.00078
19160_18511.00005
19160_18511.00021
19160_18511.00024
19160_18511.00031
19160_18511.00038
19160_18511.00045
19160_18870.00002
19160_18870.00005
19160_18870.00011
19160_18870.00021
19160_18870.00024
19160_18870.00038
19160_18870.00039
19160_18870.00051
19160_18870.00056
19160_18870.00071
19160_18870.00078
19160_18870.00079
19160_18870.00080
19160_18870.00081
19160_18870.00091
19160_18870.00094





---

## Session information

```{r}
sessionInfo()
```

