---
title: "Circle fit to intensities"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Overview/Results

Here we estimate a circle fit on the two-dimensional intensity distriubtion of GFP and RFP. 


---

## Data and packages

Packages

```{r}
library(circular)
library(conicfit)
library(Biobase)
library(dplyr)
library(matrixStats)
library(CorShrink)

source("../code/circle.intensity.fit.R")
```

Load data

```{r}
df <- readRDS(file="../data/eset-filtered.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

# log2cpm <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.rds")
# log2cpm.adjust <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.adjust.rds")

# import corrected intensities
pdata.adj <- readRDS("../output/images-normalize-anova.Rmd/pdata.adj.rds")
```



visualize intensity data on a circle. project intensity data onto radians and the visualize the fitting on circles. then use cellcycleR to perform fitting and compare the model fitting with the projection results. it seems that the predicted ordering is especially bad for the gree intensities, which have many outliers. another reason I suspect isthat it attemps to fit similar amplitude/phase parameters to the green and red intensities. 


## Circle fitting

Based on all data.

```{r, fig.width=7, fig.height=5}
source("../code/circle.intensity.fit.R")
sample_names <- rownames(pdata.adj)

pdata.adj <- pdata.adj %>% group_by(chip_id) %>% 
  mutate(rfp.z=scale(rfp.median.log10sum.adjust),
            gfp.z=scale(gfp.median.log10sum.adjust),
            dapi.z=scale(dapi.median.log10sum.adjust))
rownames(pdata.adj) <- sample_names
  
par(mfrow=c(2,3))
for(i in 1:length(unique(pdata.adj$chip_id))) {
  id <- unique(as.character(pdata.adj$chip_id))[i]
  df_sub <- subset(pdata.adj, chip_id == id, select=c(gfp.z, rfp.z))  
  
  cpred <- circle.fit(df_sub)

  xlims <- range(df_sub[,1])
  ylims <- range(df_sub[,2])
  plot(df_sub, pch=16, col="gray50", xlim=xlims, ylim=ylims, cex=.7,
       main = id, xlab="GFP", ylab="RFP")
  points(cpred[,1], cpred[,2], col="blue", type = "p")
  points(mean(cpred[,1]), mean(cpred[,2]), col="red", pch=3, cex=2)
}
```


Consider deleted residuals.

```{r, fig.width=7, fig.height=5}
resids.del <- lapply(1:length(unique(pdata.adj$chip_id)), function(i) {
  id <- unique(as.character(pdata.adj$chip_id))[i]
  df_sub <- subset(pdata.adj, chip_id == id, select=c(gfp.z, rfp.z))  
  resids <- circle.fit.resid.delete(df_sub)  
  scale(resids)
})
names(resids.del) <- unique(pdata.adj$chip_id)

par(mfrow=c(2,3))
for(i in 1:length(unique(pdata.adj$chip_id))) {
  # id <- unique(as.character(pdata.adj$chip_id))[i]
  # df_sub <- subset(pdata.adj, chip_id == id, select=c(gfp.z, rfp.z))  
  hist(resids.del[[i]], main = unique(pdata.adj$chip_id)[i])
}
```


Remove samples with standardized residuals greater than 3.

```{r, fig.width=8, fig.height=6}
resids.del.remove <- lapply(1:length(unique(pdata.adj$chip_id)), function(i) {
  which(resids.del[[i]] > 3)
})
names(resids.del.remove) <- unique(pdata.adj$chip_id)

pdata.adj.filt <- do.call(rbind, lapply(1:length(unique(pdata.adj$chip_id)), function(i) {
  id <- unique(as.character(pdata.adj$chip_id))[i]
  df_sub <- pdata.adj[which(pdata.adj$chip_id == id),]
  ii.remove <- resids.del.remove[[i]]
  df_sub_return <- df_sub[-ii.remove,]
  rownames(df_sub_return) <- (rownames(pdata.adj)[which(pdata.adj$chip_id == id)])[-ii.remove]
  data.frame(df_sub_return)
}) )
```

Visualize fit after removing outliers.

```{r, fig.width=7, fig.height=5}
par(mfrow=c(2,3))
for(i in 1:length(unique(pdata.adj.filt$chip_id))) {
  id <- unique(as.character(pdata.adj.filt$chip_id))[i]
  df_sub <- subset(pdata.adj.filt, chip_id == id, select=c(gfp.z, rfp.z))  
  
  cpred <- circle.fit(df_sub)

  xlims <- range(df_sub[,1])
  ylims <- range(df_sub[,2])
  plot(df_sub, pch=16, col="gray50", xlim=xlims, ylim=ylims, cex=.7,
       main = id, xlab="GFP", ylab="RFP")
  points(cpred[,1], cpred[,2], col="blue", type = "p")
  points(mean(cpred[,1]), mean(cpred[,2]), col="red", pch=3, cex=2)
}
```

```{r, eval=F}
saveRDS(pdata.adj.filt, 
        file = "../output/images-circle-ordering.Rmd/pdata.adj.filt.rds")
```


## Project positions

```{r}
pdata.adj.filt <- readRDS("../output/images-circle-ordering.Rmd/pdata.adj.filt.rds")

proj.res <- vector("list", length=length(unique((pdata.adj$chip_id))))

for(i in 1:length(unique((pdata.adj$chip_id)))) {
  proj.res[[i]] <- vector("list",2)
  
  id <- unique(as.character(pdata.adj.filt$chip_id))[i]
  
  df_sub <- subset(pdata.adj.filt, 
                   chip_id == id, select=c(gfp.z, rfp.z))  
#  sample_ids <-
    
  cpred <- circle.fit(df_sub)

  proj.res[[i]][[1]] <- data.frame(cpred, df_sub)
  colnames(proj.res[[i]][[1]]) <- c("pos.pred.x", "pos.pred.y", "gfp.z", "rfp.z")
  
  # convert projected coordinates to radians
  # modulo 2*pi
  proj.res[[i]][[1]]$rads <- coord2rad(cbind(proj.res[[i]][[1]]$pos.pred.x,
                                        proj.res[[i]][[1]]$pos.pred.y))
  rownames(proj.res[[i]][[1]]) <- rownames(df_sub)
  
  # compute centers
  centers <- LMcircleFit(as.matrix(df_sub), ParIni=colMeans(as.matrix(df_sub)), IterMAX=50)
  
  proj.res[[i]][[2]] <- data.frame(x.center=centers[1], y.center=centers[2])
}
names(proj.res) <- unique(pdata.adj.filt$chip_id)
```

Save output

```{r, eval=F}
saveRDS(proj.res, file = "../output/images-circle-ordering.Rmd/proj.res.rds")
```

Plot circle fit.

```{r, fig.width=9, fig.height=6}
proj.res <- readRDS(file = "../output/images-circle-ordering.Rmd/proj.res.rds")

par(mfrow=c(2,3))
for (i in 1:length(proj.res)) {
  # xlims <- range(proj.res[[i]]$gfp.z)
  # ylims <- range(proj.res[[i]]$rfp.z)
  xlims <- c(-2.5, 2.5)
  ylims <- c(-2.5, 2.5)
  plot(subset(proj.res[[i]][[1]], select=c(gfp.z, rfp.z)), 
       pch=16, col="gray50", xlim=xlims, ylim=ylims, cex=.5, 
       main = names(proj.res)[i],
       xlab = "GFP", ylab = "RFP")
  points(proj.res[[i]][[1]]$pos.pred.x, proj.res[[i]][[1]]$pos.pred.y,
         col="blue", pch=1)
  points(proj.res[[i]][[2]]$x.center, proj.res[[i]][[2]]$y.center, 
         col="red", pch=3, cex=2)
}

  
par(mfrow=c(2,3))
for (i in 1:length(proj.res)) {
  plot(proj.res[[i]][[1]]$rads, stack=TRUE, bins=90,
       main = names(proj.res)[i])
}
```


---

## Property of the circle fit

### Intensity values by circle fit

```{r, fig.width=6, fig.height=5}
pdata.adj.filtered <- readRDS("../output/images-circle-ordering.Rmd/pdata.adj.filt.rds")
proj.res <- readRDS("../output/images-circle-ordering.Rmd/proj.res.rds")

for (i in 1:length(unique(pdata.adj.filt$chip_id))) {
  par(mfrow=c(2,2), mar = c(3,2,2,1))
  ids <- unique(as.character(pdata.adj.filt$chip_id))
  p_sub <- subset(pdata.adj.filt, chip_id == ids[i])
  #all.equal(rownames(p_sub), rownames(proj.res$NA18870[[1]]))
  plot(proj.res[[i]][[1]]$rads, stack=T, bins=180, main = "Distribution")
  library(RColorBrewer)
  color <- colorRampPalette(brewer.pal(11,"Spectral"))(11)
  plot(x=as.numeric(proj.res[[i]][[1]]$rads), 
       y=p_sub$rfp.z, pch=16, cex=.5, col=color[1], ylim=c(-2.5, 2.5),
       xlab = "Position on the circle",
       ylab = "RFP", main = "RFP")
  abline(h=0, lwd=.5)
  plot(x=as.numeric(proj.res[[i]][[1]]$rads), 
       y=p_sub$gfp.z, pch=16, cex=.7, col=color[9], ylim=c(-2.5, 2.5),
       xlab = "Position on the circle",
       ylab = "GFP", main = "GFP")
  abline(h=0, lwd=.5)
  plot(x=as.numeric(proj.res[[i]][[1]]$rads), 
       y=p_sub$dapi.z, pch=16, cex=.7, col=color[10], ylim=c(-2.5, 2.5),
       xlab = "Position on the circle",
       ylab = "DAPI", main = "DAPI")
  abline(h=0, lwd=.5)
  title(names(proj.res)[i], outer=TRUE, line =-1)
}
```


### Expression variation by cell time

```{r, eval = T}
# load cell cycle genes
genes.cycle <- readRDS("../output/seqdata-select-cellcyclegenes.Rmd/genes.cycle.detect.rds")

# log2cpm
log2cpm <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.rds")
log2cpm.adjust <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.adjust.rds")

counts.cycle <- counts[rownames(counts) %in% genes.cycle, ]
log2cpm.cycle <- log2cpm[rownames(log2cpm) %in% genes.cycle, ]
log2cpm.adjust.cycle <- log2cpm.adjust[rownames(log2cpm.adjust) %in% genes.cycle, ]
```


### Pearson correlation 

```{r}
corrs <- lapply(1:length(unique(pdata.adj.filt$chip_id)), function(i) {
  
  id <- unique(pdata.adj.filt$chip_id)[i]
  
  log2cpm_sub <- log2cpm.adjust.cycle[, match(rownames(proj.res[[i]][[1]]), colnames(log2cpm.adjust.cycle))]
  
  counts_sub <- counts.cycle[, match(rownames(proj.res[[i]][[1]]), colnames(counts.cycle))]
  
  corrs <- do.call(rbind, lapply(1:nrow(counts_sub), function(g) {
    vec <- cbind(as.numeric(proj.res[[i]][[1]]$rads),
                 log2cpm_sub[g,])
    filt <- counts_sub[g,] > 1
    nsamp <- sum(filt)

    if (nsamp > ncol(counts_sub)/2) {
      vec <- vec[filt,]
      corr <- cor(vec[,1], vec[,2])
      nsam <- nrow(vec)
      data.frame(corr=corr, nsam=nsam)
    } else {
      data.frame(corr=NA, nsam=nrow(vec))
    }
    }))
  rownames(corrs) <- rownames(counts_sub)
  return(corrs) 
  })   
names(corrs) <- unique(pdata.adj.filt$chip_id)
```


```{r, fig.height=5, fig.width=7}
par(mfrow=c(2,3))
for (i in 1:length(corrs)) {
  hist(corrs[[i]]$corr, main = names(corrs)[i])            
}
title(main = "Pearson correlation", outer = TRUE, line = -1)
```

Apply CorShrink

```{r}
par(mfrow=c(2,3))
for (i in 1:length(corrs)) {
  corrs_sub <- corrs[[i]]
  corr.shrink <- CorShrinkVector(corrs_sub$corr, nsamp_vec = corrs_sub$nsam,
                                   optmethod = "mixEM", report_model = TRUE)
  names(corr.shrink$estimate) <- rownames(corrs_sub)
  plot(corr.shrink$model$result$betahat,
       corr.shrink$model$result$PosteriorMean,
       col = 1+as.numeric(corr.shrink$model$result$svalue < .01),
       xlab = "Correlation", ylab = "Shrunken estimate")
  abline(0,1)
  title(names(corrs)[i])
}
```


### Ciruclar correlation 

```{r}
source("../code/corr.cl.R")

corrs.cl <- lapply(1:length(unique(pdata.adj.filt$chip_id)), function(i) {
  
  id <- unique(pdata.adj.filt$chip_id)[i]
  
  log2cpm_sub <- log2cpm.adjust.cycle[, match(rownames(proj.res[[i]][[1]]), colnames(log2cpm.adjust.cycle))]
  
  counts_sub <- counts.cycle[, match(rownames(proj.res[[i]][[1]]), colnames(counts.cycle))]
  
  corrs <- do.call(rbind, lapply(1:nrow(counts_sub), function(g) {
    vec <- cbind(as.numeric(proj.res[[i]][[1]]$rads),
                 log2cpm_sub[g,])
    filt <- counts_sub[g,] > 1
    nsamp <- sum(filt)

    if (nsamp > ncol(counts_sub)/2) {
      vec <- vec[filt,]
      corr <- R2xtCorrCoeff(lvar=vec[,2], cvar=vec[,1])
      nsam <- nrow(vec)
      data.frame(corr=corr, nsam=nsam)
    } else {
      data.frame(corr=NA, nsam=nrow(vec))
    }
    }))
  rownames(corrs) <- rownames(counts_sub)
  return(corrs) 
  })   
names(corrs.cl) <- unique(pdata.adj.filt$chip_id)
```


```{r, fig.height=5, fig.width=7}
par(mfrow=c(2,3))
for (i in 1:length(corrs)) {
  hist(corrs.cl[[i]]$corr, main = names(corrs)[i])            
}
title(main = "Circular-linear correlation", outer = TRUE, line = -1)
```

Compute significance value

```{r}
corrs.cl.sig <- lapply(1:length(unique(pdata.adj.filt$chip_id)), function(i) {
  
  id <- unique(pdata.adj.filt$chip_id)[i]
  
  log2cpm_sub <- log2cpm.adjust.cycle[, match(rownames(proj.res[[i]][[1]]), colnames(log2cpm.adjust.cycle))]
  
  counts_sub <- counts.cycle[, match(rownames(proj.res[[i]][[1]]), colnames(counts.cycle))]
  
  corrs <- do.call(rbind, lapply(1:nrow(counts_sub), function(g) {
    vec <- cbind(as.numeric(proj.res[[i]][[1]]$rads),
                 log2cpm_sub[g,])
    filt <- counts_sub[g,] > 1
    nsamp <- sum(filt)

    if (nsamp > ncol(counts_sub)/2) {
      vec <- vec[filt,]
      corr <- R2xtIndTestRand(lvar=vec[,2], cvar=vec[,1], NR=100)
      nsam <- nrow(vec)
      return(corr)
    } else {
      return(data.frame(corr=NA, pval=NA))
    }
    }))
  rownames(corrs) <- rownames(counts_sub)
  return(corrs) 
  })   
names(corrs.cl.sig) <- unique(pdata.adj.filt$chip_id)
```


```{r, fig.height=5, fig.width=7}
par(mfrow=c(2,3))
for(i in 1:length(corrs.cl.sig)) {
  hist(corrs.cl.sig[[i]]$pval, 
       main = names(corrs.cl.sig)[i])
}
```

Consider significant ones

```{r}
for (i in 1:length(corrs.cl.sig)) {
  ii.sig <- corrs.cl.sig[[i]]$pval < .01
  print(sum(ii.sig, na.rm=TRUE))
}
```

Print some genes

```{r, fig.width=5, fig.height=5}
for (i in 1:length(corrs.cl.sig)) {
  ii.sig <- corrs.cl.sig[[i]]$pval < .01
  
  id <- unique(pdata.adj.filt$chip_id)[i]
  
  log2cpm_sub <- log2cpm.adjust.cycle[, match(rownames(proj.res[[i]][[1]]), colnames(log2cpm.adjust.cycle))]
  
  genes <- rownames(corrs.cl.sig[[i]])
  
  if (i == 5) {numgene <- 3} else {numgene <- 4}
  par(mfrow=c(2,2))
  for (g in 1:numgene) {
    gene <- genes[which(ii.sig)[g]]
    plot(x=as.numeric(proj.res[[i]][[1]]$rads),
         y = log2cpm_sub[rownames(log2cpm_sub) == gene,] ,
         xlab = "Inferred cell time",
         ylab = "log2cpm",
         main = gene)
  }
  title(names(proj.res)[i], outer = TRUE, line = -1)
}
```



---

## Session information

```{r, echo = FALSE}
sessionInfo()
```



