---
title: "Select genes indicative of cell cycle state"
author: "Joyce Hsiao"
output: 
  html_document:
    toc: TRUE
    toc_float: FALSE
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Overview/Results

Assume that cell cycle state is a latent variable. Then here we are interested in evaluating whether FUCCI intensities correlates or predicts cell cycle state, asssuming that DAPI can be employed as a proxy for cell cycle state. 

In addition, I performed some analysis on genes that are highly correlated with both DAPI and FUCCI intensities to see 1) if cell cycle ordering based on cellcycleR correlated with these genes, 2) if cell cycle ordering based on least square fit correlated with these genes, 3) if cell cycle ordering based on these genes correlate with the other two ordering at all.

---

## Data and packages

Packages

```{r}
library(CorShrink)
library(mygene)
```

Load data

```{r}
df <- readRDS(file="../data/eset-filtered.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

# import corrected intensities
pdata.adj <- readRDS("../output/images-normalize-anova.Rmd/pdata.adj.rds")

log2cpm <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.rds")
log2cpm.adjust <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.adjust.rds")

log2cpm <- log2cpm[grep("ENSG", rownames(log2cpm)), 
                   colnames(log2cpm) %in% rownames(pdata.adj)]

log2cpm.adjust <- log2cpm.adjust[grep("ENSG", rownames(log2cpm)),
                   colnames(log2cpm.adjust) %in% rownames(pdata.adj)]

all.equal(rownames(pdata.adj), colnames(log2cpm))

macosko <- readRDS(file = "../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")
pdata.adj.filt <- readRDS(file = "../output/images-circle-ordering.Rmd/pdata.adj.filt.rds")
proj.res <- readRDS(file = "../output/images-circle-ordering.Rmd/proj.res.rds")
```

---

## Correlations

compute correlation between adjusted intensities and log2 expression data.

```{r, fig.width=6, fig.height=5}
corr.rfp <- do.call(rbind, lapply(1:nrow(log2cpm), function(i) {
  vec <- cbind(pdata.adj$rfp.median.log10sum.adjust.ash,
               log2cpm[i,])
  filt <- counts[i,] > 1
  nsamp <- sum(filt)
  if (nsamp > 100) {
#    cnt <- counts[i,filt]
    vec <- vec[filt,]
    corr <- cor(vec[,1], vec[,2])
    nsam <- nrow(vec)
    data.frame(corr=corr, nsam=nsam)
  } else {
    data.frame(corr=NA, nsam=nrow(vec))
  }
  }) )


corr.gfp <- do.call(rbind, lapply(1:nrow(log2cpm), function(i) {
  vec <- cbind(pdata.adj$gfp.median.log10sum.adjust.ash,
               log2cpm[i,])
  filt <- counts[i,] > 1
  nsamp <- sum(filt)
  if (nsamp > 100) {
    vec <- vec[filt,]
    corr <- cor(vec[,1], vec[,2])
    nsam <- nrow(vec)
    data.frame(corr=corr, nsam=nsam)
  } else {
    data.frame(corr=NA, nsam=nrow(vec))
  }
  }) )


corr.dapi <- do.call(rbind, lapply(1:nrow(log2cpm), function(i) {
  vec <- cbind(pdata.adj$dapi.median.log10sum.adjust.ash,
               log2cpm[i,])
  filt <- counts[i,] > 1
  nsamp <- sum(filt)
  if (nsamp > 100) {
    vec <- vec[filt,]
    corr <- cor(vec[,1], vec[,2])
    nsam <- nrow(vec)
    data.frame(corr=corr, nsam=nsam)
  } else {
    data.frame(corr=NA, nsam=nrow(vec))
  }
  }) )
rownames(corr.rfp) <- rownames(log2cpm)
rownames(corr.gfp) <- rownames(log2cpm)
rownames(corr.dapi) <- rownames(log2cpm)


corr.rfp.val <- corr.rfp[!is.na(corr.rfp$corr),]
corr.gfp.val <- corr.gfp[!is.na(corr.gfp$corr),]
corr.dapi.val <- corr.dapi[!is.na(corr.dapi$corr),]


par(mfrow=c(2,2))
hist(corr.rfp.val$corr, main = "correlation between RFP and expression",nclass=50)
abline(v=0, col = "blue")
hist(corr.gfp.val$corr, main = "correlation between GFP and expression",nclass=50)
abline(v=0, col = "blue")
hist(corr.dapi.val$corr, main = "correlation between DAPI and expression",nclass=50)
abline(v=0, col = "blue")
```

Apply CorShrink.

```{r}
corr.rfp.shrink <- CorShrinkVector(corr.rfp.val$corr, nsamp_vec = corr.rfp.val$nsam,
                                   optmethod = "mixEM", report_model = TRUE)
names(corr.rfp.shrink$estimate) <- rownames(corr.rfp.val)

corr.gfp.shrink <- CorShrinkVector(corr.gfp.val$corr, nsamp_vec = corr.gfp.val$nsam,
                                   optmethod = "mixEM", report_model = TRUE)
names(corr.gfp.shrink$estimate) <- rownames(corr.gfp.val)

corr.dapi.shrink <- CorShrinkVector(corr.dapi.val$corr, nsamp_vec = corr.dapi.val$nsam,
                                   optmethod = "mixEM", report_model = TRUE)
names(corr.dapi.shrink$estimate) <- rownames(corr.dapi.val)


par(mfcol=c(2,2))
plot(corr.rfp.val$corr, corr.rfp.shrink$estimate,
     col=1+as.numeric(corr.rfp.shrink$model$result$svalue < .01),
     xlim=c(-.5,.5), ylim=c(-.5,.5), pch=16, cex=.8,
     xlab = "Correlation", ylab = "Shrunken estimate",
     main = "RFP")
abline(0,1)

plot(corr.gfp.val$corr, corr.gfp.shrink$estimate,
     col=1+as.numeric(corr.gfp.shrink$model$result$svalue < .01),
     xlim=c(-.5,.5), ylim=c(-.5,.5), pch=16, cex=.8,
     xlab = "Correlation", ylab = "Shrunken estimate",
     main = "GFP")
abline(0,1)

plot(corr.dapi.val$corr, corr.dapi.shrink$estimate,
     col=1+as.numeric(corr.dapi.shrink$model$result$svalue < .01),
     xlim=c(-.5,.5), ylim=c(-.5,.5), pch=16, cex=.8,
     xlab = "Correlation", ylab = "Shrunken estimate",
     main = "DAPI")
abline(0,1)
title("s-value < .01", outer = TRUE, line = -1)
```

---

## DAPI vs FUCCI

```{r}
macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")

corr.all <- data.frame(genes = rownames(corr.dapi.val),
                       corr.dapi = corr.dapi.shrink$model$result$PosteriorMean,
                       corr.gfp = corr.gfp.shrink$model$result$PosteriorMean,
                       corr.rfp = corr.rfp.shrink$model$result$PosteriorMean,
                       sval.dapi = corr.dapi.shrink$model$result$svalue,
                       sval.gfp = corr.gfp.shrink$model$result$svalue,
                       sval.rfp = corr.rfp.shrink$model$result$svalue)
rownames(corr.all) <- rownames(corr.dapi.val)

corr.all.macosko <- corr.all[which(rownames(corr.all) %in% macosko$ensembl),]
```

```{r fig.width=6, fig.height=6}
par(mfrow=c(2,2))
hist(corr.all$corr.dapi[corr.all$sval.dapi < .01], nclass = 50,
     main = "DAPI vs. expression (sval < .01)",
     xlab = "Pearson correlation")

hist(corr.all$corr.gfp[corr.all$sval.gfp < .01], nclass = 50,
     main = "GFP vs. expression (sval < .01)",
     xlab = "Pearson correlation")

hist(corr.all$corr.rfp[corr.all$sval.rfp < .01], nclass = 50,
     main = "RFP vs. expression (sval < .01)",
     xlab = "Pearson correlation")
```

Of the 469 genes previously annotated as cycle gene, we see that there's about 50% of the genes significantly associated with DAPI that are also associated with FUCCI intenssities (to both GFP and RFP), and vice versa.

```{r fig.width=4, fig.height=4}
library(VennDiagram)
library(grid)
grid.draw(venn.diagram(
  list(DAPI = rownames(corr.all.macosko)[corr.all.macosko$sval.dapi < .01],
       FUCCI = rownames(corr.all.macosko)[corr.all.macosko$sval.gfp < .01 & corr.all.macosko$sval.rfp < .01]),
             filename = NULL))
```

Of genes that are signficantly correlated with both DAPI and Fucci, see how many the expression can be predicted by FUCCI above and beyond DAPI and vice versa.

```{r}
both.genes <- rownames(corr.all.macosko)[corr.all.macosko$sval.dapi < .01 & corr.all.macosko$sval.gfp < .01 & corr.all.macosko$sval.rfp < .01]

fucci.above.dapi <- do.call(c, lapply(1:length(both.genes), function(i) {
  nm <- both.genes[i]
  fit.0 <- lm(log2cpm[which(rownames(log2cpm) %in% nm), ] ~ factor(pdata.adj$chip_id) + pdata.adj$dapi.median.log10sum.adjust.ash - 1)
  fit.1 <- lm(log2cpm[which(rownames(log2cpm) %in% nm), ] ~ factor(pdata.adj$chip_id) + pdata.adj$dapi.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash - 1)
  res <- anova(fit.0, fit.1)
  res$`Pr(>F)`[2]
}) )
names(fucci.above.dapi) <- both.genes


dapi.above.fucci <- do.call(c, lapply(1:length(both.genes), function(i) {
  nm <- both.genes[i]
  fit.0 <- lm(log2cpm[which(rownames(log2cpm) %in% nm), ] ~ factor(pdata.adj$chip_id) + pdata.adj$gfp.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash - 1)
  fit.1 <- lm(log2cpm[which(rownames(log2cpm) %in% nm), ] ~ factor(pdata.adj$chip_id) + pdata.adj$dapi.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash - 1)
  res <- anova(fit.0, fit.1)
  res$`Pr(>F)`[2]
}) )
names(dapi.above.fucci) <- both.genes
```

Of the 21 genes correlated with both DAPI and FUCCI, 6 of them the expression is predicted by FUCCI above and beyond DAPI, while 10 of them the expression is predicted by DAPI above and beyond FUCCI. In addition, 3 of them FUCCI and DAPI and equally important in predicting expression profile. 

```{r}
sum(fucci.above.dapi < .01)
sum(dapi.above.fucci < .01)
```


Of the 7844 genes that we computed correlation between expression and intensities, we see that there's about 50% of the genes significantly associated with DAPI that are also associated with FUCCI intenssities (to both GFP and RFP), and vice versa.

```{r fig.width=4, fig.height=4}
grid.draw(venn.diagram(
  list(DAPI = rownames(corr.all)[corr.all$sval.dapi < .01],
       FUCCI = rownames(corr.all)[corr.all$sval.gfp < .01 & corr.all$sval.rfp < .01]),
             filename = NULL))
```

Of genes that are signficantly correlated with both DAPI and Fucci, see how many the expression can be predicted by FUCCI above and beyond DAPI and vice versa.

```{r}
both.genes <- rownames(corr.all)[corr.all$sval.dapi < .01 & corr.all$sval.gfp < .01 & corr.all$sval.rfp < .01]

fucci.above.dapi <- do.call(c, lapply(1:length(both.genes), function(i) {
  nm <- both.genes[i]
  fit.0 <- lm(log2cpm[which(rownames(log2cpm) %in% nm), ] ~ factor(pdata.adj$chip_id) + pdata.adj$dapi.median.log10sum.adjust.ash - 1)
  fit.1 <- lm(log2cpm[which(rownames(log2cpm) %in% nm), ] ~ factor(pdata.adj$chip_id) + pdata.adj$dapi.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash - 1)
  res <- anova(fit.0, fit.1)
  res$`Pr(>F)`[2]
}) )
names(fucci.above.dapi) <- both.genes


dapi.above.fucci <- do.call(c, lapply(1:length(both.genes), function(i) {
  nm <- both.genes[i]
  fit.0 <- lm(log2cpm[which(rownames(log2cpm) %in% nm), ] ~ factor(pdata.adj$chip_id) + pdata.adj$gfp.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash - 1)
  fit.1 <- lm(log2cpm[which(rownames(log2cpm) %in% nm), ] ~ factor(pdata.adj$chip_id) + pdata.adj$dapi.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash + pdata.adj$gfp.median.log10sum.adjust.ash - 1)
  res <- anova(fit.0, fit.1)
  res$`Pr(>F)`[2]
}) )
names(dapi.above.fucci) <- both.genes
```

Of the 33 genes correlated with both DAPI and FUCCI, 10 of them the expression is predicted by FUCCI above and beyond DAPI, while 13 of them the expression is predicted by DAPI above and beyond FUCCI. In addition, 3 of them FUCCI and DAPI and equally important in predicting expression profile. 

```{r}
sum(fucci.above.dapi < .01)
sum(dapi.above.fucci < .01)
```

```{r}
sum(fucci.above.dapi < .01 & dapi.above.fucci < .01)
```

---

## Save output

```{r, eval = F}
saveRDS(corr.all, file = "../output/images-seq-correlation.Rmd/corr.all.rds")
saveRDS(corr.all.macosko, file = "../output/images-seq-correlation.Rmd/corr.all.macosko.rds")
```


---

## Compare DAPI and expression in their ability to predict FUCCI intensities

The approach is outlined as follows:

1. Find top genes correlated with DAPI (some cutoff)
2. Do PCA on these, and take the first PC
3. Use this first PC to predict GFP and RFP. Compare with using DAPI to predict GFP and RFP.
4. Then you can do the same thing reversing the role of DAPI vs GFP/RFP


```{r}
corr.all <- readRDS("../output/images-seq-correlation.Rmd/corr.all.rds")

log2cpm.dapi <- log2cpm[which(rownames(log2cpm) %in% rownames(corr.all)[corr.all$sval.dapi < .01]), ]

pca.log2cpm.dapi <- prcomp(t(log2cpm.dapi), scale. = FALSE)
(100*(pca.log2cpm.dapi$sdev^2)/sum(pca.log2cpm.dapi$sdev^2))[1:10]

# use first PC from genes correlatd with DAPI to predict FUCCI
fit.rfp.pc1.dapi <- lm(pdata.adj$rfp.median.log10sum.adjust.ash ~ pca.log2cpm.dapi$x[,1] - 1)
fit.gfp.pc1.dapi <- lm(pdata.adj$gfp.median.log10sum.adjust.ash ~ pca.log2cpm.dapi$x[,1] - 1)

fit.rfp.ints.dapi <- lm(pdata.adj$rfp.median.log10sum.adjust.ash ~ pdata.adj$dapi.median.log10sum.adjust.ash - 1)

fit.gfp.ints.dapi <- lm(pdata.adj$gfp.median.log10sum.adjust.ash ~ pdata.adj$dapi.median.log10sum.adjust.ash - 1)



get.loglik.fit <- function(fit) {
  get.loglik <- function(residuals, sigma) {
    residuals.scaled <- -(residuals^2)
    residuals.scaled <- residuals.scaled/2/(sigma^2)
    out <- sum(residuals.scaled - log(sigma) - 0.5*log(2*pi))
    return(out)
  }
  get.loglik(fit$residuals, summary(fit)$sigma)
}


# fit <- fit.dapi.pc.fucci
# residuals <- fit$residuals
# sigma <- summary(fit)$sigma

# make an output table
data.frame(GFP = c(get.loglik.fit(fit.gfp.pc1.dapi), get.loglik.fit(fit.gfp.ints.dapi)),
           RFP = c(get.loglik.fit(fit.rfp.pc1.dapi), get.loglik.fit(fit.rfp.ints.dapi)),
           row.names = c("DAPI PC1 (6%)", "DAPI intensity"))

data.frame(GFP = c(unlist(anova(fit.gfp.pc1.dapi)[5])[1],
                   unlist(anova(fit.gfp.ints.dapi)[5])[1]),
           RFP = c(unlist(anova(fit.rfp.pc1.dapi)[5])[1], 
                   unlist(anova(fit.rfp.ints.dapi)[5])[1]),
           row.names = c("DAPI PC1", "DAPI intensity"))
```


FUCCI predicting DAPI verus expression predicting DAPI.

```{r}
corr.all <- readRDS("../output/images-seq-correlation.Rmd/corr.all.rds")

log2cpm.fucci <- log2cpm[which(rownames(log2cpm) %in% rownames(corr.all)[corr.all$sval.gfp < .01 & corr.all$sval.rfp < .01]), ]

pca.log2cpm.fucci <- prcomp(t(log2cpm.fucci), scale. = FALSE)
(100*(pca.log2cpm.fucci$sdev^2)/sum(pca.log2cpm.fucci$sdev^2))[1:10]

# use first PC from genes correlatd with DAPI to predict FUCCI
fit.dapi.pc.fucci <- lm(pdata.adj$dapi.median.log10sum.adjust.ash ~ pca.log2cpm.fucci$x[,1] -1)

fit.dapi.ints.fucci <- lm(pdata.adj$dapi.median.log10sum.adjust.ash ~ pdata.adj$gfp.median.log10sum.adjust.ash + pdata.adj$rfp.median.log10sum.adjust.ash -1)

# make an output table
data.frame(DAPI = c(get.loglik.fit(fit.dapi.pc.fucci), get.loglik.fit(fit.dapi.ints.fucci)),
           row.names = c("FUCCI PC1 (11%)", "FUCCI intensity"))

data.frame(DAPI = c(unlist(anova(fit.dapi.pc.fucci)[5])[1],
                   unlist(anova(fit.dapi.ints.fucci)[5])[1]),
           row.names = c("FUCCI PC1 (11%)", "FUCCI intensity"))
```

---

## Association between gene expression and projected cell times

There are three clusters in projected cell times (lowest BIC). I then fit spherical regression to find genes that are significant predictors of projected cell times fo each cluster. I only found one signfican genes, which is CDK1. 

**Note**: It's important to consider the assumption that sin(theta) and cos(theta) are uncorrelated. May be unrealistic. 

```{r}
log2cpm.sig.macosko <- log2cpm[which(rownames(log2cpm) %in% rownames(corr.all.macosko)[corr.all.macosko$sval.dapi < .01 & corr.all.macosko$sval.rfp < .01 & corr.all.macosko$sval.gfp < .01]), ]

log2cpm.sig.macosko.filt <- log2cpm.sig.macosko[,match(rownames(pdata.adj.filt), colnames(log2cpm.sig.macosko)) ]

Theta <- do.call(c, lapply(proj.res, function(x) as.numeric(x[[1]]$rads)))
names(Theta) <- do.call(c, lapply(proj.res, function(x) rownames(x[[1]])))

hist(Theta, nclass = 50)
```

Predict cell times by annotated cell cycle gens that are correlated with both DAPI and FUCCI. Only two genes of these 21 show significant association with the projected cell time. 

```{r}
library(Rfast)
x <- data.frame(chip_id = pdata.adj.filt$chip_id,
                t(log2cpm.sig.macosko.filt))
fit <- spml.reg(y=Theta, x=x, seb = TRUE)
round(2*pnorm(fit$be/fit$seb, lower.tail = FALSE),4)
```

Could there be multiple modes in the distribution and hence the terrible prediction?

```{r}
library(movMF)

x <- cbind(sin(Theta), cos(Theta))
fit.2 <- movMF(x, k = 2, nruns = 100)
fit.3 <- movMF(x, k = 3, nruns = 100)
fit.4 <- movMF(x, k = 4, nruns = 100)

sapply(list(fit.2, fit.3, fit.4), BIC)
```

For each cluster, fit spml. 

```{r}
library(ashr)
fit.3.predict <- predict(fit.3)
par(mfrow =c(2,2))
hist(Theta[fit.3.predict == 1], xlim = c(0,2*pi), main = "Cluster 1")
hist(Theta[fit.3.predict == 2], xlim = c(0,2*pi), main = "Cluster 2")
hist(Theta[fit.3.predict == 3], xlim = c(0,2*pi), main = "Cluster 3")

x <- data.frame(chip_id = pdata.adj.filt$chip_id,
                t(log2cpm.sig.macosko.filt))

fit.clust1 <- spml.reg(y=Theta[which(fit.3.predict == 1)], 
                       x=x[which(fit.3.predict == 1), ], seb = TRUE)
fit.clust1.sval <- cbind(ash(fit.clust1$be[,1],fit.clust1$seb[,1])$result$svalue,
                         ash(fit.clust1$be[,2],fit.clust1$seb[,2])$result$svalue)

fit.clust2 <- spml.reg(y=Theta[which(fit.3.predict == 2)], 
                       x=x[which(fit.3.predict == 2), ], seb = TRUE)
fit.clust2.sval <- cbind(ash(fit.clust2$be[,1],fit.clust2$seb[,1])$result$svalue,
                         ash(fit.clust2$be[,2],fit.clust2$seb[,2])$result$svalue)

fit.clust3 <- spml.reg(y=Theta[which(fit.3.predict == 3)], 
                       x=x[which(fit.3.predict == 3), ], seb = TRUE)
fit.clust3.sval <- cbind(ash(fit.clust3$be[,1],fit.clust3$seb[,1])$result$svalue,
                         ash(fit.clust3$be[,2],fit.clust3$seb[,2])$result$svalue)

cbind(round(fit.clust1.sval,4),
      round(fit.clust2.sval,4),
      round(fit.clust3.sval,4))
```

None is signifcantly associated with sin(theta). But cdk1 for the third cluster (cyclin dependent kinase 1) is significantly associated with all three clusters (sval < .01).

```{r}
rownames(fit.clust3$be)[24]
```

Check how different the effect sizes are

```{r}
round(cbind(fit.clust1$be, fit.clust2$be, fit.clust3$be),4)
```


Correlation between RFP and GFP.

```{r}
plot(x=pdata.adj.filt$gfp.median.log10sum.adjust,
     y=pdata.adj.filt$rfp.median.log10sum.adjust,
     xlab = "GFP", ylab = "RFP")
```

---

## Session information

```{r, echo = FALSE}
sessionInfo()
```



