---
title: "Quantile-normalize intensity measurements"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Introduction

Previously when investigating measurement variation (GFP/RFP/FAPI), we learned that there's significant variation between batches in the distributions of background-corrected pixel intensites. See [here](../docs/images-qc-labels.html).

$~$

**Approach:**

In this document, I apply quantile normalization to intensity measurements on a per-channel basis. The approach is as follows

1. Construct a reference intensity. Estimate k-quantiles of reference intensities $Q^{R,k}= \big( {q^{R,k}_{[1]}, \dots, q^{R,k}_{[n_k]}} \big)$.

2. For each plate $i$, estimate $l$-quantiles of intensities on a per-plate basis $Q^{i,l}= \big( {q^{i,l}_{[1]}, \dots, q^{i,l}_{[n_l]}} \big)$. 

3. For each plate $i$, compare the intensity value $F_{ij}$ with the quantile values $\big( {q^{i,l}_{[1]}, \dots, q^{i,l}_{[n_l]}} \big)$ and assign image/well $j$ with the quantile that has the closest intensity value, say $q^{i,l}_{[m]}$ if $m= argmin_{(1,\dots,n_l)} |F_{ij}- q^{i,l}_{[n_l]}|$. Then subsitute the intensity value with the $m$-th quantile value in the reference intensity $q^{R,k}_{[m]}$. 

$~$

I tried two methods for constructing reference intensity vector, and the results are vastly different depends on the method that we choose. 

*Method 1*: Aggregate intensity values aross plates.

*Method 2*: Take the average of $n$-quantiles across plates. 

$~$

**Results:**

1. We chose 1/.005 quantiles for all three channels. See the document for our exploratory analysis of intensities from all three channels.

2. Method 1 versus Method 2: Because in Method 1, the distribution of Green/Red is more dense toward low and high-valued intensities, we see that the normalized values are closer toward the boundaries. 

3. Method 2 of constructing the reference produces better results. We see that the relationship between Green/Red is preserved before versus after normalization. 

4. In Method 2, after normalization, the range of intensities is the same between plates for each of the three channels (Green, Red, DAPI). As a result, many of the images/wells with low intensties decreased in intensity values.

5. Because of 4, the distances between samples in many plates increase rather than decrease. We were looking for decrease in the distances between samples, i.e., tighter clusters or smaller within-cluster distance...


---

## Load packages and data

```{r, message=FALSE, warning=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(cowplot)
library(wesanderson)
library(RColorBrewer)
library(Biobase)

plates <- c("18511_18855","18855_19101","18855_19160","18870_18511",
            "18870_18855","18870_19101","18870_19160","19098_18511",
            "19098_18870","19098_19160","19101_18511","19101_19098",
            "19160_18870","19101_19160","19160_18511", "18855_19098")

##Combine intensity stats from different plates.
ints <- do.call(rbind, lapply(1:length(plates), function(index) {
  tmp <- readRDS(paste0("/project2/gilad/fucci-seq/intensities_stats/",plates[index],".stats.rds"))
  tmp <- data.frame(plate=plates[index], 
                    image=as.character(droplevels(tmp$wellID)),
    rfp.back.zoom.mean=tmp$rfp.back.zoom.mean,
    rfp.back.zoom.median=tmp$rfp.back.zoom.q05,
    gfp.back.zoom.mean=tmp$gfp.back.zoom.mean,
    gfp.back.zoom.median=tmp$gfp.back.zoom.q05,
    dapi.back.zoom.mean=tmp$dapi.back.mean,
    dapi.back.zoom.median=tmp$dapi.back.q05,
    rfp.mean.log10sum=with(tmp, log10((rfp.fore.zoom.mean-rfp.back.zoom.mean)*size.zoom)),
    gfp.mean.log10sum=with(tmp, log10((gfp.fore.zoom.mean-gfp.back.zoom.mean)*size.zoom)),
    dapi.mean.log10sum=with(tmp, log10((dapi.fore.mean-dapi.back.mean)*size)),
    rfp.median.log10sum=with(tmp, log10((rfp.fore.zoom.mean-rfp.back.zoom.q05)*size.zoom)),
    gfp.median.log10sum=with(tmp, log10((gfp.fore.zoom.mean-gfp.back.zoom.q05)*size.zoom)),
    dapi.median.log10sum=with(tmp, log10((dapi.fore.mean-dapi.back.q05)*size)) )

  tmp$rfp.mean.log10sum[which(tmp$rfp.mean.log10sum == "NaN")] <- min(tmp$rfp.mean.log10sum, na.rm=TRUE)
  tmp$gfp.mean.log10sum[which(tmp$gfp.mean.log10sum == "NaN")] <- min(tmp$gfp.mean.log10sum, na.rm=TRUE)
  tmp$dapi.mean.log10sum[which(tmp$dapi.mean.log10sum == "NaN")] <- min(tmp$dapi.mean.log10sum, na.rm=TRUE)
  tmp$rfp.median.log10sum[which(tmp$rfp.median.log10sum == "NaN")] <- min(tmp$rfp.median.log10sum, na.rm=TRUE)
  tmp$gfp.median.log10sum[which(tmp$gfp.median.log10sum == "NaN")] <- min(tmp$gfp.median.log10sum, na.rm=TRUE)
  tmp$dapi.median.log10sum[which(tmp$dapi.median.log10sum == "NaN")] <- min(tmp$dapi.median.log10sum, na.rm=TRUE)

  return(tmp)  
}) )
#saveRDS(ints, "/project2/gilad/fucci-seq/intensities_matrix2stats/ints.rds")

##Load the expression set info.
eset_fls <- list.files("../data/eset", full.names=TRUE)

anno <- do.call(rbind, lapply(1:length(eset_fls), function(index) {
  eset_index <- readRDS(eset_fls[[index]])
  pdata_index <- pData(eset_index)
  return(pdata_index)
}))

# make unique id in both
ints$unique <- paste0(ints$plate,"_",as.character(ints$image))
anno$unique <- paste0(anno$image_individual,"_", sprintf("%05d",as.numeric(anno$image_label)))

# subset annotation file
subset_index <- which(anno$unique %in% ints$unique)
anno_subset <- anno[subset_index,]

# subset intensities
ints_subset_index <- match(anno_subset$unique, ints$unique)
ints_subset <- ints[ints_subset_index,]

#all.equal(anno_subset$unique, ints_subset$unique)
ints_subset$chip_id <- anno_subset$chip_id
```

---

$~$

## Densit plots

$~$

First, look at the distribution of all batches combined versus each batch. 

$~$

```{r, echo = FALSE}
plot_grid(
  ggplot(ints_subset, aes(x=rfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=rfp.median.log10sum, col = as.factor(plate))) +
    labs(title = "RFP by plate",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Red channel log10 pixel sum", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_subset, aes(x=gfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=gfp.median.log10sum, col = as.factor(plate))) +
    labs(title = "GFP by plate",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Green channel log10 pixel sum", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_subset, aes(x=dapi.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=dapi.median.log10sum, col = as.factor(plate))) +
    labs(title = "DAPI by plate",
         subtitle = "Black line indicates distribution aggregating plates",
         x="DAPI channel log10 pixel sum", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

---

$~$

## Normalization

$~$

Code for one single sample

$~$

```{r}
my_quantnorm <- function(reference, sample, span=.01) {
  # quantiles for intensities all samples across plates
  quants_reference <- quantile(reference, probs=seq(0,1,span))
  # intensities for a given plate
  # quantiles for intensities at each plate
  quants_sample <- quantile(sample, probs=seq(0,1,span))
  # empty vector for normalized values
  sample_normed <- vector("numeric", length=length(sample))
  
  for (index in 1:length(sample)) {
    # for each sample, find the closest sample quantile
    sample_order <- names(which.min(abs(sample[index]-quants_sample)))
    # # get the reference intensity value of the closet quantile
    ref_order_value <- reference[which(names(quants_reference)==sample_order)] 
    # assign the reference intensity value to the sample
    sample_normed[index] <- ref_order_value 
  }
  return(sample_normed)
}
```

$~$

### Choosing quantile interval

$~$

**RFP**

$~$

```{r, echo = FALSE}
ints_subset$plate <- as.character(ints_subset$plate)
ints_rfp <- data.frame(plate=ints_subset$plate,
                       rfp.median.log10sum = ints_subset$rfp.median.log10sum)
ints_rfp$span1 <- NULL
ints_rfp$span2 <- NULL
ints_rfp$span3 <- NULL
ints_rfp$span4 <- NULL

for (index in 1:length(unique(ints_rfp$plate))) {
  plate <- as.character(unique(ints_rfp$plate))[index]
#  print(plate)
  tmp <- my_quantnorm(reference = ints_rfp$rfp.median.log10sum,
                      sample = ints_rfp$rfp.median.log10sum[ints_rfp$plate == plate],
                      span=.02)
  ints_rfp$span1[ints_rfp$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_rfp$rfp.median.log10sum,
                      sample = ints_rfp$rfp.median.log10sum[ints_rfp$plate == plate],
                      span=.01)
  ints_rfp$span2[ints_rfp$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = ints_rfp$rfp.median.log10sum,
                      sample = ints_rfp$rfp.median.log10sum[ints_rfp$plate == plate],
                      span=.005)
  ints_rfp$span3[ints_rfp$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_rfp$rfp.median.log10sum,
                    sample = ints_rfp$rfp.median.log10sum[ints_rfp$plate == plate],
                    span=.001)
  ints_rfp$span4[ints_rfp$plate == plate] <- tmp
}

plot_grid(
  ggplot(ints_rfp, aes(x=rfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span1, col = as.factor(plate))) +
    labs(title = "Span=.02",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_rfp, aes(x=rfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span2, col = as.factor(plate))) +
    labs(title = "Span=.01",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_rfp, aes(x=rfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span3, col = as.factor(plate))) +
    labs(title = "Span=.005",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_rfp, aes(x=rfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span4, col = as.factor(plate))) +
    labs(title = "Span=.001",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

$~$

**GFP**

$~$

```{r, echo = FALSE}
ints_subset$plate <- as.character(ints_subset$plate)
ints_gfp <- data.frame(plate=ints_subset$plate,
                       gfp.median.log10sum = ints_subset$gfp.median.log10sum)
ints_gfp$span1 <- NULL
ints_gfp$span2 <- NULL
ints_gfp$span3 <- NULL
ints_gfp$span4 <- NULL

for (index in 1:length(unique(ints_gfp$plate))) {
  plate <- unique(ints_gfp$plate)[index]
#  print(plate)
  tmp <- my_quantnorm(reference = ints_gfp$gfp.median.log10sum,
                      sample = ints_gfp$gfp.median.log10sum[ints_gfp$plate == plate],
                      span=.02)
  ints_gfp$span1[ints_gfp$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_gfp$gfp.median.log10sum,
                      sample = ints_gfp$gfp.median.log10sum[ints_gfp$plate == plate],
                      span=.01)
  ints_gfp$span2[ints_gfp$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = ints_gfp$gfp.median.log10sum,
                      sample = ints_gfp$gfp.median.log10sum[ints_gfp$plate == plate],
                      span=.005)
  ints_gfp$span3[ints_gfp$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_gfp$gfp.median.log10sum,
                    sample = ints_gfp$gfp.median.log10sum[ints_gfp$plate == plate],
                    span=.001)
  ints_gfp$span4[ints_gfp$plate == plate] <- tmp
}

plot_grid(
  ggplot(ints_gfp, aes(x=gfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span1, col = as.factor(plate))) +
    labs(title = "Span=.02",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_gfp, aes(x=gfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span2, col = as.factor(plate))) +
    labs(title = "Span=.01",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_gfp, aes(x=gfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span3, col = as.factor(plate))) +
    labs(title = "Span=.005",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_gfp, aes(x=gfp.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span4, col = as.factor(plate))) +
    labs(title = "Span=.001",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

$~$

**DAPI**

$~$

```{r, echo = FALSE}
ints_subset$plate <- as.character(ints_subset$plate)
ints_dapi <- data.frame(plate=ints_subset$plate,
                       dapi.median.log10sum = ints_subset$dapi.median.log10sum)
ints_dapi$span1 <- NULL
ints_dapi$span2 <- NULL
ints_dapi$span3 <- NULL
ints_dapi$span4 <- NULL

for (index in 1:length(unique(ints_dapi$plate))) {
  plate <- unique(ints_dapi$plate)[index]
#  print(plate)
  tmp <- my_quantnorm(reference = ints_dapi$dapi.median.log10sum,
                      sample = ints_dapi$dapi.median.log10sum[ints_dapi$plate == plate],
                      span=.02)
  ints_dapi$span1[ints_dapi$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_dapi$dapi.median.log10sum,
                      sample = ints_dapi$dapi.median.log10sum[ints_dapi$plate == plate],
                      span=.01)
  ints_dapi$span2[ints_dapi$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = ints_dapi$dapi.median.log10sum,
                      sample = ints_dapi$dapi.median.log10sum[ints_dapi$plate == plate],
                      span=.005)
  ints_dapi$span3[ints_dapi$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_dapi$dapi.median.log10sum,
                    sample = ints_dapi$dapi.median.log10sum[ints_dapi$plate == plate],
                    span=.001)
  ints_dapi$span4[ints_dapi$plate == plate] <- tmp
}

plot_grid(
  ggplot(ints_dapi, aes(x=dapi.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span1, col = as.factor(plate))) +
    labs(title = "Span=.02",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_dapi, aes(x=dapi.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span2, col = as.factor(plate))) +
    labs(title = "Span=.01",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_dapi, aes(x=dapi.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span3, col = as.factor(plate))) +
    labs(title = "Span=.005",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_dapi, aes(x=dapi.median.log10sum), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span4, col = as.factor(plate))) +
    labs(title = "Span=.001",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

$~$

### Method 1

Method 1 constructs a vector of refernece intensity by aggregating all image intensity values across plates. 

$~$

```{r, echo = FALSE}
#DAPI choose .005; Green choose .005; Red choose .005
ints_subset$plate <- as.character(ints_subset$plate)
ints_normed <- data.frame(plate=ints_subset$plate,
                          chip_id=ints_subset$chip_id,
                       dapi.median.log10sum = ints_subset$dapi.median.log10sum,
                       rfp.median.log10sum = ints_subset$rfp.median.log10sum,
                       gfp.median.log10sum = ints_subset$gfp.median.log10sum)
ints_normed$rfp.normed <- NULL
ints_normed$gfp.normed <- NULL
ints_normed$dapi.normed <- NULL

for (index in 1:length(unique(ints_normed$plate))) {
  plate <- unique(ints_normed$plate)[index]

  tmp <- my_quantnorm(reference = ints_normed$dapi.median.log10sum,
                      sample = ints_normed$dapi.median.log10sum[ints_normed$plate == plate],
                      span=.005)
  ints_normed$dapi.normed[ints_normed$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = ints_normed$rfp.median.log10sum,
                      sample = ints_normed$rfp.median.log10sum[ints_normed$plate == plate],
                      span=.005)
  ints_normed$rfp.normed[ints_normed$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_normed$gfp.median.log10sum,
                    sample = ints_normed$gfp.median.log10sum[ints_normed$plate == plate],
                    span=.005)
  ints_normed$gfp.normed[ints_normed$plate == plate] <- tmp
}
```

$~$

**Distribution of the reference**

$~$

```{r, fig.width=5, fig.height=5, echo = FALSE}
ggplot(ints_normed, aes(x=rfp.median.log10sum,y=gfp.median.log10sum)) +
  geom_point(alpha = .5, cex = .7) + 
#    facet_wrap(~as.factor(chip_id), ncol=4) + 
  labs(title = "All plates raw intensities",
       x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none")
```

$~$

**After normalization**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
plot_grid(
  ggplot(ints_normed, aes(x=gfp.normed), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=gfp.normed, col = as.factor(plate))) +
    labs(title = "GFP",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed GFP intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_normed, aes(x=rfp.normed), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=rfp.normed, col = as.factor(plate))) +
    labs(title = "RFP",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed RFP intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_normed, aes(x=dapi.normed), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=dapi.normed, col = as.factor(plate))) +
    labs(title = "DAPI",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed DAPI intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

$~$

**Green versus Red intensties by plate, labeled by DAPI**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
ints_normed <- ints_normed %>% 
  mutate(dapi_normed_4quant=ntile(dapi.normed,4),
         dapi_normed_3quant=ntile(dapi.normed,3),
         dapi_4quant=ntile(dapi.median.log10sum, 4),
         dapi_3quant=ntile(dapi.median.log10sum, 3))

plot_grid(
  ggplot(ints_normed, aes(x=rfp.median.log10sum,
                          y=gfp.median.log10sum, 
                          col = as.factor(dapi_3quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(title = "Raw intensities",
         subtitle = "labedl by DAPI 3-quantile before normalization",
         x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none"),  
  ggplot(ints_normed, aes(x=rfp.normed,
                          y=gfp.normed, 
                          col = as.factor(dapi_normed_3quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(title = "Normed intensities",
         subtitle = "labeled by DAPI 3-quantiles after normalization",
         x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none") )
```

$~$

**Green versus Red intensties by individual, labeled by DAPI**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
ints_normed <- ints_normed %>% 
  mutate(dapi_normed_4quant=ntile(dapi.normed,4),
         dapi_normed_3quant=ntile(dapi.normed,3),
         dapi_4quant=ntile(dapi.median.log10sum, 4),
         dapi_3quant=ntile(dapi.median.log10sum, 3))

plot_grid(
  ggplot(ints_normed, aes(x=rfp.median.log10sum,
                          y=gfp.median.log10sum, 
                          col = as.factor(dapi_3quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(title = "Raw intensities",
         subtitle = "labedl by DAPI 3-quantile before normalization",
         x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none"),  
  ggplot(ints_normed, aes(x=rfp.normed,
                          y=gfp.normed, 
                          col = as.factor(dapi_normed_3quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(title = "Normed intensities",
         subtitle = "labeled by DAPI 3-quantiles after normalization",
         x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none") )
```

$~$

**Green**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
plot_grid(
  ggplot(ints_normed, aes(x=gfp.median.log10sum, y=gfp.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "GFP by plate") + 
    theme_gray() + theme(legend.position="none"), 
  ggplot(ints_normed, aes(x=gfp.median.log10sum, y=gfp.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "GFP by individual") + 
    theme_gray() + theme(legend.position="none"),
  ncol=2)
```

$~$

**Red**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
plot_grid(
  ggplot(ints_normed, aes(x=rfp.median.log10sum, y=rfp.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "RFP by plate") + 
    theme_gray() + theme(legend.position="none"), 
  ggplot(ints_normed, aes(x=rfp.median.log10sum, y=rfp.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "RFP by individual") + 
    theme_gray() + theme(legend.position="none"),
  ncol=2)
```

$~$

**DAPI**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
plot_grid(
  ggplot(ints_normed, aes(x=dapi.median.log10sum, y=dapi.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "DAPI by plate") + 
    theme_gray() + theme(legend.position="none"), 
  ggplot(ints_normed, aes(x=dapi.median.log10sum, y=dapi.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "DAPI by individual") + 
    theme_gray() + theme(legend.position="none"),
  ncol=2)
```

---

$~$

### Method 2

$~$

Reference intensity vector: average of quantile values across plates.

$~$

```{r, echo = FALSE}
# constructs reference
quant_reference_rfp <- rowMeans(do.call(cbind, lapply(1:length(unique(as.character(ints_subset$plate))), function(index) {
  plate <- unique(as.character(ints_subset$plate))[index]
  quantile(ints_subset$rfp.median.log10sum[ints_subset$plate == plate], prob=seq(0,1,.005))
})))
quant_reference_gfp <- rowMeans(do.call(cbind, lapply(1:length(unique(as.character(ints_subset$plate))), function(index) {
  plate <- unique(as.character(ints_subset$plate))[index]
  quantile(ints_subset$gfp.median.log10sum[ints_subset$plate == plate], prob=seq(0,1,.005))
})))
quant_reference_dapi <- rowMeans(do.call(cbind, lapply(1:length(unique(as.character(ints_subset$plate))), function(index) {
  plate <- unique(as.character(ints_subset$plate))[index]
  quantile(ints_subset$dapi.median.log10sum[ints_subset$plate == plate], prob=seq(0,1,.005))
})))


#DAPI choose .005; Green choose .005; Red choose .005
ints_normed_2 <- data.frame(plate=ints_subset$plate,
                          chip_id=ints_subset$chip_id,
                       dapi.median.log10sum = ints_subset$dapi.median.log10sum,
                       rfp.median.log10sum = ints_subset$rfp.median.log10sum,
                       gfp.median.log10sum = ints_subset$gfp.median.log10sum)
ints_normed_2$rfp.normed <- NULL
ints_normed_2$gfp.normed <- NULL
ints_normed_2$dapi.normed <- NULL 

for (index in 1:length(unique(ints_normed_2$plate))) {
  plate <- unique(ints_normed_2$plate)[index]

  tmp <- my_quantnorm(reference = quant_reference_dapi,
                      sample = ints_normed_2$dapi.median.log10sum[ints_normed_2$plate == plate],
                      span=.005)
  ints_normed_2$dapi.normed[ints_normed_2$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = quant_reference_rfp,
                      sample = ints_normed_2$rfp.median.log10sum[ints_normed_2$plate == plate],
                      span=.005)
  ints_normed_2$rfp.normed[ints_normed_2$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = quant_reference_gfp,
                    sample = ints_normed_2$gfp.median.log10sum[ints_normed_2$plate == plate],
                    span=.005)
  ints_normed_2$gfp.normed[ints_normed_2$plate == plate] <- tmp
}
```

$~$

**After normalization**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
plot_grid(
  ggplot(ints_normed_2, aes(x=gfp.normed, col=as.factor(plate))) + 
    geom_density(alpha = .5, cex = .7) + 
    labs(title = "GFP",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed GFP intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_normed_2, aes(x=rfp.normed, col=as.factor(plate))) + 
    geom_density(alpha = .5, cex = .7) + 
    labs(title = "RFP",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed RFP intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_normed_2, aes(x=dapi.normed, col=as.factor(plate))) + 
    geom_density(alpha = .5, cex = .7) + 
    labs(title = "DAPI",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed DAPI intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

$~$

**Green versus Red intensties by plate, labeled by DAPI**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
ints_normed_2 <- ints_normed_2 %>% 
  mutate(dapi_normed_4quant=ntile(dapi.normed,4),
         dapi_normed_3quant=ntile(dapi.normed,3),
         dapi_4quant=ntile(dapi.median.log10sum, 4),
         dapi_3quant=ntile(dapi.median.log10sum, 3))

plot_grid(
  ggplot(ints_normed_2, aes(x=rfp.median.log10sum,
                          y=gfp.median.log10sum, 
                          col = as.factor(dapi_3quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(title = "Raw intensities",
         subtitle = "labedl by DAPI 3-quantile before normalization",
         x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none"),  
  ggplot(ints_normed_2, aes(x=rfp.normed,
                          y=gfp.normed, 
                          col = as.factor(dapi_normed_3quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(title = "Normed intensities",
         subtitle = "labeled by DAPI 3-quantiles after normalization",
         x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none") )
```

$~$

**Green versus Red intensties by individual, labeled by DAPI**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
ints_normed_2 <- ints_normed_2 %>% 
  mutate(dapi_normed_4quant=ntile(dapi.normed,4),
         dapi_normed_3quant=ntile(dapi.normed,3),
         dapi_4quant=ntile(dapi.median.log10sum, 4),
         dapi_3quant=ntile(dapi.median.log10sum, 3))

plot_grid(
  ggplot(ints_normed_2, aes(x=rfp.median.log10sum,
                          y=gfp.median.log10sum, 
                          col = as.factor(dapi_3quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(title = "Raw intensities",
         subtitle = "labedl by DAPI 3-quantile before normalization",
         x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none"),  
  ggplot(ints_normed_2, aes(x=rfp.normed,
                          y=gfp.normed, 
                          col = as.factor(dapi_normed_3quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(title = "Normed intensities",
         subtitle = "labeled by DAPI 3-quantiles after normalization",
         x="RFP", y = "GFP") + theme_gray()+ theme(legend.position="none") )
```

$~$

**Green**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
plot_grid(
  ggplot(ints_normed_2, aes(x=gfp.median.log10sum, y=gfp.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "GFP by plate") + 
    theme_gray() + theme(legend.position="none"), 
  ggplot(ints_normed_2, aes(x=gfp.median.log10sum, y=gfp.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "GFP by individual") + 
    theme_gray() + theme(legend.position="none"),
  ncol=2)
```

$~$

**Red**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
plot_grid(
  ggplot(ints_normed_2, aes(x=rfp.median.log10sum, y=rfp.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "RFP by plate") + 
    theme_gray() + theme(legend.position="none"), 
  ggplot(ints_normed_2, aes(x=rfp.median.log10sum, y=rfp.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "RFP by individual") + 
    theme_gray() + theme(legend.position="none"),
  ncol=2)
```

$~$

**DAPI**

$~$

```{r, fig.width=12, fig.height=5, echo = FALSE}
plot_grid(
  ggplot(ints_normed_2, aes(x=dapi.median.log10sum, y=dapi.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(plate), ncol=4) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "DAPI by plate") + 
    theme_gray() + theme(legend.position="none"), 
  ggplot(ints_normed_2, aes(x=dapi.median.log10sum, y=dapi.normed)) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(x="Raw intensity", y = "Normed intensity", title = "DAPI by individual") + 
    theme_gray() + theme(legend.position="none"),
  ncol=2)
```




---

## Session information

```{r, echo=FALSE}
sessionInfo()
```

