---
title: "Circle fit to intensities: follow-up"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Goals

1. Estimate positions based on PCA

2. Correlation positions with genes and DAPI

---

## Data and packages

Packages

```{r}
library(circular)
library(conicfit)
library(Biobase)
library(dplyr)
library(matrixStats)
library(CorShrink)
```

Load data

```{r}
df <- readRDS(file="../data/eset-filtered.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.rds")
# log2cpm.adjust <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.adjust.rds")

# import corrected intensities
pdata.adj <- readRDS("../output/images-normalize-anova.Rmd/pdata.adj.rds")

macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")
```


---

## Projected normal on PCs of Red/Green


```{r, fig.width=7, fig.height=5}
pc.fucci <- prcomp(subset(pdata.adj, 
                          select=c("rfp.median.log10sum.adjust",
                                   "gfp.median.log10sum.adjust")),
                   center = T, scale. = T)
```

The first two PC are about the sample in explaining variation.

```{r}
pc.fucci$sdev[1:2]
```

Green and red are similarly correlated with PC1 and PC2 across individuals. 

```{r}
pcs.rfp <- sapply(1:2, function(i) {
  res <- summary(lm(pc.fucci$x[,i]~pdata.adj$rfp.median.log10sum.adjust))
  res$adj.r.squared
})
pcs.rfp

pcs.gfp <- sapply(1:2, function(i) {
  res <- summary(lm(pc.fucci$x[,i]~pdata.adj$gfp.median.log10sum.adjust))
  res$adj.r.squared
})
pcs.gfp
```

```{r, fig.width=7, fig.height=7}
library(circular)
Theta.fucci <- coord2rad(pc.fucci$x)

par(mfrow=c(2,2), mar=c(4,4,3,1))
plot(circular(Theta.fucci), stack = TRUE)
plot(as.numeric(Theta.fucci), pdata.adj$rfp.median.log10sum.adjust,
     col = "firebrick", pch = 16, cex = .7,
     xlab = "Theta (projected cell time)", 
     ylab = "RFP (log10 intensity)")
plot(as.numeric(Theta.fucci), pdata.adj$gfp.median.log10sum.adjust,
     col = "forestgreen", pch = 16, cex = .7,
     xlab = "Theta (projected cell time)", 
     ylab = "GFP (log10 intensity)")
plot(as.numeric(Theta.fucci), pdata.adj$gfp.median.log10sum.adjust,
     col = "mediumblue", pch = 16, cex = .7,
     xlab = "Theta (projected cell time)", 
     ylab = "DAPI (log10 intensity)")
```

Correlation with DAPI, RFP and GFP.

```{r}
library(Rfast)
library(circular)

# normalize intensities to between 0 to 2*pi
pdata.adj$dapi.circ <- with(pdata.adj, {
  normed <- (dapi.median.log10sum.adjust-min(dapi.median.log10sum.adjust))/(max(dapi.median.log10sum.adjust)- min(dapi.median.log10sum.adjust))
  normed*2*pi
} )
pdata.adj$gfp.circ <- with(pdata.adj, {
  normed <- (gfp.median.log10sum-min(gfp.median.log10sum.adjust))/(max(gfp.median.log10sum.adjust)- min(gfp.median.log10sum.adjust))
  normed*2*pi
} )
pdata.adj$rfp.circ <- with(pdata.adj, {
  normed <- (rfp.median.log10sum.adjust-min(rfp.median.log10sum.adjust))/(max(rfp.median.log10sum.adjust)- min(rfp.median.log10sum.adjust))
  normed*2*pi
} )


cor.dapi.cl <- sqrt(circlin.cor(as.numeric(Theta.fucci), pdata.adj$dapi.median.log10sum.adjust)[1])
cor.gfp.cl <- sqrt(circlin.cor(as.numeric(Theta.fucci), pdata.adj$gfp.median.log10sum.adjust)[1])
cor.rfp.cl <- sqrt(circlin.cor(as.numeric(Theta.fucci), pdata.adj$rfp.median.log10sum.adjust)[1])

cor.dapi.cc <- cor.circular(as.numeric(Theta.fucci), pdata.adj$dapi.median.log10sum.adjust)
cor.gfp.cc <- cor.circular(as.numeric(Theta.fucci), pdata.adj$gfp.median.log10sum.adjust)
cor.rfp.cc <- cor.circular(as.numeric(Theta.fucci), pdata.adj$rfp.median.log10sum.adjust)

out <- data.frame(cbind(rbind(cor.dapi.cl, cor.gfp.cl, cor.rfp.cl),
                 rbind(cor.dapi.cc, cor.gfp.cc, cor.rfp.cc)),
           row.names = c("DAPI", "GFP", "RFP"))
colnames(out) <- c("cor.circ.linear", "cor.circ.circ")
print(out)
```


The high correlation of GFP and RFP in the first column follows the use of PCs of GFP and RFP to construct cell time. Results in the second column seem interesting. The values capture the relationship between these measures in the plot. GFP and DAPI are negatively associated with cell time and RFP is positive associated with cell time.



Correlation with DAPI, RFP and GFP.

```{r}
library(Rfast)
library(circular)

# normalize intensities to between 0 to 2*pi
pdata.adj$dapi.circ <- with(pdata.adj, {
  normed <- (dapi.median.log10sum.adjust-min(dapi.median.log10sum.adjust))/(max(dapi.median.log10sum.adjust)- min(dapi.median.log10sum.adjust))
  normed*2*pi
} )
pdata.adj$gfp.circ <- with(pdata.adj, {
  normed <- (gfp.median.log10sum-min(gfp.median.log10sum.adjust))/(max(gfp.median.log10sum.adjust)- min(gfp.median.log10sum.adjust))
  normed*2*pi
} )
pdata.adj$rfp.circ <- with(pdata.adj, {
  normed <- (rfp.median.log10sum.adjust-min(rfp.median.log10sum.adjust))/(max(rfp.median.log10sum.adjust)- min(rfp.median.log10sum.adjust))
  normed*2*pi
} )


cor.dapi.cl <- sqrt(circlin.cor(as.numeric(Theta.fucci), pdata.adj$dapi.median.log10sum.adjust)[1])
cor.gfp.cl <- sqrt(circlin.cor(as.numeric(Theta.fucci), pdata.adj$gfp.median.log10sum.adjust)[1])
cor.rfp.cl <- sqrt(circlin.cor(as.numeric(Theta.fucci), pdata.adj$rfp.median.log10sum.adjust)[1])

cor.dapi.cc <- cor.circular(as.numeric(Theta.fucci), pdata.adj$dapi.median.log10sum.adjust)
cor.gfp.cc <- cor.circular(as.numeric(Theta.fucci), pdata.adj$gfp.median.log10sum.adjust)
cor.rfp.cc <- cor.circular(as.numeric(Theta.fucci), pdata.adj$rfp.median.log10sum.adjust)

out <- data.frame(cbind(rbind(cor.dapi.cl, cor.gfp.cl, cor.rfp.cl),
                 rbind(cor.dapi.cc, cor.gfp.cc, cor.rfp.cc)),
           row.names = c("DAPI", "GFP", "RFP"))
colnames(out) <- c("cor.circ.linear", "cor.circ.circ")
print(out)
```

---

## Circular-circular correlation between projected time and gene expression


```{r, fig.width=7, fig.height=4}
library(Rfast)
library(circular)

counts.log2cpm <- log10((10^6)*t(t(counts)/colSums(counts))+1)
counts.log2cpm.normed <- 2*pi*(counts.log2cpm)/max(counts.log2cpm)

cor.cc.log2cpm <- sapply(1:nrow(counts.log2cpm), function(g) {
  cor.circular(counts.log2cpm.normed[g,],Theta.fucci)
})
nsam.nonzero <- rowSums(counts.log2cpm > 0)

library(CorShrink)
corshrink.nonzero.cc.log2cpm <- CorShrinkVector(cor.cc.log2cpm,
                    nsamp_vec = nsam.nonzero, report_model=TRUE)

sval.cc <- corshrink.nonzero.cc.log2cpm$model$result$svalue

par(mfrow=c(1,2))
hist(cor.cc.log2cpm, nclass=50,
     main = "Correlation with Theta")
with(corshrink.nonzero.cc.log2cpm$model$result, { 
     plot(betahat,PosteriorMean, pch = 16, col = "gray40", cex=.7,
          xlab = "Observed effect size", ylab = "Shrunked effect size");
     abline(0,1, col = "mediumblue");
     points(betahat,PosteriorMean, col = as.numeric(sval.cc<.01)+1, lwd=.8)
     title(paste("sig. at svalue < .01", sum(sval.cc<.01), "genes"))
     })
```

Check cell cycle enrichment with s-value < .01.

```{r}
genes.sig.cc <- (rownames(counts.log2cpm.normed)[sval.cc < .01])[order(cor.cc.log2cpm[which(sval.cc < .01)], decreasing = F)]
sig.cycle.cc <- mean(genes.sig.cc %in% macosko$ensembl)
nonsig.cycle.cc <- mean(setdiff(rownames(counts.log2cpm.normed), genes.sig.cc) %in% macosko$ensembl)
sig.cycle.cc/nonsig.cycle.cc     

rho.sig.cc <- (cor.cc.log2cpm[sval.cc < .01])[order(cor.cc.log2cpm[which(sval.cc < .01)], decreasing = F)]

genes.sig.cycle.cc <- genes.sig.cc[genes.sig.cc %in% macosko$ensembl]
rho.sig.cycle.cc <- rho.sig.cc[genes.sig.cc %in% macosko$ensembl]
gene.names.sig.cycle.cc <- sapply(1:length(genes.sig.cycle.cc), function(g) {
  nm <- with(macosko, hgnc[ensembl == genes.sig.cycle.cc[g]]) 
  return(nm[1])})
```

Check cell cycle enrichment in top 200 genes.

```{r}
genes.sig.cc <- (rownames(counts.log2cpm.normed)[order(sval.cc) < 201])[order(cor.cc.log2cpm[which(order(sval.cc) < 201)], decreasing = F)]
sig.cycle.cc <- mean(genes.sig.cc %in% macosko$ensembl)
nonsig.cycle.cc <- mean(setdiff(rownames(counts.log2cpm.normed), genes.sig.cc) %in% macosko$ensembl)
sig.cycle.cc/nonsig.cycle.cc     

rho.sig.cc <- (cor.cc.log2cpm[order(sval.cc) < 201])[order(cor.cc.log2cpm[which(order(sval.cc) < 201)], decreasing = F)]

genes.sig.cycle.cc <- genes.sig.cc[genes.sig.cc %in% macosko$ensembl]
rho.sig.cycle.cc <- rho.sig.cc[genes.sig.cc %in% macosko$ensembl]
gene.names.sig.cycle.cc <- sapply(1:length(genes.sig.cycle.cc), function(g) {
  nm <- with(macosko, hgnc[ensembl == genes.sig.cycle.cc[g]]) 
  return(nm[1])})
```

Visualize expression of significant genes.

```{r,fig.width=6, fig.height=11}
par(mfrow=c(4,2))
for(i in 1:8) {
  plot(counts.log2cpm.normed[which(rownames(counts.log2cpm) == genes.sig.cycle.cc[i]),
                             order(as.numeric(Theta.fucci))],
     ylab = "Angle (0,2pi)", xlab = "Cells ordered by Theta",
     pch = 16, cex = .7, ylim = c(0,2*pi))
  title(paste(gene.names.sig.cycle.cc[i], "; corr = ", round(rho.sig.cycle.cc[i],digits=2)))
  points(as.numeric(Theta.fucci)[order(Theta.fucci)], col = "gray40", cex=.5, pch=16)
}
title("Expression pattern by Theta", outer=T, line=-1)

# sin versus sin
par(mfrow=c(4,2))
for(i in 1:8) {
  xx <- as.numeric(Theta.fucci)[order(Theta.fucci)]
  yy <- counts.log2cpm.normed[which(rownames(counts.log2cpm) == genes.sig.cycle.cc[i]),order(as.numeric(Theta.fucci))]
  xx <- sin(xx-mean(xx))
  yy <- sin(yy-mean(yy))
  plot(x=xx, y=yy,
     pch = 16, cex = .7, ylim=c(-1,1),
     xlab = "sin(Theta)", ylab="sin(gene angle)") 
  title(paste(gene.names.sig.cycle.cc[i], "; corr = ", round(rho.sig.cycle.cc[i],digits=2)))
}
title("sin(Theta) vs. sin(normalized gene expression)", outer=T, line=-1)

# write.table(cbind(genes.sig.cycle, rho.sig.cycle), file = "output_tmp/genes.sig.cycle.txt", quote=F, sep = "\t", col.names=F, row.names=F)
# 
# write.table(cbind(genes.sig, rho.sig), file = "output_tmp/genes.sig.txt", quote=F, sep = "\t", col.names=F, row.names=F)
```



## Circular-linear correlation between projected time and gene expression


```{r, fig.width=7, fig.height=4}
library(Rfast)
library(circular)

counts.log2cpm <- log10((10^6)*t(t(counts)/colSums(counts))+1)

cor.cl.log2cpm <- sapply(1:nrow(counts.log2cpm), function(g) {
    sqrt(circlin.cor(counts.log2cpm[g,],Theta.fucci)[1])
})
nsam.nonzero <- rowSums(counts.log2cpm > 0)

library(CorShrink)
corshrink.nonzero.cl.log2cpm <- CorShrinkVector(cor.cl.log2cpm, nsamp_vec = nsam.nonzero, report_model=TRUE)

sval.cl <- corshrink.nonzero.cl.log2cpm$model$result$svalue

par(mfrow=c(1,2))
hist(cor.cl.log2cpm, nclass=50,
     main = "Correlation with Theta")
with(corshrink.nonzero.cl.log2cpm$model$result, { 
     plot(betahat,PosteriorMean, pch = 16, col = "gray40", cex=.7,
          xlab = "Observed effect size", ylab = "Shrunked effect size");
     abline(0,1, col = "mediumblue");
     points(betahat,PosteriorMean, col = as.numeric(sval.cl<.01)+1, lwd=.8)
     title(paste("sig. at svalue < .01", sum(svalue<.01), "genes"))
     })
```

Check cell cycle enrichment in genes with s-value < .01.

```{r}
genes.sig.cl <- (rownames(counts.log2cpm)[sval.cl < .01])[order(cor.cl.log2cpm[which(sval.cl < .01)], decreasing = F)]
sig.cycle.cl <- mean(genes.sig.cl %in% macosko$ensembl)
nonsig.cycle.cl <- mean(setdiff(rownames(counts.log2cpm), genes.sig.cl) %in% macosko$ensembl)
sig.cycle.cl/nonsig.cycle.cl

rho.sig.cl <- (cor.cl.log2cpm[sval.cl < .01])[order(cor.cl.log2cpm[which(sval.cl < .01)], decreasing = F)]

genes.sig.cycle.cl <- genes.sig.cl[genes.sig.cl %in% macosko$ensembl]
rho.sig.cycle.cl <- rho.sig.cl[genes.sig.cl %in% macosko$ensembl]
gene.names.sig.cycle.cl <- sapply(1:length(genes.sig.cycle.cl), function(g) {
  nm <- with(macosko, hgnc[ensembl == genes.sig.cycle.cl[g]]) 
  return(nm[1])})


## these 8 are also significant in circular-circular correlation
sval.cc[rownames(counts.log2cpm.normed) %in% genes.sig.cycle.cl[1:8]]
```

Check cell cycle enrichment in genes with s-value in top 200.


```{r}
genes.sig.cl <- (rownames(counts.log2cpm)[order(sval.cl) < 201])[order(cor.cl.log2cpm[which(order(sval.cl) < 201)], decreasing = F)]
sig.cycle.cl <- mean(genes.sig.cl %in% macosko$ensembl)
nonsig.cycle.cl <- mean(setdiff(rownames(counts.log2cpm), genes.sig.cl) %in% macosko$ensembl)
sig.cycle.cl/nonsig.cycle.cl

rho.sig.cl <- (cor.cl.log2cpm[order(sval.cl) < 201])[order(cor.cl.log2cpm[which(order(sval.cl) < 201)], decreasing = F)]

genes.sig.cycle.cl <- genes.sig.cl[genes.sig.cl %in% macosko$ensembl]
rho.sig.cycle.cl <- rho.sig.cl[genes.sig.cl %in% macosko$ensembl]
gene.names.sig.cycle.cl <- sapply(1:length(genes.sig.cycle.cl), function(g) {
  nm <- with(macosko, hgnc[ensembl == genes.sig.cycle.cl[g]]) 
  return(nm[1])})


## these 8 are also significant in circular-circular correlation
sval.cc[rownames(counts.log2cpm.normed) %in% genes.sig.cycle.cl[1:8]]
```

Visualize expression of the significant genes.

```{r cir-liear-plots, fig.width=6, fig.height=11}
par(mfrow=c(4,2))
for(i in 1:min(length(genes.sig.cycle.cl),8)) {
  plot(x=as.numeric(Theta.fucci),
       y=counts.log2cpm.normed[which(rownames(counts.log2cpm) == genes.sig.cycle.cl[i]),
                             order(as.numeric(Theta.fucci))],
     ylab = "Angle (0,2pi)", xlab = "Sin(Theta)",
     pch = 16, cex = .7, ylim = c(1,5))
  title(paste(gene.names.sig.cycle.cl[i], "; corr = ", round(rho.sig.cycle.cl[i],digits=2)))
}
title("sinusoidal pattern by Theta?", outer=T, line=-1)

# sin versus data
par(mfrow=c(4,2))
for(i in 1:min(length(genes.sig.cycle.cl),8)) {
  xx <- as.numeric(Theta.fucci)[order(Theta.fucci)]
  yy <- counts.log2cpm.normed[which(rownames(counts.log2cpm) == genes.sig.cycle.cl[i]),order(as.numeric(Theta.fucci))]
  xx <- sin(xx-mean(xx))
  yy <- yy-mean(yy)
  plot(x=xx, y=yy,
     pch = 16, cex = .7, ylim=c(-1,1),
     xlab = "sin(Theta)", ylab="sin(gene angle)") 
  title(paste(gene.names.sig.cycle.cl[i], "; corr = ", round(rho.sig.cycle.cl[i],digits=2)))
}
title("Sin(Theta) vs. data", outer=TRUE, line=-1)

# cosine versus data
par(mfrow=c(4,2))
for(i in 1:min(length(genes.sig.cycle.cl),8)) {
  xx <- as.numeric(Theta.fucci)[order(Theta.fucci)]
  yy <- counts.log2cpm.normed[which(rownames(counts.log2cpm) == genes.sig.cycle.cl[i]),order(as.numeric(Theta.fucci))]
  xx <- cos(xx-mean(xx))
  yy <- yy-mean(yy)
  plot(x=xx, y=yy,
     pch = 16, cex = .7, ylim=c(-1,1),
     xlab = "cos(Theta)", ylab="sin(gene angle)") 
  title(paste(gene.names.sig.cycle.cl[i], "; corr = ", round(rho.sig.cycle.cl[i],digits=2)))
}
title("Cos(Theta) vs. data", outer=TRUE, line=-1)
```





---

## Comparae results between circ-linear and cir-circ correlation


1. There are almost no overlap in the top 200 genes between the two methods. Going up to top 500 genes, we see 19 of the 481 genes overlap between the two methods. In other words, a variable that is significantly corrleated with sin(theta) doesn't necessary have a high correlation in sin(variable value) with sin(theta). This also implies that all previous examples that I plotted for each correlation case is only significant for that particular category.

2. Results above on gene enrichment showed that when considering top 200 genes, cir-cir results are enriched two-fold with cell cycle genes (in Macosko data), while cir-linear are enriched with .5 with cell cycle genes, i.e., the top 200 have fewer cell cycle genes than the other genes. When consider s-values, cir-cir has about 11 fold of enrichment and cir-linear has about 2 fold enrichment...


```{r}
# so many more when considering circular-linear correlation
# probably because it's picking up sinuisoidal patterns, hence osciallating genes
# find some examples...
table(order(sval.cc)<201, order(sval.cl)<201)

table(order(sval.cc)<501, order(sval.cl)<501)

plot(rank(sval.cc), rank(sval.cl))
```




---


```{r, echo = FALSE}
sessionInfo()
```



