---
title: "Quantile-normalize intensity measurements"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Introduction

Previously when investigating measurement variation (GFP/RFP/FAPI), we learned that there's significant variation between batches in the distributions of background-corrected pixel intensites. See [here](../docs/images-qc-labels.html).

In this document, I apply quantile normalization to intensity measurements. Bad news, it did not work well. While quantile normalization helps to remove variation between batches in intensity measurements, but by doing so, we increases the variation between individuals. The cells do not form any clusters in the Green versus Red plot. In hindsight, this seems to make sense. By normalizing samples within plate, we artifically inflated differences between individuals...

Next to try linear mixed model...

---

## Load packages and data

```{r, message=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(cowplot)
library(wesanderson)
library(RColorBrewer)
library(Biobase)
library(Lmoments)
```

Name all plates.

```{r}
plates <- c("18511_18855","18855_19101","18855_19160","18870_18511",
            "18870_18855","18870_19101","18870_19160","19098_18511",
            "19098_18870","19098_19160","19101_18511","19101_19098",
            "19160_18870","19101_19160","19160_18511", "18855_19098")
```

Combine intensity stats from different plates.

```{r, eval = TRUE, message=FALSE, warning=FALSE}
# make the negative ones be the samllest one within its own plate
ints <- do.call(rbind, lapply(1:length(plates), function(index) {
  tmp <- readRDS(paste0("/project2/gilad/fucci-seq/intensities_stats/",plates[index],".stats.rds"))
  tmp <- data.frame(plate=plates[index], 
                    well=as.character(droplevels(tmp$wellID)),
         rfp.sum.zoom.mean.log10=log10(tmp$rfp.sum.zoom.mean),
         gfp.sum.zoom.mean.log10=log10(tmp$gfp.sum.zoom.mean),
         dapi.sum.zoom.mean.log10=log10(tmp$dapi.sum.zoom.mean))

  tmp$rfp.sum.zoom.mean.log10[which(tmp$rfp.sum.zoom.mean.log10 == "NaN")] <- min(tmp$rfp.sum.zoom.mean.log10, na.rm=TRUE)
  tmp$gfp.sum.zoom.mean.log10[which(tmp$gfp.sum.zoom.mean.log10 == "NaN")] <- min(tmp$gfp.sum.zoom.mean.log10, na.rm=TRUE)
  tmp$dapi.sum.zoom.mean.log10[which(tmp$dapi.sum.zoom.mean.log10 == "NaN")] <- min(tmp$dapi.sum.zoom.mean.log10, na.rm=TRUE)
  return(tmp)  
}) )
```

Load the expression set info.

```{r}
eset_fls <- list.files("../data/eset", full.names=TRUE)

anno <- do.call(rbind, lapply(1:length(eset_fls), function(index) {
  eset_index <- readRDS(eset_fls[[index]])
  pdata_index <- pData(eset_index)
  return(pdata_index)
}))

# make unique id in both
ints$unique <- paste0(ints$plate,"_",as.numeric(ints$well))
anno$unique <- paste0(anno$image_individual,"_",anno$image_label)

subset_index1 <- which(anno$unique %in% ints$unique)
anno_subset <- anno[subset_index1,]

subset_index2 <- match(anno_subset$unique, ints$unique)
ints_tmp <- ints[subset_index2,]
all.equal(ints_tmp$unique, anno_subset$unique)

ints_tmp$chip_id <- anno_subset$chip_id
```

---


## Densit plots

First, look at the distribution of all batches combined versus each batch. 

```{r, echo = FALSE}
plot_grid(
  ggplot(ints_tmp, aes(x=gfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=gfp.sum.zoom.mean.log10, col = as.factor(plate))) +
    labs(title = "Green (log10 pixel sum) by plate",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Green channel log10 pixel sum", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_tmp, aes(x=rfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=rfp.sum.zoom.mean.log10, col = as.factor(plate))) +
    labs(title = "Red (log10 pixel sum) by plate",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Red channel log10 pixel sum", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_tmp, aes(x=dapi.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=dapi.sum.zoom.mean.log10, col = as.factor(plate))) +
    labs(title = "DAPI (log10 pixel sum) by plate",
         subtitle = "Black line indicates distribution aggregating plates",
         x="DAPI channel log10 pixel sum", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

---

## Normalization

Code for one single sample

```{r}
my_quantnorm <- function(reference, sample, span=.01) {
  # quantiles for intensities all samples across plates
  quants_reference <- quantile(reference, probs=seq(0,1,span))
  # intensities for a given plate
  # quantiles for intensities at each plate
  quants_sample <- quantile(sample, probs=seq(0,1,span))
  # empty vector for normalized values
  sample_normed <- vector("numeric", length=length(sample))
  
  for (index in 1:length(sample)) {
    # compare the sample intensity values with the refernece sample intensity values 
    # at each quantile; get the closes reference quantlie in terms of measurement values
    sample_order <- names(which.min(abs(sample[index]-quants_sample)))
    # get the reference intensity value of the closet quantile
    ref_order_value <- reference[which(names(quants_reference)==sample_order)] 
    # assign the reference intensity value to the sample
    sample_normed[index] <- ref_order_value 
  }
  return(sample_normed)
}
```

**Choose smoothing parameter for RFP**

```{r, echo = FALSE}
ints_tmp$plate <- as.character(ints_tmp$plate)
ints_rfp <- data.frame(plate=ints_tmp$plate,
                       rfp.sum.zoom.mean.log10 = ints_tmp$rfp.sum.zoom.mean.log10)
ints_rfp$span1 <- NULL
ints_rfp$span2 <- NULL
ints_rfp$span3 <- NULL
ints_rfp$span4 <- NULL

for (index in 1:length(unique(ints_rfp$plate))) {
  plate <- unique(ints_rfp$plate)[index]
#  print(plate)
  tmp <- my_quantnorm(reference = ints_rfp$rfp.sum.zoom.mean.log10,
                      sample = ints_rfp$rfp.sum.zoom.mean.log10[ints_rfp$plate == plate],
                      span=.02)
  ints_rfp$span1[ints_rfp$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_rfp$rfp.sum.zoom.mean.log10,
                      sample = ints_rfp$rfp.sum.zoom.mean.log10[ints_rfp$plate == plate],
                      span=.01)
  ints_rfp$span2[ints_rfp$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = ints_rfp$rfp.sum.zoom.mean.log10,
                      sample = ints_rfp$rfp.sum.zoom.mean.log10[ints_rfp$plate == plate],
                      span=.005)
  ints_rfp$span3[ints_rfp$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_rfp$rfp.sum.zoom.mean.log10,
                    sample = ints_rfp$rfp.sum.zoom.mean.log10[ints_rfp$plate == plate],
                    span=.001)
  ints_rfp$span4[ints_rfp$plate == plate] <- tmp
}

plot_grid(
  ggplot(ints_rfp, aes(x=rfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span1, col = as.factor(plate))) +
    labs(title = "Span=.02",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_rfp, aes(x=rfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span2, col = as.factor(plate))) +
    labs(title = "Span=.01",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_rfp, aes(x=rfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span3, col = as.factor(plate))) +
    labs(title = "Span=.005",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_rfp, aes(x=rfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span4, col = as.factor(plate))) +
    labs(title = "Span=.001",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```


**Choose smoothing parameter for GFP**

```{r, echo = FALSE}
ints_tmp$plate <- as.character(ints_tmp$plate)
ints_gfp <- data.frame(plate=ints_tmp$plate,
                       gfp.sum.zoom.mean.log10 = ints_tmp$gfp.sum.zoom.mean.log10)
ints_gfp$span1 <- NULL
ints_gfp$span2 <- NULL
ints_gfp$span3 <- NULL
ints_gfp$span4 <- NULL

for (index in 1:length(unique(ints_gfp$plate))) {
  plate <- unique(ints_gfp$plate)[index]
#  print(plate)
  tmp <- my_quantnorm(reference = ints_gfp$gfp.sum.zoom.mean.log10,
                      sample = ints_gfp$gfp.sum.zoom.mean.log10[ints_gfp$plate == plate],
                      span=.02)
  ints_gfp$span1[ints_gfp$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_gfp$gfp.sum.zoom.mean.log10,
                      sample = ints_gfp$gfp.sum.zoom.mean.log10[ints_gfp$plate == plate],
                      span=.01)
  ints_gfp$span2[ints_gfp$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = ints_gfp$gfp.sum.zoom.mean.log10,
                      sample = ints_gfp$gfp.sum.zoom.mean.log10[ints_gfp$plate == plate],
                      span=.005)
  ints_gfp$span3[ints_gfp$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_gfp$gfp.sum.zoom.mean.log10,
                    sample = ints_gfp$gfp.sum.zoom.mean.log10[ints_gfp$plate == plate],
                    span=.001)
  ints_gfp$span4[ints_gfp$plate == plate] <- tmp
}

plot_grid(
  ggplot(ints_gfp, aes(x=gfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span1, col = as.factor(plate))) +
    labs(title = "Span=.02",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_gfp, aes(x=gfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span2, col = as.factor(plate))) +
    labs(title = "Span=.01",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_gfp, aes(x=gfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span3, col = as.factor(plate))) +
    labs(title = "Span=.005",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_gfp, aes(x=gfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span4, col = as.factor(plate))) +
    labs(title = "Span=.001",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

**Choose smoothing parameter for DAPI**

```{r, echo = FALSE}
ints_tmp$plate <- as.character(ints_tmp$plate)
ints_dapi <- data.frame(plate=ints_tmp$plate,
                       dapi.sum.zoom.mean.log10 = ints_tmp$dapi.sum.zoom.mean.log10)
ints_dapi$span1 <- NULL
ints_dapi$span2 <- NULL
ints_dapi$span3 <- NULL
ints_dapi$span4 <- NULL

for (index in 1:length(unique(ints_dapi$plate))) {
  plate <- unique(ints_dapi$plate)[index]
#  print(plate)
  tmp <- my_quantnorm(reference = ints_dapi$dapi.sum.zoom.mean.log10,
                      sample = ints_dapi$dapi.sum.zoom.mean.log10[ints_dapi$plate == plate],
                      span=.02)
  ints_dapi$span1[ints_dapi$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_dapi$dapi.sum.zoom.mean.log10,
                      sample = ints_dapi$dapi.sum.zoom.mean.log10[ints_dapi$plate == plate],
                      span=.01)
  ints_dapi$span2[ints_dapi$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = ints_dapi$dapi.sum.zoom.mean.log10,
                      sample = ints_dapi$dapi.sum.zoom.mean.log10[ints_dapi$plate == plate],
                      span=.005)
  ints_dapi$span3[ints_dapi$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_dapi$dapi.sum.zoom.mean.log10,
                    sample = ints_dapi$dapi.sum.zoom.mean.log10[ints_dapi$plate == plate],
                    span=.001)
  ints_dapi$span4[ints_dapi$plate == plate] <- tmp
}

plot_grid(
  ggplot(ints_dapi, aes(x=dapi.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span1, col = as.factor(plate))) +
    labs(title = "Span=.02",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_dapi, aes(x=dapi.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span2, col = as.factor(plate))) +
    labs(title = "Span=.01",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_dapi, aes(x=dapi.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span3, col = as.factor(plate))) +
    labs(title = "Span=.005",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_dapi, aes(x=dapi.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=span4, col = as.factor(plate))) +
    labs(title = "Span=.001",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```


**Normalize intensity for each channel**

DAPI choose .001; Green choose .01; Red choose .01

```{r, echo = FALSE}
ints_tmp$plate <- as.character(ints_tmp$plate)
ints_normed <- data.frame(plate=ints_tmp$plate,
                          chip_id=ints_tmp$chip_id,
                       dapi.sum.zoom.mean.log10 = ints_tmp$dapi.sum.zoom.mean.log10,
                       rfp.sum.zoom.mean.log10 = ints_tmp$rfp.sum.zoom.mean.log10,
                       gfp.sum.zoom.mean.log10 = ints_tmp$gfp.sum.zoom.mean.log10)
ints_normed$rfp.normed <- NULL
ints_normed$gfp.normed <- NULL
ints_normed$dapi.normed <- NULL

for (index in 1:length(unique(ints_normed$plate))) {
  plate <- unique(ints_normed$plate)[index]

  tmp <- my_quantnorm(reference = ints_normed$dapi.sum.zoom.mean.log10,
                      sample = ints_normed$dapi.sum.zoom.mean.log10[ints_normed$plate == plate],
                      span=.001)
  ints_normed$dapi.normed[ints_dapi$plate == plate] <- tmp

  tmp <- my_quantnorm(reference = ints_normed$rfp.sum.zoom.mean.log10,
                      sample = ints_normed$rfp.sum.zoom.mean.log10[ints_normed$plate == plate],
                      span=.01)
  ints_normed$rfp.normed[ints_normed$plate == plate] <- tmp
  
  tmp <- my_quantnorm(reference = ints_normed$gfp.sum.zoom.mean.log10,
                    sample = ints_normed$gfp.sum.zoom.mean.log10[ints_normed$plate == plate],
                    span=.01)
  ints_normed$gfp.normed[ints_normed$plate == plate] <- tmp
}

plot_grid(
  ggplot(ints_normed, aes(x=gfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=gfp.normed, col = as.factor(plate))) +
    labs(title = "GFP",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed GFP intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_normed, aes(x=rfp.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=rfp.normed, col = as.factor(plate))) +
    labs(title = "RFP",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed RFP intensity", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  ggplot(ints_normed, aes(x=dapi.sum.zoom.mean.log10), col="black") + 
    geom_density(alpha = .5, cex = .7) + 
    geom_density(aes(x=dapi.normed, col = as.factor(plate))) +
    labs(title = "DAPI",
         subtitle = "Black line indicates distribution aggregating plates",
         x="Normed DAPI intensities", y = "Density") + 
    theme_gray() + theme(legend.position="none"),
  nrow = 2)
```

After normalization: Green versus Red by individual

```{r, fig.width=12, fig.height=5, echo = FALSE}
ints_normed <- ints_normed %>% 
  mutate(dapi_4quant=ntile(dapi.normed,4),
         dapi_3quant=ntile(dapi.normed,3))

plot_grid(
  ggplot(ints_normed, aes(x=rfp.normed,y=gfp.normed, col=as.factor(dapi_4quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(title = "Normalized intensities labeled by DAPI 4-quantiles",
         x="RFP", y = "GFP") + theme_gray(), 
  ggplot(ints_normed, aes(x=rfp.normed,y=gfp.normed, col=as.factor(dapi_3quant))) +
  geom_point(alpha = .5, cex = .7) + 
  facet_wrap(~as.factor(chip_id), ncol=3) + 
  labs(title = "Normalized intensities labeled by DAPI 3-quantiles",
       x="RFP", y = "GFP") + theme_gray(), ncol=2)
```


Before normalization: Green versus Red by individual

```{r, fig.width=12, fig.height=5, echo = FALSE}
ints_normed <- ints_normed %>% 
  mutate(dapi_4quant=ntile(dapi.sum.zoom.mean.log10,4),
         dapi_3quant=ntile(dapi.sum.zoom.mean.log10,3))

plot_grid(
  ggplot(ints_normed, aes(x=rfp.sum.zoom.mean.log10,
                          y=gfp.sum.zoom.mean.log10, col=as.factor(dapi_4quant))) +
    geom_point(alpha = .5, cex = .7) + 
    facet_wrap(~as.factor(chip_id), ncol=3) + 
    labs(title = "Raw intensities labeled by DAPI 4-quantiles",
         x="RFP", y = "GFP") + theme_gray(), 
  ggplot(ints_normed, aes(x=rfp.sum.zoom.mean.log10,
                          y=gfp.sum.zoom.mean.log10, col=as.factor(dapi_3quant))) +
  geom_point(alpha = .5, cex = .7) + 
  facet_wrap(~as.factor(chip_id), ncol=3) + 
  labs(title = "Raw intensities labeled by DAPI 3-quantiles",
       x="RFP", y = "GFP") + theme_gray(), ncol=2)
```

---

## Session information

```{r, echo=FALSE}
sessionInfo()
```

