---
title: "parametric cellcycleR for image intensities: convergence assessment"
author: "Joyce Hsiao"
output: workflowr::wflow_html
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```


---

## Introduction/Summary

In this document, we assess the convergence of `cellcycleR` across all samples and all batches on fitting intensity data - three intensity vectors for ~1,200 single cell samples. 


**Results**:

20 runs of `sin_cell_ordering_class` on 20 different random seeds. Each 500 iterations. We assess the range of log-likelihood, and model estimates $\hat{\alpha}_g$, $\hat{\phi}_g$, and $\hat{\sigma}_g$.


1. Large deviation in the estimates of $\phi_g$: because $\phi_g$ is not identifiable.

2. Not much deviation between the 20 runs in the estimates for $\alpha_g$ amplitude or $\sigma_g$ or log-likelihood. 

3. Re. speed of convergence, across random seeds, it takes about less than 50 iterations to reach convergence. Note that the observation vector of each cell is a three-dimensional vector (green, red, and dapi). The number of iterations will go up as we increase the dimension of the observations (e.g., number of genes).

---

## Data

Combined intensity data (see [combine-intensity-data.R](/project2/gilad/joycehsiao/fucci-seq/code/combine-intensity-data.R)).

```{r, message=FALSE, warning=FALSE, echo = FALSE}
library(ggplot2)
library(cellcycleR)
library(data.table)
library(dplyr)
library(cowplot)
library(wesanderson)
library(RColorBrewer)
library(Biobase)
library(plotrix)

ints <- readRDS(file="/project2/gilad/joycehsiao/fucci-seq/data/intensity.rds")
ints$rfp.z <- scale(ints$rfp.median.log10sum, center=TRUE, scale=TRUE)
ints$gfp.z <- scale(ints$gfp.median.log10sum, center=TRUE, scale=TRUE)
ints$dapi.z <- scale(ints$dapi.median.log10sum, center=TRUE, scale=TRUE)

ints_tmp <- data.frame(rfp.z=ints$rfp.z,
                       gfp.z=ints$gfp.z,
                       dapi.z=ints$dapi.z)
```

---

## Model overview

For $S$ cells, let the vector of true cell time be $t_S$. We use an iterative scheme to estimate cell time parameters $t_S$, and curve parameters $\alpha_g$, $\phi_g$, and $\sigma_g$. First, we consider $T$ time classes which is a set of uniformly spaced time points on $(0, 2\pi)$. In the first iteration, we assign each cell $s$ to a time class $t_s^{(0)}$ and estimate the curve parameters $\alpha_g$, $\phi_g$, and $\sigma_g$. In the second iteratio, we estimate $t_s^{(1)}$ based on the model estimates from the first iteration ($\hat{\alpha}_g$, $\hat{\phi}_g$, and $\hat{\sigma}_g$). We continue this iterative scheme until convergence. For any $n$ iterations, starting from 0, we fit the following model for gene $g$ and and cell $s$.


$$
Y_{sg} = \alpha_g sin(t_s^{(n)} + \phi_g) + \epsilon_{sg}
$$
where $\epsilon_{sg} \sim N(0, \sigma^2_g)$, and frequency is 1.


---

## Analysis

20 runs of `sin_cell_ordering_class` on 20 different random seeds. Each 500 iterations.

We assess the range of log-likelihood, and model estimates $\hat{\alpha}_g$, $\hat{\phi}_g$, and $\hat{\sigma}_g$.


```{r, echo = FALSE, eval = FALSE}
# cellcycleR analysis
iter <- seq(1:20)
seeds <- c(2,3,5,7,11,13,17,19,23,29,31,37,41,43,47,53,59,61,67,71)

for (index in iter) {
  set.seed(seeds[index])
  out <- sin_cell_ordering_class(as.matrix(ints_tmp), 
                                 celltime_levels=300, 
                                 num_shuffle=1, maxiter=500, tol=1e-6,
                                 fix.phase = FALSE, phase_in=NULL,
                                 n_cores=5)
  saveRDS(out, 
          file = paste0("/project2/gilad/joycehsiao/fucci-seq/output/images-cellcycleR-convergence.Rmd/out","_", sprintf("%02d", index),".rds") )
}
```


```{r, echo = FALSE, eval = TRUE}
fname <- Sys.glob("../output/images-cellcycleR-convergence.Rmd/*.rds")
objs <- lapply(fname, readRDS)
loglik <- sapply(objs, "[[", "loglik")

amp <- do.call(rbind, lapply(objs, "[[", "amp"))
phi <- do.call(rbind, lapply(objs, "[[", "phi"))
sigma <- do.call(rbind, lapply(objs, "[[", "sigma"))

colnames(amp) <- c("Red", "Green", "DAPI")
colnames(phi) <- c("Red", "Green", "DAPI")
colnames(sigma) <- c("Red", "Green", "DAPI")
```

```{r}
summary(loglik)
summary(amp)
summary(phi)
summary(sigma)
```

log-liklihood distribution

```{r}
plot(loglik)
```


```{r, fig.width=10, fig.height=12}
par(mfrow=c(5,4), mar=c(1,1,2.5,1), oma=rep(1,4))
for (i in 1:length(objs)) {
plot(objs[[i]]$loglik_iter[-1],
     main = paste("round no.", i))
}
title("log-likelihood", outer=TRUE)
```

speed of convergence

```{r}
maxiter <- vector("numeric", length(objs))
for (i in 1:length(maxiter)) {
  maxiter[i] <- which.max(objs[[i]]$loglik_iter[-1]) 
}

boxplot(maxiter, outpch = NA, main = "no. iterations reached for convergence") 
stripchart(maxiter,  
            vertical = TRUE, method = "jitter", 
            pch = 21, col = "maroon", bg = "bisque", 
            add = TRUE) 
```
---


```{r, echo = FALSE}
sessionInfo()
```


