---
title: "Compile training results"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Two training scenarios

1. Cell times derived from fucci: mixed individua predict mixed individual, mixed individual predict one individual

2. Cell times derived from fucci + dapi: mixed individua predict mixed individual, mixed individual predict one individual

---

## Codes

There are two main steps in the analysis.

1. `code/working/job_run_methods.cyclical.R`: runs for one fold at a time the nonparametic regression, estimates the cyclical functions for every single gene and saves the PVE. This generates `data/results/data_training_cyclical_genes.fold.K.rds`

2. `code/working/job_run_methods.train.R`: runs for one fold at the time, takes the output from the previous step, and computes the prediction error of the top X cyclical genes. This generates `data/results/results_train.fold.K.topX.rds`.

For the training results on random error, we compute the prediction error of the top X cyclical genes on permuted cell times. This generates `data/results/results_train_permute_oncyclical.K.topX.rds`.

For the different individual prediction scenario, we perform the training five times. Each time uses data from four individuals to estimate per-gene cyclical gene expression fucntions and estimates the prediction error of top X genes in a different individual. 

Note that these two sets of training results are evaluated differently. Namely, in the case where we learn genes that are predictive of cell times for the same individuals, we evaluate the selected genes in a group of mixed individuals. And, in the case where we learn genes that are predictive of cell times for a different individual, we evaluate the selected genes in yet another individual. Hence, we get six sets of evaluation results for the different individual scenario and one set of evaluation results for the same individual scenario. 


---

## Make top cyclical genes lists

* Choose across the 5 folds (training sets), the genes that appear among the top X cyclical genes in at least 4 training sets.

* Overlap between random prediction and individual prediction: a gene has to appear in the random list and in 4 or more individuals in the individual list.


```{r, eval=F}
# cell times based on fucci: mixed individuals ---------------------------
data_training_cyclical_genes.fold.1 <- readRDS("../data/results/data_training_cyclical_genes.fold.1.rds")
data_training_cyclical_genes.fold.2 <- readRDS("../data/results/data_training_cyclical_genes.fold.2.rds")
data_training_cyclical_genes.fold.3 <- readRDS("../data/results/data_training_cyclical_genes.fold.3.rds")
data_training_cyclical_genes.fold.4 <- readRDS("../data/results/data_training_cyclical_genes.fold.4.rds")
data_training_cyclical_genes.fold.5 <- readRDS("../data/results/data_training_cyclical_genes.fold.5.rds")
data_cyclical_list <- list(data_training_cyclical_genes.fold.1,
                           data_training_cyclical_genes.fold.2,
                           data_training_cyclical_genes.fold.3,
                           data_training_cyclical_genes.fold.4,
                           data_training_cyclical_genes.fold.5)

ngenes <- c(5, seq(10, nrow(data_cyclical_list[[1]]), 10))
genes_list <- lapply(1:length(ngenes), function(i) {
  ngene <- ngenes[i]
  tmp <- do.call(cbind, lapply(1:5, function(fold) {
    top_list <- rownames(data_cyclical_list[[fold]])[order(data_cyclical_list[[fold]]$pve,
                                                        decreasing = T)[1:ngene]]
    rownames(data_cyclical_list[[fold]]) %in% top_list
  }) )
  rownames(tmp) <- rownames(data_cyclical_list[[fold]])
  return(rownames(tmp)[rowSums(tmp)>=4])
})
names(genes_list) <- ngenes

saveRDS(genes_list, 
        file = "../output/method-train-summary-output.Rmd/double_topgenes_mixed.rds")
```


Below runs `code/working/job_run_methods.ind.R`.

```{r, eval=F}
# cell times based on fucci: one individual ---------------------------
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
ngenes <- c(5, seq(10, 11040, 10))
for (j in 1:length(inds)) {
  ind <- inds[j]
  gene_names <- rownames(readRDS(paste0("../data/results/ind_",ind,"_data_training_cyclical_genes.fold.",
                   1,".rds")))
  genes_list <- lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    tmp <- do.call(cbind, lapply(1:5, function(fold) {
      fl_name <- paste0("../data/results/ind_",ind,"_data_training_cyclical_genes.fold.",
                   fold,".rds")
      df <- readRDS(fl_name)
      top_list <- rownames(df)[order(df$pve,decreasing = T)[1:ngene]]
      rownames(df) %in% top_list
    }) )
    rownames(tmp) <- gene_names
    return(rownames(tmp)[rowSums(tmp)>=4])
  })
  names(genes_list) <- ngenes
  saveRDS(genes_list, 
          file = paste0("../output/method-train-summary-output.Rmd/double_topgenes_",ind,".rds"))
}
```


Hold off results on including DAPI as cell time measure.

```{r, eval=F}
# # cell times based on fucci an dapi: mixed individuals ---------------------------
# data_training_cyclical_genes.fold.1 <- readRDS("../data/results/triple_alge_data_training_cyclical_genes.fold.1.rds")
# data_training_cyclical_genes.fold.2 <- readRDS("../data/results/triple_alge_data_training_cyclical_genes.fold.2.rds")
# data_training_cyclical_genes.fold.3 <- readRDS("../data/results/triple_alge_data_training_cyclical_genes.fold.3.rds")
# data_training_cyclical_genes.fold.4 <- readRDS("../data/results/triple_alge_data_training_cyclical_genes.fold.4.rds")
# data_training_cyclical_genes.fold.5 <- readRDS("../data/results/triple_alge_data_training_cyclical_genes.fold.5.rds")
# data_cyclical_list <- list(data_training_cyclical_genes.fold.1,
#                            data_training_cyclical_genes.fold.2,
#                            data_training_cyclical_genes.fold.3,
#                            data_training_cyclical_genes.fold.4,
#                            data_training_cyclical_genes.fold.5)
# 
# ngenes <- c(5, seq(10, nrow(data_cyclical_list[[1]]), 10))
# genes_list <- lapply(1:length(ngenes), function(i) {
#   ngene <- ngenes[i]
#   tmp <- do.call(cbind, lapply(1:5, function(fold) {
#     top_list <- rownames(data_cyclical_list[[fold]])[order(data_cyclical_list[[fold]]$pve,
#                                                         decreasing = T)[1:ngene]]
#     rownames(data_cyclical_list[[fold]]) %in% top_list
#   }) )
#   rownames(tmp) <- rownames(data_cyclical_list[[fold]])
#   return(rownames(tmp)[rowSums(tmp)>=4])
# })
# names(genes_list) <- ngenes
# 
# saveRDS(genes_list, 
#         file = "../output/method-train-summary-output.Rmd/triple_topgenes_mixed.rds")
# 
# 
# 
# # cell times based on fucci and dapi: one individual ---------------------------
# inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
# ngenes <- c(5, seq(10, 11040, 10))
# for (j in 1:length(inds)) {
#   ind <- inds[j]
#   gene_names <- rownames(readRDS(paste0("../data/results/triple_alge_ind_",ind,"_data_training_cyclical_genes.fold.",
#                    1,".rds")))
#   genes_list <- lapply(1:length(ngenes), function(i) {
#     ngene <- ngenes[i]
#     tmp <- do.call(cbind, lapply(1:5, function(fold) {
#       fl_name <- paste0("../data/results/triple_alge_ind_",ind,"_data_training_cyclical_genes.fold.",
#                    fold,".rds")
#       df <- readRDS(fl_name)
#       top_list <- rownames(df)[order(df$pve,decreasing = T)[1:ngene]]
#       rownames(df) %in% top_list
#     }) )
#     rownames(tmp) <- gene_names
#     return(rownames(tmp)[rowSums(tmp)>=4])
#   })
#   names(genes_list) <- ngenes
#   saveRDS(genes_list, 
#           file = paste0("../output/method-train-summary-output.Rmd/triple_topgenes_",ind,".rds"))
# }
```


---

## Prediction error by top X cyclical genes

code for summarize results

```{r, eval=F}
diff_time_wrapper <- function(results_list) {
  
  methods_list <- sapply(names(results_list),
                         function(x) strsplit(x, split=".", fixed=TRUE)[[1]][2])
  
  diff_time_list <- do.call(rbind, lapply(1:length(results_list), function(i) {
    diff_time <- results_list[[i]]$diff_time
    diff_mean <- mean(diff_time/2/pi)

    return(data.frame(diff_mean=diff_mean,
#                      diff_se=diff_se,
                      methods=methods_list[i]))
  }) )
  return(diff_time_list)  
}


# cell times based on FUCCI, mixed individuals -------------------------------------------
ngenes <- c(5, seq(10, 1000, by=10))
train_top <- do.call(rbind, lapply(1:length(ngenes), function(i) {
  ngene <- ngenes[i]
  train_topX <- do.call(rbind, lapply(1:5, function(fold) {
    fl_name <- paste0("../data/results/results_train.fold.",fold,".top",ngene,".rds")
    df <- readRDS(fl_name)
    out <- diff_time_wrapper(df$fit.test)
    out$fold <- fold
    return(out)
  }) )
  train_topX$ngenes <- ngene
  #return(train_topX)  
  agg_mn <- aggregate(diff_mean ~ methods,
                    data=train_topX, FUN=mean)
  agg_sd <- aggregate(diff_mean ~ methods,
                    data=train_topX, FUN=sd)

  obj <- data.frame(methods=agg_mn$methods, 
                    diff_mean=agg_mn$diff_mean,
                    diff_se=agg_sd$diff_mean/sqrt(5))
  obj$ngenes <- ngene
  return(obj)
}) )
saveRDS(train_top, "../output/method-train-summary-output.Rmd/double_diff_time_mixed.rds")


# cell times based on FUCCI, mixed individuals, permuted labels ---------------------------
ngenes <- c(5, seq(10,1000, by=10))
train_top_permute <- do.call(rbind, lapply(1:length(ngenes), function(i) {
  ngene <- ngenes[i]
  train_topX <- do.call(rbind, lapply(1:5, function(fold) {
    fl_name <- paste0("../data/results/results_train_permute_oncyclical.fold.",
                      fold,".top",ngene,".rds")
    df <- readRDS(fl_name)
    out <- diff_time_wrapper(df$fit.test)
    out$fold <- fold
    return(out)
  }) )
  train_topX$ngenes <- ngene
  #return(train_topX)  
  agg_mn <- aggregate(diff_mean ~ methods,
                    data=train_topX, FUN=mean)
  agg_sd <- aggregate(diff_mean ~ methods,
                    data=train_topX, FUN=sd)

  obj <- data.frame(methods=agg_mn$methods, 
                    diff_mean=agg_mn$diff_mean,
                    diff_se=agg_sd$diff_mean/sqrt(5))
  obj$ngenes <- ngene
  return(obj)
}) )

saveRDS(train_top_permute, 
    file = "../output/method-train-summary-output.Rmd/double_diff_time_mixed_permute.rds")



# cell times based on fucci, predicting one individual ----------------------------------
# some problems with NA18870, fold 4, 810 gene
# some problem with NA18855, fold 1, 770 gene
ngenes <- c(5, seq(10,700, by=10))
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
train_top <- lapply(1:length(inds), function(j) {
  ind <- inds[j]
  out <- do.call(rbind, lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    train_topX <- do.call(rbind, lapply(1:5, function(fold) {
      print(ind)
      print(ngene)
      print(fold)
      fl_name <- paste0("../data/results/ind_",ind,"_results_train.fold.",fold,
                        ".top",ngene,".rds")
      df <- readRDS(fl_name)
      out <- diff_time_wrapper(df$fit.test)
      out$fold <- fold
      return(out)
    }) )
    train_topX$ngenes <- ngene
    #return(train_topX)  
    agg_mn <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=mean)
    agg_sd <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=sd)
  
    obj <- data.frame(methods=agg_mn$methods, 
                      diff_mean=agg_mn$diff_mean,
                      diff_se=agg_sd$diff_mean/sqrt(5))
    obj$ngenes <- ngene
    return(obj)
  }) )
  out$ind <- ind
  return(out)
})
names(train_top) <- inds

saveRDS(train_top, 
  file = "../output/method-train-summary-output.Rmd/double_diff_time_ind.rds")
```



---


```{r}
sessionInfo()
```

