---
title: "Evaluate the withheld sample on top X cyclical"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Run on supervised model

test cluster code

```{r, eval=F}
ngenes=101

dir <-"/project2/gilad/joycehsiao/fucci-seq"
source(file.path(dir,"code/working/run_methods.R"))

data_training <- readRDS(file=file.path(dir, "data/results/data_training.rds"))
data_withheld <-readRDS(file=file.path(dir, "data/results/data_withheld.rds"))

sig.genes <- readRDS(file=file.path(dir,
    "output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.476.rds"))

# make prediction parameters
Y_train_topX <- data_training$log2cpm.quant.nonvalid[
  rownames(data_training$log2cpm.quant.nonvalid) %in% rownames(sig.genes)[1:ngenes], ]

training_topX <- cycle.npreg.insample(Y = Y_train_topX,
                                        theta = data_training$theta.nonvalid,
                                        polyorder=2,
                                        ncores=15,
                                        method.trend="trendfilter")


seurat.genes <- readLines(
  con = file.path(dir,                                    "data/cellcycle-genes-previous-studies/seurat_cellcycle/regev_lab_cell_cycle_genes.txt"))
seurat.genes <- list(s.genes=seurat.genes[1:43],
                     g2m.genes=seurat.genes[44:97])

training_model <- training_topX
Y_test_normed <- data_withheld$log2cpm.quant.valid
  
cycle.genes <- rownames(training_model$Y)
Y_test_normed.cycle <- Y_test_normed[which(rownames(Y_test_normed) %in% cycle.genes),]
#Y_test.cycle <- Y_test[which(rownames(Y_test) %in% cycle.genes),]

### supervised methods
message("Begin supervised method...")
fit.supervised <- cycle.npreg.outsample(Y_test=Y_test_normed.cycle,
                                        sigma_est=training_model$sigma_est,
                                        funs_est=training_model$funs_est,
                                        method.grid = "uniform",
                                        method.trend="trendfilter",
                                        polyorder=2,
                                        ncores=ncores)
  
pred_time <- fit.supervised$cell_times_est
pred_time_shift <- with(fit.supervised, rotation(data_withheld$theta.valid,
                                                 cell_times_est)$y2shift)
#plot(pred_time, pred_time_shift)
source("../peco/R/utility.R")
pv <- get.pve(data_withheld$pdata.valid$gfp.median.log10sum.adjust[order(pred_time_shift)])


bb <- readRDS("../data/results/results_train_top101.rds")

bb1.pve <- get.pve(bb$fold.1$fit.supervised$dapi[order(bb$fold.1$fit.supervised$pred_time_shift)])
bb2.pve <- get.pve(bb$fold.2$fit.supervised$dapi[order(bb$fold.2$fit.supervised$pred_time_shift)])
bb3.pve <- get.pve(bb$fold.3$fit.supervised$dapi[order(bb$fold.3$fit.supervised$pred_time_shift)])
bb4.pve <- get.pve(bb$fold.4$fit.supervised$dapi[order(bb$fold.4$fit.supervised$pred_time_shift)])
bb5.pve <- get.pve(bb$fold.5$fit.supervised$dapi[order(bb$fold.5$fit.supervised$pred_time_shift)])


mean(c(bb1.pve$pve, bb2.pve$pve, bb3.pve$pve, bb4.pve$pve, bb5.pve$pve))
```

---

## Load previously computed results

```{r, eval=F}
results_eval_top101 <- readRDS("../data/results/results_eval_top101.rds")
results_eval_top10 <- readRDS("../data/results/results_eval_top10.rds")
```

---

## Before removing moisy labels, on top 10 cyclical genes

```{r, eval=F}
source("../peco/R/utility.R")
source("../peco/R/fit.trendfilter.generic.R")
source("../peco/R/run_seurat.R")

mean(results_eval_top10$fit.supervised$diff_time)/2/pi
mean(results_eval_top10$fit.trend2.unsup$diff_time)/2/pi
mean(results_eval_top10$fit.seurat$diff_time)/2/pi


pve.wrapper <- function(results_list, methods_list) {
  res <- lapply(1:length(results_list), 
       function(i) {
          obj <- results_list[[i]]
          out <- data.frame(
            dapi=with(obj, get.pve(dapi[order(pred_time_shift)])$pve),
            gfp=with(obj, get.pve(gfp[order(pred_time_shift)])$pve),
            rfp=with(obj, get.pve(rfp[order(pred_time_shift)])$pve) )
  })
  names(res) <- methods_list
  return(res)
}


results_list <- results_eval_top10
methods_list <- sapply(names(results_list), function(x) strsplit(x, split=".", fixed=TRUE)[[1]][2])

pve_eval_top10 <- do.call(rbind, pve.wrapper(results_list=results_eval_top10, 
                               methods_list=methods_list))
pve_eval_top10$genes_used <- "top10"
pve_eval_top10$methods <- methods_list
saveRDS(pve_eval_top10,
        "../output/method-eval-withheld-explore.Rmd/pve_eval_top10.rds")
```


```{r}
pve_eval_top10 <- readRDS("../output/method-eval-withheld-explore.Rmd/pve_eval_top10.rds")
print(pve_eval_top10)
```

```{r, eval=F}
source("../peco/R/utility.R")
source("../peco/R/fit.trendfilter.generic.R")
source("../peco/R/run_seurat.R")

get.aov(yy=results_eval_top10$fit.seurat$dapi,
        xx=results_eval_top10$fit.seurat$assignments)

get.aov(yy=results_eval_top10$fit.seurat$gfp,
        xx=results_eval_top10$fit.seurat$assignments)

get.aov(yy=results_eval_top10$fit.seurat$rfp,
        xx=results_eval_top10$fit.seurat$assignments)
```

### Seurat scores vs cell time predictions

```{r, eval=F}
seurat.S.sup <- with(results_eval_top10,
                     get.pve(fit.seurat$S[order(fit.supervised$pred_time_shift)]))
seurat.S.unsup <- with(results_eval_top10,
                     get.pve(fit.seurat$S[order(fit.trend2.unsup$pred_time_shift)]))
seurat.G2M.sup <- with(results_eval_top10,
                     get.pve(fit.seurat$G2M[order(fit.supervised$pred_time_shift)]))
seurat.G2M.unsup <- with(results_eval_top10,
                     get.pve(fit.seurat$G2M[order(fit.trend2.unsup$pred_time_shift)]))

seurat.S.owntime <- with(results_eval_top10,
                     get.pve(fit.seurat$S[order(fit.seurat$cell_times_est)]))
seurat.G2M.owntime <- with(results_eval_top10,
                     get.pve(fit.seurat$G2M[order(fit.seurat$cell_times_est)]))

save(seurat.S.sup, seurat.S.unsup,
     seurat.G2M.sup, seurat.G2M.unsup,
     seurat.S.owntime, seurat.G2M.owntime,
     file = "../output/method-eval-withheld-explore.Rmd/seurat.time.top10.rda")
```


```{r}
load(file="../output/method-eval-withheld-explore.Rmd/seurat.time.top10.rda")

c(seurat.S.sup$pve, seurat.G2M.sup$pve)
c(seurat.S.unsup$pve, seurat.G2M.unsup$pve)
c(seurat.S.owntime$pve, seurat.G2M.owntime$pve)


# with(results_eval_top10,
#   get.aov(yy=fit.seurat$G2M,xx=fit.seurat$assignments))
# 
# with(results_eval_top10,
#   get.aov(yy=fit.seurat$S,xx=fit.seurat$assignments))
```


### Seurat cell class vs unsupervsied predicted time

```{r, eval=F}
cols <- c("orange", "red", "brown")
par(mfrow=c(1,3))
with(results_eval_top10, 
     hist(fit.trend2.unsup$pred_time_shift[fit.seurat$assignments=="G1"],
     nclass=5, col=cols[1], main = "G1, 10 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
with(results_eval_top10, hist(fit.trend2.unsup$pred_time_shift[fit.seurat$assignments=="S"],
     nclass=10, col=cols[2],
     main = "S, 66 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
with(results_eval_top10, hist(fit.trend2.unsup$pred_time_shift[fit.seurat$assignments=="G2M"],
     nclass=10, col=cols[3],
     main = "G2M, 57 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
title("Unsupervised cell time and Seurat classes", outer=TRUE, line=-1)
```


### Seurat clases vs supervsied cell times

```{r, eval=F}
cols <- c("orange", "red", "brown")
par(mfrow=c(1,3))
with(results_eval_top10, 
     hist(fit.supervised$pred_time_shift[fit.seurat$assignments=="G1"],
     nclass=5, col=cols[1], main = "G1, 10 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
with(results_eval_top10, hist(fit.supervised$pred_time_shift[fit.seurat$assignments=="S"],
     nclass=10, col=cols[2],
     main = "S, 66 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
with(results_eval_top10, hist(fit.supervised$pred_time_shift[fit.seurat$assignments=="G2M"],
     nclass=10, col=cols[3],
     main = "G2M, 57 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
title("Supervised cell time and Seurat classes", outer=TRUE, line=-1)
```


---



## Before removing moisy labels, on top 101 cyclical genes

```{r, eval=F}
source("../peco/R/utility.R")
source("../peco/R/fit.trendfilter.generic.R")
source("../peco/R/run_seurat.R")

mean(results_eval_top101$fit.supervised$diff_time)/2/pi
mean(results_eval_top101$fit.trend2.unsup$diff_time)/2/pi
mean(results_eval_top101$fit.seurat$diff_time)/2/pi


pve.wrapper <- function(results_list, methods_list) {
  res <- lapply(1:length(results_list), 
       function(i) {
          obj <- results_list[[i]]
          out <- data.frame(
            dapi=with(obj, get.pve(dapi[order(pred_time_shift)])$pve),
            gfp=with(obj, get.pve(gfp[order(pred_time_shift)])$pve),
            rfp=with(obj, get.pve(rfp[order(pred_time_shift)])$pve) )
  })
  names(res) <- methods_list
  return(res)
}


results_list <- results_eval_top101
methods_list <- sapply(names(results_list), function(x) strsplit(x, split=".", fixed=TRUE)[[1]][2])

pve_eval_top101 <- do.call(rbind, pve.wrapper(results_list=results_eval_top101, 
                               methods_list=methods_list))
pve_eval_top101$genes_used <- "top101"
pve_eval_top101$methods <- methods_list
saveRDS(pve_eval_top101,
        "../output/method-eval-withheld-explore.Rmd/pve_eval_top101.rds")
```


```{r}
pve_eval_top101 <- readRDS("../output/method-eval-withheld-explore.Rmd/pve_eval_top101.rds")
print(pve_eval_top101)
```

```{r, eval=F}
get.aov(yy=results_eval_top101$fit.seurat$dapi,
        xx=results_eval_top101$fit.seurat$assignments)

get.aov(yy=results_eval_top101$fit.seurat$gfp,
        xx=results_eval_top101$fit.seurat$assignments)

get.aov(yy=results_eval_top101$fit.seurat$rfp,
        xx=results_eval_top101$fit.seurat$assignments)
```

### Seurat scores vs cell time predictions

```{r, eval=F}
seurat.S.sup <- with(results_eval_top101,
                     get.pve(fit.seurat$S[order(fit.supervised$pred_time_shift)]))
seurat.S.unsup <- with(results_eval_top101,
                     get.pve(fit.seurat$S[order(fit.trend2.unsup$pred_time_shift)]))
seurat.G2M.sup <- with(results_eval_top101,
                     get.pve(fit.seurat$G2M[order(fit.supervised$pred_time_shift)]))
seurat.G2M.unsup <- with(results_eval_top101,
                     get.pve(fit.seurat$G2M[order(fit.trend2.unsup$pred_time_shift)]))

seurat.S.owntime <- with(results_eval_top101,
                     get.pve(fit.seurat$S[order(fit.seurat$cell_times_est)]))
seurat.G2M.owntime <- with(results_eval_top101,
                     get.pve(fit.seurat$G2M[order(fit.seurat$cell_times_est)]))

save(seurat.S.sup, seurat.S.unsup,
     seurat.G2M.sup, seurat.G2M.unsup,
     seurat.S.owntime, seurat.G2M.owntime,
     file = "../output/method-eval-withheld-explore.Rmd/seurat.time.top101.rda")
```


```{r}
load(file="../output/method-eval-withheld-explore.Rmd/seurat.time.top101.rda")

c(seurat.S.sup$pve, seurat.G2M.sup$pve)
c(seurat.S.unsup$pve, seurat.G2M.unsup$pve)
c(seurat.S.owntime$pve, seurat.G2M.owntime$pve)


# with(results_eval_top101,
#   get.aov(yy=fit.seurat$G2M,xx=fit.seurat$assignments))
# with(results_eval_top101,
#   get.aov(yy=fit.seurat$S,xx=fit.seurat$assignments))

```


### Seurat cell class vs unsupervsied predicted time

```{r, eval=F}
cols <- c("orange", "red", "brown")
par(mfrow=c(1,3))
with(results_eval_top101, 
     hist(fit.trend2.unsup$pred_time_shift[fit.seurat$assignments=="G1"],
     nclass=5, col=cols[1], main = "G1, 10 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
with(results_eval_top101, hist(fit.trend2.unsup$pred_time_shift[fit.seurat$assignments=="S"],
     nclass=10, col=cols[2],
     main = "S, 66 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
with(results_eval_top101, hist(fit.trend2.unsup$pred_time_shift[fit.seurat$assignments=="G2M"],
     nclass=10, col=cols[3],
     main = "G2M, 57 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
title("Unsupervised cell time and Seurat classes", outer=TRUE, line=-1)
```


### Seurat clases vs supervsied cell times

```{r, eval=F}
cols <- c("orange", "red", "brown")
par(mfrow=c(1,3))
with(results_eval_top101, 
     hist(fit.supervised$pred_time_shift[fit.seurat$assignments=="G1"],
     nclass=5, col=cols[1], main = "G1, 10 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
with(results_eval_top101, hist(fit.supervised$pred_time_shift[fit.seurat$assignments=="S"],
     nclass=10, col=cols[2],
     main = "S, 66 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
with(results_eval_top101, hist(fit.supervised$pred_time_shift[fit.seurat$assignments=="G2M"],
     nclass=10, col=cols[3],
     main = "G2M, 57 cells", xlim=c(0,2*pi), ylim=c(0,30),
     xlab="Predicted cell time"))
title("Supervised cell time and Seurat classes", outer=TRUE, line=-1)
```




---


```{r}
sessionInfo()
```

