---
title: "Training model using using trendiflter"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

## Packages

```{r}
library(Biobase)
```

---

## Use the top 10 genes identified to have cyclical expression patterns, in one training dataset

Get data

```{r}
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")

log2cpm.all <- log2cpm.all[,order(pdata$theta)]
pdata <- pdata[order(pdata$theta),]

log2cpm.quant <- readRDS("../output/npreg-trendfilter-quantile.Rmd/log2cpm.quant.rds")
# import previously identifid cell cycle genes
# cyclegenes <- readRDS("../output/npreg-methods.Rmd/cyclegenes.rds")
# cyclegenes.names <- colnames(cyclegenes)[2:6]

# select external validation samples
set.seed(99)
nvalid <- round(ncol(log2cpm.quant)*.15)
ii.valid <- sample(1:ncol(log2cpm.quant), nvalid, replace = F)
ii.nonvalid <- setdiff(1:ncol(log2cpm.quant), ii.valid)

log2cpm.quant.nonvalid <- log2cpm.quant[,ii.nonvalid]
log2cpm.quant.valid <- log2cpm.quant[,ii.valid]
theta <- pdata$theta
names(theta) <- rownames(pdata)
theta.nonvalid <- theta[ii.nonvalid]
theta.valid <- theta[ii.valid]


sig.genes <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.rds")
expr.sig <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% rownames(sig.genes), ]

# set training samples
source("../peco/R/primes.R")
source("../peco/R/partitionSamples.R")
parts <- partitionSamples(1:ncol(log2cpm.quant.nonvalid), runs=5,
                          nsize.each = rep(151,5))
part_indices <- parts$partitions
```

```{r}
source("../peco/R/fit.trendfilter.generic.R")
source("../peco/R/cycle.npreg.R")
source("../peco/R/cycle.corr.R")

Y_train <- expr.sig[,part_indices[[1]]$train]
theta_train <- theta.nonvalid[part_indices[[1]]$train]
```


* The shape of the predicted expression levels is similar between results based on fucci-labels and reuslts based on PCA of the expression levels. Hence, no need to reverse the axis.

```{r, eval=F}
Y_train <- expr.sig[,part_indices[[1]]$train]
theta_train <- theta.nonvalid[part_indices[[1]]$train]
theta_train_pca <- initialize_cell_times(Y_train)
```

```{r, eval = F}
fit.train.nobin <- cycle.npreg.insample(Y = Y_train, 
                                  theta = theta_train, 
                                  nbins = NULL, ncores=10)

fit.train.pca.nobin <- cycle.npreg.insample(Y = Y_train, 
                                  theta = theta_train_pca, 
                                  nbins = NULL, ncores=10)

saveRDS(fit.train.nobin, "../output/method-npreg.Rmd/fit.train.nobin.rds")
saveRDS(fit.train.pca.nobin, "../output/method-npreg.Rmd/fit.train.pca.nobin.rds")
```


```{r}
fit.train.nobin <- readRDS("../output/method-npreg.Rmd/fit.train.nobin.rds")
fit.train.pca.nobin <- readRDS("../output/method-npreg.Rmd/fit.train.pca.nobin.rds")
fit.train.nobin$loglik_est
fit.train.pca.nobin$loglik_est

par(mfcol=c(2,5))
for (g in 1:5) {
  plot(fit.train.nobin$Y_ordered[g,])
  points(fit.train.nobin$mu_est[g,], col="blue", cex=.6, pch=16)
  plot(fit.train.pca.nobin$Y_ordered[g,])
  points(fit.train.pca.nobin$mu_est[g,], col="blue", cex=.6, pch=16)
}
```

Fisher-Lee correlation coefficient for rotational dependence supports a significant rotational dependency between the cell times re-estimated based on FUCCI and the cell times re-estimated based on PCA.

```{r}
par(mfrow=c(1,1))
plot(fit.train.nobin$cell_times_est, 
     fit.train.pca.nobin$cell_times_est[match(names(fit.train.nobin$cell_times_est),
                                        names(fit.train.pca.nobin$cell_times_est))])

rtest_pval <- rFL.IndTestRand(fit.train.nobin$cell_times_est,
                     fit.train.pca.nobin$cell_times_est[match(names(fit.train.nobin$cell_times_est),
                                                  names(fit.train.pca.nobin$cell_times_est))],
                     NR=9999)
# rtest_boot <- rhoFLCIBoot(fit.train.nobin$cell_times_est,
#                   fit.train.pca.nobin$cell_times_est[match(names(fit.train.nobin$cell_times_est),
#                                               names(fit.train.pca.nobin$cell_times_est))],
#                           ConfLevel = 95, B=9999)
# 
# rtest_js <- JSTestRand(fit.train.nobin$cell_times_est,
#               fit.train.pca.nobin$cell_times_est[match(names(fit.train.nobin$cell_times_est),
#                                                 names(fit.train.pca.nobin$cell_times_est))], 
#               NR=9999)
rtest_pval
```

Fisher-Lee correlation coefficient for rotational dependence supports a significant rotational dependency between the FUCCI cell times and cell times re-estimated based on FUCCI.

```{r}
par(mfrow=c(1,1))
plot(fit.train.nobin$cell_times_est, 
     theta_train[match(names(fit.train.nobin$cell_times_est),
                                        names(theta_train))])

rtest_pval <- rFL.IndTestRand(fit.train.nobin$cell_times_est,
                             theta_train[match(names(fit.train.nobin$cell_times_est),
                                        names(theta_train))],
                             NR=9999)

# rtest_boot <- rhoFLCIBoot(fit.train.nobin$cell_times_est,
#                           theta_train[match(names(fit.train.nobin$cell_times_est),
#                                       names(theta_train))],
#                           ConfLevel = 95, B=9999)
# # a version in the circular package
# rtest_js <- JSTestRand(fit.train.nobin$cell_times_est,
#                        theta_train[match(names(fit.train.nobin$cell_times_est),
#                                     names(theta_train))], NR=9999)

rtest_pval
# rtest_boot
# rtest_js
```


---

## Consider the above fitting applying to the test sample

Compare for test samples, the ablity to predict cell times 1) using cyclical patterns for each gene predicted from the training samples initialized by fucci-labels, 2) using cyclical patterns from each gene predicted from training samples initialized by PCA, 3) using the cyclial patterns for each gene in the test samples initialized by fucc-labels, 4) using the cyclial patterns for each gene in the test samples initialized by PCA.

```{r, eval=T}
Y_test <- expr.sig[,part_indices[[1]]$test]
theta_test <- theta.nonvalid[part_indices[[1]]$test]
#theta_test_pca <- initialize_cell_times(Y_test)
```

```{r, eval=F}
fit.test.bytrain.fucci.bin <- vector("list", 5)
for (i in 1:5) {
  fit.test.bytrain.fucci.bin[[i]] <- cycle.npreg.outsample(Y_test,
                                  theta_est=fit.train.bin$cell_times_est,
                                  mu_est=fit.train.bin$mu_est, 
                                  sigma_est=fit.train.bin$sigma_est)
}

saveRDS(fit.test.bytrain.fucci.bin,
        "../output/method-npreg.Rmd/fit.test.bytrain.fucci.bin.rds")
```



```{r, eval = F}
## not specifying bins when predicting
fit.test.bytrain.fucci.nobin <- cycle.npreg.outsample(Y_test,
                                  theta_est=fit.train.nobin$cell_times_est,
                                  mu_est=fit.train.nobin$mu_est, 
                                  sigma_est=fit.train.nobin$sigma_est)
saveRDS(fit.test.bytrain.fucci,
        "../output/method-npreg.Rmd/fit.test.bytrain.fucci.nobin.rds")

fit.test.insample.fucci.nobin <- cycle.npreg.insample(Y = Y_test, 
                                                theta = theta_test, 
                                                nbins = NULL, ncores=10)
saveRDS(fit.test.insample.fucci.nobin, 
        "../output/method-npreg.Rmd/fit.test.insample.fucci.nobin.rds")


fit.test.insample.fucci.bin <- cycle.npreg.insample(Y = Y_test, 
                                                theta = theta_test, 
                                                nbins = 100, ncores=10)
saveRDS(fit.test.insample.fucci.bin, 
        "../output/method-npreg.Rmd/fit.test.insample.fucci.bin.rds")


fit.train.bin <- cycle.npreg.insample(Y = Y_train, 
                                  theta = theta_train, 
                                  nbins = 100, ncores=10)

saveRDS(fit.train.bin, "../output/method-npreg.Rmd/fit.train.bin.rds")

fit.test.bytrain.fucci.bin <- cycle.npreg.outsample(Y_test,
                                  theta_est=fit.train.bin$cell_times_est,
                                  mu_est=fit.train.bin$mu_est, 
                                  sigma_est=fit.train.bin$sigma_est)
saveRDS(fit.test.bytrain.fucci.bin,
        "../output/method-npreg.Rmd/fit.test.bytrain.fucci.bin.rds")

```


```{r}
fit.test.bytrain.fucci.nobin <- readRDS("../output/method-npreg.Rmd/fit.test.bytrain.fucci.nobin.rds")
fit.test.bytrain.fucci.bin <- readRDS("../output/method-npreg.Rmd/fit.test.bytrain.fucci.bin.rds")
fit.test.insample.fucci <- readRDS("../output/method-npreg.Rmd/fit.test.insample.fucci.rds")
fit.test.insample.fucci.bin <- readRDS("../output/method-npreg.Rmd/fit.test.insample.fucci.bin.rds")

par(mfcol=c(2,5))
for (i in 1:10) {
  plot(fit.test.bytrain.fucci.nobin$Y[i,])
  points(fit.test.bytrain.fucci.nobin$mu_est[i,], col = "blue", cex=.6, pch=16)
}

par(mfcol=c(2,5))
for (i in 1:10) {
  plot(fit.test.insample.fucci$Y[i,])
  points(fit.test.insample.fucci$mu_est[i,], col = "blue", cex=.6, pch=16)
}

fit.test.insample.fucci.bin$loglik_est
fit.test.bytrain.fucci.bin$loglik_est
```

Correlation between cell time label: use grid points to estimate gene means in the training sample

```{r}
# JSTestRand(fit.test.bytrain.fucci.nobin$cell_times_est,
#            theta_test[match(names(fit.test.bytrain.fucci.nobin$cell_times_est),
#                             names(theta_test))], NR=9999)

# JSTestRand(fit.test.insample.fucci.nobin$cell_times_est,
           # theta_test[match(names(fit.test.insample.fucci.nobin$cell_times_est),
                            # names(theta_test))], NR=9999)
```

Correlation between cell time labels: not use grid points to estimate gene means in the training sample

```{r}
# JSTestRand(fit.test.bytrain.fucci.bin$cell_times_est,
#            theta_test[match(names(fit.test.bytrain.fucci.bin$cell_times_est),
#                             names(theta_test))], NR=9999)

# JSTestRand(fit.test.insample.fucci.bin$cell_times_est,
#            theta_test[match(names(fit.test.insample.fucci.bin$cell_times_est),
#                             names(theta_test))], NR=9999)

```

---

## Develop a classification performance measure

1. For every sample i in the reference label, identify the nearst k labels at a neighborhood symmetric with respect to the sample i, such that they are at [i-k, i+k]

2. Then, for every sample i in the set for testing, consider also the k nearest neighbors as above.

3. Compute for every sample i the fraction of samples in the neighborhood that are in the reference neighborhood, and the fraction of samples that are not in the neighborhood that are in the reference neighborhood. The ratio of these two gives the fold enrichment of the neighborhood samples that are in the reference neighborhood. 


```{r, eval=F}
circ.dist.neighbors <- function(labels, k) {
  mat_neighbors <- matrix(0, ncol=length(labels), nrow=length(labels))  
  colnames(mat_neighbors) <- labels
  rownames(mat_neighbors) <- labels
  N <- length(labels)
  band <- round(k/2)
  
  for (i in 1:ncol(mat_neighbors)) {
    if (i == 1) {
    neighbors <- c(labels[c((N-band+i):N)], labels[c((i+1):(i+band))])
    mat_neighbors[,i] <- rownames(mat_neighbors) %in% neighbors
    }
    if (i > 1 & i <= band) {
    neighbors <- c(labels[c(1:(i-1), c((N-band+i):N))], labels[c((i+1):(i+band))])
    mat_neighbors[,i] <- rownames(mat_neighbors) %in% neighbors
    }
    if (i > band & i <= (N-band)) {
    neighbors <- c(labels[c((i-band):(i-1))], labels[c((i+1):(i+band))])
    mat_neighbors[,i] <- rownames(mat_neighbors) %in% neighbors
    }
    if (i > (N-band)) {
    neighbors <- c(labels[c((i-band):(i-1))], labels[c((i+1):N, (band -(N-i)):1)])
    mat_neighbors[,i] <- rownames(mat_neighbors) %in% neighbors
    }
    if (i == N) {
    neighbors <- c(labels[c((N-band):(N-1))], labels[c(1:band)])
    mat_neighbors[,i] <- rownames(mat_neighbors) %in% neighbors
    }

  }
  return(mat_neighbors)
}


## test the code
labels_to_test <- names(theta_test)
labels_ref <- names(theta_test)
dist_mat_ref <- circ.dist.neighbors(labels_ref, k=10)
dist_mat_to_test <- circ.dist.neighbors(labels_to_test, k=10)
ii.match <- match(colnames(dist_mat_ref), colnames(dist_mat_to_test))
dist_mat_to_test <- dist_mat_to_test[ii.match, ii.match]

dist_enrich <- sapply(1:N, function(i) {
  lab_self <- rownames(dist_mat_ref)[i]
  labs_ref <- rownames(dist_mat_ref)[which(dist_mat_ref[i,]==1)]
  labs_to_test <- rownames(dist_mat_to_test)[which(dist_mat_to_test[i,]==1)]
  labs_to_test_nonneighbors <- setdiff(rownames(dist_mat_to_test), c(labs_to_test, lab_self))
  prop_neighbors_in_ref <- sum(labs_to_test %in% labs_ref)/length(labs_to_test)
  prop_nonneighbors_in_ref <- sum(labs_to_test_nonneighbors %in% labs_ref)/length(labs_to_test_nonneighbors)
  return(prop_neighbors_in_ref/prop_nonneighbors_in_ref)
})
names(dist_enrich) <- rownames(dist_mat_ref)
summary(dist_enrich)


## try on some data
labels_to_test <- names(fit.test.bytrain.fucci.nobin$cell_times_est)
labels_ref <- names(theta_test)
dist_mat_ref <- circ.dist.neighbors(labels_ref, k=10)
dist_mat_to_test <- circ.dist.neighbors(labels_to_test, k=10)
ii.match <- match(colnames(dist_mat_ref), colnames(dist_mat_to_test))
dist_mat_to_test <- dist_mat_to_test[ii.match, ii.match]

dist_enrich <- sapply(1:N, function(i) {
  lab_self <- rownames(dist_mat_ref)[i]
  labs_ref <- rownames(dist_mat_ref)[which(dist_mat_ref[i,]==1)]
  labs_to_test <- rownames(dist_mat_to_test)[which(dist_mat_to_test[i,]==1)]
  labs_to_test_nonneighbors <- setdiff(rownames(dist_mat_to_test), c(labs_to_test, lab_self))
  prop_neighbors_in_ref <- sum(labs_to_test %in% labs_ref)/length(labs_to_test)
  prop_nonneighbors_in_ref <- sum(labs_to_test_nonneighbors %in% labs_ref)/length(labs_to_test_nonneighbors)
  return(prop_neighbors_in_ref/prop_nonneighbors_in_ref)
})
names(dist_enrich) <- rownames(dist_mat_ref)
summary(dist_enrich)



# compare with predictions obtained from training sample (binned)
labels_to_test <- names(fit.test.bytrain.fucci.bin$cell_times_est)
labels_ref <- names(theta_test)
dist_mat_ref <- circ.dist.neighbors(labels_ref, k=10)
dist_mat_to_test <- circ.dist.neighbors(labels_to_test, k=10)
ii.match <- match(colnames(dist_mat_ref), colnames(dist_mat_to_test))
dist_mat_to_test <- dist_mat_to_test[ii.match, ii.match]

dist_enrich <- sapply(1:N, function(i) {
  lab_self <- rownames(dist_mat_ref)[i]
  labs_ref <- rownames(dist_mat_ref)[which(dist_mat_ref[i,]==1)]
  labs_to_test <- rownames(dist_mat_to_test)[which(dist_mat_to_test[i,]==1)]
  labs_to_test_nonneighbors <- setdiff(rownames(dist_mat_to_test), c(labs_to_test, lab_self))
  prop_neighbors_in_ref <- sum(labs_to_test %in% labs_ref)/length(labs_to_test)
  prop_nonneighbors_in_ref <- sum(labs_to_test_nonneighbors %in% labs_ref)/length(labs_to_test_nonneighbors)
  return(prop_neighbors_in_ref/prop_nonneighbors_in_ref)
})
names(dist_enrich) <- rownames(dist_mat_ref)
summary(dist_enrich)


# compare with predictions obtained from training sample (binned)
labels_to_test <- names(fit.test.insample.fucci.nobin$cell_times_est)
labels_ref <- names(theta_test)
dist_mat_ref <- circ.dist.neighbors(labels_ref, k=10)
dist_mat_to_test <- circ.dist.neighbors(labels_to_test, k=10)
ii.match <- match(colnames(dist_mat_ref), colnames(dist_mat_to_test))
dist_mat_to_test <- dist_mat_to_test[ii.match, ii.match]

dist_enrich <- sapply(1:N, function(i) {
  lab_self <- rownames(dist_mat_ref)[i]
  labs_ref <- rownames(dist_mat_ref)[which(dist_mat_ref[i,]==1)]
  labs_to_test <- rownames(dist_mat_to_test)[which(dist_mat_to_test[i,]==1)]
  labs_to_test_nonneighbors <- setdiff(rownames(dist_mat_to_test), c(labs_to_test, lab_self))
  prop_neighbors_in_ref <- sum(labs_to_test %in% labs_ref)/length(labs_to_test)
  prop_nonneighbors_in_ref <- sum(labs_to_test_nonneighbors %in% labs_ref)/length(labs_to_test_nonneighbors)
  return(prop_neighbors_in_ref/prop_nonneighbors_in_ref)
})
names(dist_enrich) <- rownames(dist_mat_ref)
summary(dist_enrich)

# compare with predictions obtained from in-sample prediction (binned)
labels_to_test <- names(fit.test.insample.fucci.bin$cell_times_est)
labels_ref <- names(theta_test)
dist_mat_ref <- circ.dist.neighbors(labels_ref, k=10)
dist_mat_to_test <- circ.dist.neighbors(labels_to_test, k=10)
ii.match <- match(colnames(dist_mat_ref), colnames(dist_mat_to_test))
dist_mat_to_test <- dist_mat_to_test[ii.match, ii.match]

dist_enrich <- sapply(1:N, function(i) {
  lab_self <- rownames(dist_mat_ref)[i]
  labs_ref <- rownames(dist_mat_ref)[which(dist_mat_ref[i,]==1)]
  labs_to_test <- rownames(dist_mat_to_test)[which(dist_mat_to_test[i,]==1)]
  labs_to_test_nonneighbors <- setdiff(rownames(dist_mat_to_test), c(labs_to_test, lab_self))
  prop_neighbors_in_ref <- sum(labs_to_test %in% labs_ref)/length(labs_to_test)
  prop_nonneighbors_in_ref <- sum(labs_to_test_nonneighbors %in% labs_ref)/length(labs_to_test_nonneighbors)
  return(prop_neighbors_in_ref/prop_nonneighbors_in_ref)
})
names(dist_enrich) <- rownames(dist_mat_ref)
summary(dist_enrich)
```


---

## Running on a larger dataset

Will run the job on cluster. See code in `code/method-npreg.Rmd`.

```{r, eval=T}
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")

log2cpm.all <- log2cpm.all[,order(pdata$theta)]
pdata <- pdata[order(pdata$theta),]

log2cpm.quant <- readRDS("../output/npreg-trendfilter-quantile.Rmd/log2cpm.quant.rds")
# import previously identifid cell cycle genes
# cyclegenes <- readRDS("../output/npreg-methods.Rmd/cyclegenes.rds")
# cyclegenes.names <- colnames(cyclegenes)[2:6]

# select external validation samples
set.seed(99)
nvalid <- round(ncol(log2cpm.quant)*.15)
ii.valid <- sample(1:ncol(log2cpm.quant), nvalid, replace = F)
ii.nonvalid <- setdiff(1:ncol(log2cpm.quant), ii.valid)

log2cpm.quant.nonvalid <- log2cpm.quant[,ii.nonvalid]
log2cpm.quant.valid <- log2cpm.quant[,ii.valid]
theta <- pdata$theta
names(theta) <- rownames(pdata)
theta.nonvalid <- theta[ii.nonvalid]
theta.valid <- theta[ii.valid]


sig.genes <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.rds")
expr.sig <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% rownames(sig.genes)[1:10], ]

# set training samples
source("../peco/R/primes.R")
source("../peco/R/partitionSamples.R")
parts <- partitionSamples(1:ncol(log2cpm.quant.nonvalid), runs=5,
                          nsize.each = rep(151,5))
part_indices <- parts$partitions
```

```{r}
fold.train <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.train.fold.",fold,".rds")) })
fold.test.bytrain.fucci <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.test.bytrain.fucci.fold.", fold, ".rds"))})
fold.test.insample.fucci <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.test.insample.fucci.fold.",fold,".rds"))}) 

do.call(rbind, lapply(1:5, function(f) {
  data.frame(train.fucci=fold.train[[f]]$loglik_est/4,
             test.bytrain.fucci=fold.test.bytrain.fucci[[f]]$loglik_est,
             test.insample.fucci=fold.test.insample.fucci[[f]]$loglik_est)
}))
```

compute two measures to evaluate the results:

1. correlation between labels

2. fold enrichment of neighborhood samples defined in the reference label

```{r, eval = F}
source("../code/utility.R")
eval_cor <- vector("list", 5)
eval_neighbor <- vector("list", 5)

for (f in 1:5) {
  eval_cor[[f]] <- JSTestRand(theta.nonvalid[part_indices[[f]]$test],
                        fold.test.bytrain.fucci[[f]]$cell_times_est, NR=9999)
}
eval_cor <- do.call(rbind, eval_cor)
colnames(eval_cor) <- c("rho", "pval")


for (f in 1:5) {
  labels_ref <- names(theta.nonvalid[part_indices[[f]]$test])
  labels_to_test <- names(fold.test.bytrain.fucci[[f]]$cell_times_est)
  dist_mat_ref <- circ.dist.neighbors(labels_ref, k=10)
  dist_mat_to_test <- circ.dist.neighbors(labels_to_test, k=10)
  ii.match <- match(colnames(dist_mat_ref), colnames(dist_mat_to_test))
  dist_mat_to_test <- dist_mat_to_test[ii.match, ii.match]
  N <- length(labels_ref)
  
  dist_enrich <- sapply(1:N, function(i) {
      lab_self <- rownames(dist_mat_ref)[i]
      labs_ref <- rownames(dist_mat_ref)[which(dist_mat_ref[i,]==1)]
      labs_to_test <- rownames(dist_mat_to_test)[which(dist_mat_to_test[i,]==1)]
#      labs_to_test_nonneighbors <- setdiff(rownames(dist_mat_to_test), c(labs_to_test, lab_self))
      #prop_neighbors_in_ref <- sum(labs_to_test %in% labs_ref)/length(labs_to_test)
      #prop_nonneighbors_in_ref <- sum(labs_to_test_nonneighbors %in% labs_ref)/length(labs_to_test_nonneighbors)
#      return(prop_neighbors_in_ref/prop_nonneighbors_in_ref)
      TP <- sum(dist_mat_ref[i,]==1 & dist_mat_to_test[i,]==1)
      FP <- sum(dist_mat_ref[i,]==0 & dist_mat_to_test[i,]==1)
      FN <- sum(dist_mat_ref[i,]==1 & dist_mat_to_test[i,]==0)
      precision <- TP/(TP+FP)
      recall <- TP/(TP+FN)
      if ((precision+recall)==0) {
        F1score <- 0
      } else {
        F1score <- 2*precision*recall/(precision+recall)
      }
      return(F1score)
    })
  names(dist_enrich) <- rownames(dist_mat_ref)
  
  eval_neighbor[[f]] <- dist_enrich
}
saveRDS(eval_cor, "../output/method-npreg.Rmd/eval_cor.rds")
saveRDS(eval_neighbor, "../output/method-npreg.Rmd/eval_neighbor.rds")
```


```{r}
eval_cor <- readRDS("../output/method-npreg.Rmd/eval_cor.rds")
eval_neighbor <- readRDS("../output/method-npreg.Rmd/eval_neighbor.rds")

eval_cor

lapply(eval_neighbor, function(x) table(x))

eval_neighbor_table <- matrix(0, nrow=5, ncol=8)
for (i in 1:5 ) {
  eval_neighbor_table[i,1:length(table(eval_neighbor[[i]]))] <- as.numeric(table(eval_neighbor[[i]]))
}
colnames(eval_neighbor_table) <- c("0", "0.1", "0.2", "0.3", "0.4", "0.5", "mean", "sd")
#eval_neighbor_table <- do.call(rbind, lapply(eval_neighbor, function(x) table(x)))
eval_neighbor_table[,7] <- sapply(eval_neighbor, function(x) round(mean(x), digits=2))
eval_neighbor_table[,8] <- sapply(eval_neighbor, function(x) round(sd(x), digits=2))
eval_neighbor_table
```


---

## Running on 476 genes

Will run the job on cluster. See code in `code/method-npreg.Rmd`.

```{r, eval=F}
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")

log2cpm.all <- log2cpm.all[,order(pdata$theta)]
pdata <- pdata[order(pdata$theta),]

log2cpm.quant <- readRDS("../output/npreg-trendfilter-quantile.Rmd/log2cpm.quant.rds")
# import previously identifid cell cycle genes
# cyclegenes <- readRDS("../output/npreg-methods.Rmd/cyclegenes.rds")
# cyclegenes.names <- colnames(cyclegenes)[2:6]

# select external validation samples
set.seed(99)
nvalid <- round(ncol(log2cpm.quant)*.15)
ii.valid <- sample(1:ncol(log2cpm.quant), nvalid, replace = F)
ii.nonvalid <- setdiff(1:ncol(log2cpm.quant), ii.valid)

log2cpm.quant.nonvalid <- log2cpm.quant[,ii.nonvalid]
log2cpm.quant.valid <- log2cpm.quant[,ii.valid]
theta <- pdata$theta
names(theta) <- rownames(pdata)
theta.nonvalid <- theta[ii.nonvalid]
theta.valid <- theta[ii.valid]


# sig.genes <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.rds")
# expr.sig <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% rownames(sig.genes)[1:10], ]

# set training samples
source("../peco/R/primes.R")
source("../peco/R/partitionSamples.R")
parts <- partitionSamples(1:ncol(log2cpm.quant.nonvalid), runs=5,
                          nsize.each = rep(151,5))
part_indices <- parts$partitions
```

```{r}
fold.train <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.train.fold.",fold,".all.rds")) })
fold.test.bytrain.fucci <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.test.bytrain.fucci.fold.", fold, ".all.rds"))})
fold.test.insample.fucci <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.test.insample.fucci.fold.",fold,".all.rds"))}) 

do.call(rbind, lapply(1:5, function(f) {
  data.frame(train.fucci=fold.train[[f]]$loglik_est/4,
             test.bytrain.fucci=fold.test.bytrain.fucci[[f]]$loglik_est,
             test.insample.fucci=fold.test.insample.fucci[[f]]$loglik_est)
}))
```

compute two measures to evaluate the results:

1. correlation between labels

2. fold enrichment of neighborhood samples defined in the reference label

```{r, eval = F}
source("../code/utility.R")
eval_cor <- vector("list", 5)
eval_neighbor <- vector("list", 5)

for (f in 1:5) {
  eval_cor[[f]] <- JSTestRand(theta.nonvalid[part_indices[[f]]$test],
                        fold.test.bytrain.fucci[[f]]$cell_times_est, NR=9999)
}
eval_cor <- do.call(rbind, eval_cor)
colnames(eval_cor) <- c("rho", "pval")


for (f in 1:5) {
  labels_ref <- names(theta.nonvalid[part_indices[[f]]$test])
  labels_to_test <- names(fold.test.bytrain.fucci[[f]]$cell_times_est)
  dist_mat_ref <- circ.dist.neighbors(labels_ref, k=10)
  dist_mat_to_test <- circ.dist.neighbors(labels_to_test, k=10)
  ii.match <- match(colnames(dist_mat_ref), colnames(dist_mat_to_test))
  dist_mat_to_test <- dist_mat_to_test[ii.match, ii.match]
  N <- length(labels_ref)
  
  dist_enrich <- sapply(1:N, function(i) {
      lab_self <- rownames(dist_mat_ref)[i]
      labs_ref <- rownames(dist_mat_ref)[which(dist_mat_ref[i,]==1)]
      labs_to_test <- rownames(dist_mat_to_test)[which(dist_mat_to_test[i,]==1)]
#      labs_to_test_nonneighbors <- setdiff(rownames(dist_mat_to_test), c(labs_to_test, lab_self))
      #prop_neighbors_in_ref <- sum(labs_to_test %in% labs_ref)/length(labs_to_test)
      #prop_nonneighbors_in_ref <- sum(labs_to_test_nonneighbors %in% labs_ref)/length(labs_to_test_nonneighbors)
#      return(prop_neighbors_in_ref/prop_nonneighbors_in_ref)
      TP <- sum(dist_mat_ref[i,]==1 & dist_mat_to_test[i,]==1)
      FP <- sum(dist_mat_ref[i,]==0 & dist_mat_to_test[i,]==1)
      FN <- sum(dist_mat_ref[i,]==1 & dist_mat_to_test[i,]==0)
      precision <- TP/(TP+FP)
      recall <- TP/(TP+FN)
      if ((precision+recall)==0) {
        F1score <- 0
      } else {
        F1score <- 2*precision*recall/(precision+recall)
      }
      return(F1score)
    })
  names(dist_enrich) <- rownames(dist_mat_ref)
  
  eval_neighbor[[f]] <- dist_enrich
}

saveRDS(eval_cor, "../output/method-npreg.Rmd/eval_cor.all.rds")
saveRDS(eval_neighbor, "../output/method-npreg.Rmd/eval_neighbor.all.rds")
```


```{r}
eval_cor <- readRDS("../output/method-npreg.Rmd/eval_cor.all.rds")
eval_neighbor <- readRDS("../output/method-npreg.Rmd/eval_neighbor.all.rds")

eval_cor

sapply(eval_neighbor, function(x) table(x))

eval_neighbor_table <- matrix(0, nrow=5, ncol=7)
for (i in 1:5 ) {
  eval_neighbor_table[i,1:length(table(eval_neighbor[[i]]))] <- as.numeric(table(eval_neighbor[[i]]))
}
colnames(eval_neighbor_table) <- c("0", "0.1", "0.2", "0.3", "0.4", "mean", "sd")
#eval_neighbor_table <- do.call(rbind, lapply(eval_neighbor, function(x) table(x)))
eval_neighbor_table[,6] <- sapply(eval_neighbor, function(x) round(mean(x), digits=2))
eval_neighbor_table[,7] <- sapply(eval_neighbor, function(x) round(sd(x), digits=2))
eval_neighbor_table
```



---


## Running on Macosko genes

Will run the job on cluster. See code in `code/method-npreg.Rmd`.

```{r, eval=F}
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")

log2cpm.all <- log2cpm.all[,order(pdata$theta)]
pdata <- pdata[order(pdata$theta),]

log2cpm.quant <- readRDS("../output/npreg-trendfilter-quantile.Rmd/log2cpm.quant.rds")
# import previously identifid cell cycle genes
# cyclegenes <- readRDS("../output/npreg-methods.Rmd/cyclegenes.rds")
# cyclegenes.names <- colnames(cyclegenes)[2:6]

# select external validation samples
set.seed(99)
nvalid <- round(ncol(log2cpm.quant)*.15)
ii.valid <- sample(1:ncol(log2cpm.quant), nvalid, replace = F)
ii.nonvalid <- setdiff(1:ncol(log2cpm.quant), ii.valid)

log2cpm.quant.nonvalid <- log2cpm.quant[,ii.nonvalid]
log2cpm.quant.valid <- log2cpm.quant[,ii.valid]
theta <- pdata$theta
names(theta) <- rownames(pdata)
theta.nonvalid <- theta[ii.nonvalid]
theta.valid <- theta[ii.valid]


# sig.genes <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.rds")
# expr.sig <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% rownames(sig.genes)[1:10], ]

# set training samples
source("../peco/R/primes.R")
source("../peco/R/partitionSamples.R")
parts <- partitionSamples(1:ncol(log2cpm.quant.nonvalid), runs=5,
                          nsize.each = rep(151,5))
part_indices <- parts$partitions
```

```{r}
fold.train <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.train.fold.",fold,".macosko.rds")) })
fold.test.bytrain.fucci <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.test.bytrain.fucci.fold.", fold, ".macosko.rds"))})
fold.test.insample.fucci <- lapply(1:5, function(fold) {
  readRDS(paste0("../output/method-npreg.Rmd/fold.test.insample.fucci.fold.",fold,".macosko.rds"))}) 

do.call(rbind, lapply(1:5, function(f) {
  data.frame(train.fucci=fold.train[[f]]$loglik_est/4,
             test.bytrain.fucci=fold.test.bytrain.fucci[[f]]$loglik_est,
             test.insample.fucci=fold.test.insample.fucci[[f]]$loglik_est)
}))
```

compute two measures to evaluate the results:

1. correlation between labels

2. fold enrichment of neighborhood samples defined in the reference label

```{r, eval = F}
source("../code/utility.R")
eval_cor <- vector("list", 5)
eval_neighbor <- vector("list", 5)

for (f in 1:5) {
  eval_cor[[f]] <- JSTestRand(theta.nonvalid[part_indices[[f]]$test],
                        fold.test.bytrain.fucci[[f]]$cell_times_est, NR=9999)
}
eval_cor <- do.call(rbind, eval_cor)
colnames(eval_cor) <- c("rho", "pval")


for (f in 1:5) {
  labels_ref <- names(theta.nonvalid[part_indices[[f]]$test])
  labels_to_test <- names(fold.test.bytrain.fucci[[f]]$cell_times_est)
  dist_mat_ref <- circ.dist.neighbors(labels_ref, k=10)
  dist_mat_to_test <- circ.dist.neighbors(labels_to_test, k=10)
  ii.match <- match(colnames(dist_mat_ref), colnames(dist_mat_to_test))
  dist_mat_to_test <- dist_mat_to_test[ii.match, ii.match]
  N <- length(labels_ref)
  
  dist_enrich <- sapply(1:N, function(i) {
      lab_self <- rownames(dist_mat_ref)[i]
      labs_ref <- rownames(dist_mat_ref)[which(dist_mat_ref[i,]==1)]
      labs_to_test <- rownames(dist_mat_to_test)[which(dist_mat_to_test[i,]==1)]
#      labs_to_test_nonneighbors <- setdiff(rownames(dist_mat_to_test), c(labs_to_test, lab_self))
      #prop_neighbors_in_ref <- sum(labs_to_test %in% labs_ref)/length(labs_to_test)
      #prop_nonneighbors_in_ref <- sum(labs_to_test_nonneighbors %in% labs_ref)/length(labs_to_test_nonneighbors)
#      return(prop_neighbors_in_ref/prop_nonneighbors_in_ref)
      TP <- sum(dist_mat_ref[i,]==1 & dist_mat_to_test[i,]==1)
      FP <- sum(dist_mat_ref[i,]==0 & dist_mat_to_test[i,]==1)
      FN <- sum(dist_mat_ref[i,]==1 & dist_mat_to_test[i,]==0)
      precision <- TP/(TP+FP)
      recall <- TP/(TP+FN)
      if ((precision+recall)==0) {
        F1score <- 0
      } else {
        F1score <- 2*precision*recall/(precision+recall)
      }
      return(F1score)
    })
  names(dist_enrich) <- rownames(dist_mat_ref)
  
  eval_neighbor[[f]] <- dist_enrich
}

saveRDS(eval_cor, "../output/method-npreg.Rmd/eval_cor.macosko.rds")
saveRDS(eval_neighbor, "../output/method-npreg.Rmd/eval_neighbor.macosko.rds")
```


```{r}
eval_cor <- readRDS("../output/method-npreg.Rmd/eval_cor.macosko.rds")
eval_neighbor <- readRDS("../output/method-npreg.Rmd/eval_neighbor.macosko.rds")

eval_cor

lapply(eval_neighbor, function(x) summary(x))
lapply(eval_neighbor, function(x) table(x))
eval_neighbor_table <- matrix(0, nrow=5, ncol=7)
for (i in 1:5 ) {
  eval_neighbor_table[i,1:length(table(eval_neighbor[[i]]))] <- as.numeric(table(eval_neighbor[[i]]))
}
colnames(eval_neighbor_table) <- c("0", "0.1", "0.2", "0.3", "0.4", "mean", "sd")
#eval_neighbor_table <- do.call(rbind, lapply(eval_neighbor, function(x) table(x)))
eval_neighbor_table[,6] <- sapply(eval_neighbor, function(x) round(mean(x),digits=2))
eval_neighbor_table[,7] <- sapply(eval_neighbor, function(x) round(sd(x),digits=2))
eval_neighbor_table
```




---


```{r}
sessionInfo()
```

