---
title: "Evaluate training labels"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Results

* First we get a sense of the associations between the intensities. We do this by computing PVE of one intensities given another. 

* Then, we compute cell time labels based on GFP, RPF and based on DAPI, GFP, RFP. We then compute the PVE of intensities by these two sets of labels. Three together don't do better in explaining RFP and GFP than two together. 

* Then we consider in the training dataset, the associations between the PC score of the samples versus the predicted cell times. Are there more errors for cells at the center than the others?
    * I looked at prediction error vs PC scores and dont' see a clear association. To the point that there's no need to statistical analysis.
    * I then compute each cell's point to the origin on the PC plot and see if prediction error is a function of distance of the origin. Again there doesn't seem to be an obvious pattern for the relationship between prediction error and distance to the origin.
    

* Notes: predicted time mostly doesn't need rotations


---

## Extract data from the top 101 genes identified 

```{r}
library(Biobase)
source("../peco/R/cycle.corr.R")
source("../peco/R/cycle.npreg.R")

df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

# compare three cell times
theta_pca_triple <- prcomp(cbind(pdata$dapi.median.log10sum.adjust,
                                     pdata$rfp.median.log10sum.adjust,
                                pdata$gfp.median.log10sum.adjust), scale=TRUE)
library(circular)
pdata$theta_triple <- coord2rad(cbind(theta_pca_triple$x[,1],
                                    theta_pca_triple$x[,2]))

pdata$theta_triple <- as.numeric(pdata$theta_triple)

y1=pdata$theta
y2=pdata$theta_triple
y2shift <- rotation(y1, y2)

pdata$theta_triple <- y2shift

par(mfrow=c(1,1))
plot(y1,
     y2shift, main = "post rotation",
     xlab="based on GFP, RFP",
     ylab="based on DAPI, GFP, RFP")
```

fucci time distribution

```{r}
hist(pdata$theta, breaks=40,
     xlim=c(0,6.5), xlab="fucci time",
     main = "fucci time distribution")
```

intensity time distribution

```{r}
hist(pdata$theta_triple, breaks=40,
     xlim=c(0,6.5), xlab="intensity time",
     main = "intensity time distribution")
```


intensities by fucci time

```{r}
library(scales)
plot(pdata$theta,
     pdata$dapi.median.log10sum.adjust,
     col=alpha("blue", .5), cex=.7, pch=16,
     ylim=range(c(pdata$rfp.median.log10sum.adjust,
                  pdata$gfp.median.log10sum.adjust,
                  pdata$dapi.median.log10sum.adjust)),
     xlab="fucci time",
     ylab="batch-adjusted log intensites",
     main ="Intensities by fucci time")
points(pdata$theta,
     pdata$gfp.median.log10sum.adjust,
     col=alpha("forestgreen", .5), cex=.7, pch=16)
points(pdata$theta,
     pdata$rfp.median.log10sum.adjust,
     col=alpha("red", .5), cex=.7, pch=16)



plot(pdata$theta_triple,
     pdata$dapi.median.log10sum.adjust,
     col=alpha("blue", .5), cex=.7, pch=16,
     ylim=range(c(pdata$rfp.median.log10sum.adjust,
                  pdata$gfp.median.log10sum.adjust,
                  pdata$dapi.median.log10sum.adjust)),
     xlab="intensity time (fucci + daip)",
     ylab="batch-adjusted log intensites",
     main ="Intensities by intensity time")
points(pdata$theta,
     pdata$gfp.median.log10sum.adjust,
     col=alpha("forestgreen", .5), cex=.7, pch=16)
points(pdata$theta,
     pdata$rfp.median.log10sum.adjust,
     col=alpha("red", .5), cex=.7, pch=16)

# rr <- rFLIndTestRand(pdata$theta, pdata$theta_triple, NR=10000)
# cor(y1, y2shift)



# select external validation samples
log2cpm.quant <- readRDS("../output/npreg-trendfilter-quantile.Rmd/log2cpm.quant.rds")

set.seed(99)
nvalid <- round(ncol(log2cpm.quant)*.15)
ii.valid <- sample(1:ncol(log2cpm.quant), nvalid, replace = F)
ii.nonvalid <- setdiff(1:ncol(log2cpm.quant), ii.valid)

log2cpm.quant.nonvalid <- log2cpm.quant[,ii.nonvalid]
log2cpm.quant.valid <- log2cpm.quant[,ii.valid]

sig.genes <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.101.rds")
expr.sig <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% rownames(sig.genes), ]


# get predicted times
# set training samples
source("../peco/R/primes.R")
source("../peco/R/partitionSamples.R")
parts <- partitionSamples(1:ncol(log2cpm.quant.nonvalid), runs=5,
                          nsize.each = rep(151,5))
part_indices <- parts$partitions
```



---


## PVE between intensities in all samples

```{r}
source("../peco/R/fit.trendfilter.generic.R")
source("../peco/R/utility.R")

with(pdata, summary(lm(dapi.median.log10sum.adjust~rfp.median.log10sum.adjust))$adj.r.squared)
with(pdata, summary(lm(dapi.median.log10sum.adjust~gfp.median.log10sum.adjust))$adj.r.squared)

with(pdata, summary(lm(rfp.median.log10sum.adjust~gfp.median.log10sum.adjust))$adj.r.squared)
with(pdata, summary(lm(gfp.median.log10sum.adjust~rfp.median.log10sum.adjust))$adj.r.squared)

with(pdata, summary(lm(rfp.median.log10sum.adjust~dapi.median.log10sum.adjust))$adj.r.squared)
with(pdata, summary(lm(gfp.median.log10sum.adjust~dapi.median.log10sum.adjust))$adj.r.squared)

plot(pdata$gfp.median.log10sum.adjust,
     pdata$rfp.median.log10sum.adjust,
     xlab="GFP log10 intensities",
     ylab="RFP log10 intensities",
     main = "RFP vs. GFP")
```

---

## PVE of cell times in all samples

```{r, eval=F}
rfp_theta <- with(pdata, get.pve(pdata$rfp.median.log10sum.adjust[order(pdata$theta)]))
rfp_theta_triple <- with(pdata, get.pve(pdata$rfp.median.log10sum.adjust[order(pdata$theta_triple)]))

gfp_theta <- with(pdata, get.pve(pdata$gfp.median.log10sum.adjust[order(pdata$theta)]))
gfp_theta_triple <- with(pdata, get.pve(pdata$gfp.median.log10sum.adjust[order(pdata$theta_triple)]))

dapi_theta <- with(pdata, get.pve(pdata$dapi.median.log10sum.adjust[order(pdata$theta)]))
dapi_theta_triple <- with(pdata, get.pve(pdata$dapi.median.log10sum.adjust[order(pdata$theta_triple)]))

save(rfp_theta, rfp_theta_triple,
     gfp_theta, gfp_theta_triple,
     dapi_theta, dapi_theta_triple,
     file = "../output/method-labels.Rmd/pve.rda")
```


```{r}
load(file = "../output/method-labels.Rmd/pve.rda")

c(rfp_theta, rfp_theta_triple,
     gfp_theta, gfp_theta_triple,
     dapi_theta, dapi_theta_triple)
```


---

## PC scores of samples versus prediction error

Fitting using 101 genes

```{r}
# first check the theta in pdata
pca <- prcomp(cbind(pdata$rfp.median.log10sum.adjust,
                    pdata$gfp.median.log10sum.adjust), scale=TRUE)
pca_df <- cbind(pca$x[,1],pca$x[,2])
rownames(pca_df) <- rownames(pdata)

theta_check <- as.numeric(coord2rad(pca_df))
theta_check <- 2*pi-theta_check
plot(theta_check, pdata$theta)
names(theta_check) <- rownames(pdata)
```

```{r, eval =F}
fits <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.sig[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 

  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr.sig[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test)
}

for (i in 1:5) {
  fits[[i]]$theta_est_shift <- rotation(fits[[i]]$theta_test, fits[[i]]$fit.test$cell_times_est)$y2shift
}
  
saveRDS(fits, file = "../output/method-labels.Rmd/fits.rds")
```


```{r}
fits <- readRDS(file = "../output/method-labels.Rmd/fits.rds")
source("../peco/R/utility.R")

diff_time <- lapply(1:length(fits), function(i) {
  pmin(abs(fits[[i]]$theta_est_shift-fits[[i]]$theta_test),
    abs(fits[[i]]$theta_est_shift-(2*pi-fits[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits[[i]]$theta_test),rownames(pdata))]
  get.pve(dap[order(fits[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```

explore PCA outliers

```{r}
pca_df_sub <- pca_df[match(names(fits[[2]]$theta_test),rownames(pca_df)),] 
diff_bins <- cut(diff_time[[2]], 
                 breaks=quantile(diff_time[[2]], prob=seq(0,1,.25)), include.lowest=TRUE)
summary(diff_bins)
.43/2/pi

library(RColorBrewer)
library(scales)
cols <- brewer.pal(9, "YlGn")[c(3, 5, 7, 9)]
par(mfrow=c(1,2))
plot(fits[[2]]$theta_test,
   diff_time[[2]]/2/pi, pch=16,
   xlab="cell time labels", ylab="margin of error (% arc length)",
   col=alpha(c("gray50", "royalblue")[(as.numeric(diff_bins)==4|as.numeric(diff_bins)==3)+1],.7),
   main="Prediction error")
plot(pca_df_sub[,1], pca_df_sub[,2],
     col=alpha(c("gray50", "royalblue")[(as.numeric(diff_bins)==4|as.numeric(diff_bins)==3)+1],.7), pch=16,
     xlab="PC1", ylab="PC2",
     main="PCs based on GFP, RFP")
abline(h=0,v=0, lty=3)
title(main="blue dots: diff time > 50th percentile", outer=TRUE, line=-1)


par(mfrow=c(1,2))
dist_to_origin <- sqrt(pca_df_sub[,1]^2+pca_df_sub[,2]^2)
#which(abs(scale(dist_to_origin))>2)

plot(dist_to_origin, diff_time[[2]]/2/pi,
     xlab="Distance to the origin in the PC1 vs PC2 space",
     ylab="Maring of error (% arc length)",
     main = "Prediction error", pch=16,
     col=alpha(c("gray50", "forestgreen"),.7)[(abs(scale(dist_to_origin))>1)+1])

plot(pca_df_sub[,1], pca_df_sub[,2],
     col=alpha(c("gray50", "forestgreen"),.7)[(abs(scale(dist_to_origin))>1)+1], 
     pch=16,
     xlab="PC1", ylab="PC2",
     main="PC1 vs. PC2 based on GFP, RFP")
abline(h=0,v=0, lty=3)

title(main="Green points: 1 +/- SD of mean distance to the origin", outer=TRUE, line=-1)
```

---

## Re-fitting after removing PC outliers, remove outliers from the same 5 folds


```{r}
# first check the theta in pdata
pca <- prcomp(cbind(pdata$rfp.median.log10sum.adjust,
                    pdata$gfp.median.log10sum.adjust), scale=TRUE)
pca_df <- cbind(pca$x[,1],pca$x[,2])
rownames(pca_df) <- rownames(pdata)

theta_check <- as.numeric(coord2rad(pca_df))
theta_check <- 2*pi-theta_check
plot(theta_check, pdata$theta)
names(theta_check) <- rownames(pdata)

dist_to_origin <- sqrt(pca_df[,1]^2+pca_df[,2]^2)
which_out <- rownames(pdata)[which(abs(scale(dist_to_origin))>1)]

plot(pca_df[,1],
     pca_df[,2],
     col=c("gray50", "forestgreen")[(abs(scale(dist_to_origin))>1)+1], pch=16,
     xlab="PC1",
     ylab="PC2",
     main="abs(Distance-to-Origin) > +/- 1")
```


```{r, eval=F}
fits_sub <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.sig[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 
  
  Y_train_sub <- Y_train[,which(!(colnames(Y_train) %in% which_out))]
  theta_train_sub <- theta_train[which(!(names(theta_train) %in% which_out))]

  fit.train <- cycle.npreg.insample(Y = Y_train_sub, 
                                    theta = theta_train_sub, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr.sig[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  Y_test_sub <- Y_test[,which(!(colnames(Y_test) %in% which_out))]
  theta_test_sub <- theta_test[which(!(names(theta_test) %in% which_out))]

  fit.test <- cycle.npreg.outsample(Y_test=Y_test_sub,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_sub[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test_sub)
}

for (i in 1:5) {
  fits_sub[[i]]$theta_est_shift <- rotation(fits_sub[[i]]$theta_test, fits_sub[[i]]$fit.test$cell_times_est)$y2shift
}

saveRDS(fits_sub, file = "../output/method-labels.Rmd/fits_sub.rds")
```


```{r}
fits_sub <- readRDS(file = "../output/method-labels.Rmd/fits_sub.rds")

diff_time <- lapply(1:5, function(i) {
  pmin(abs(fits_sub[[i]]$theta_est_shift-fits_sub[[i]]$theta_test),
    abs(fits_sub[[i]]$theta_est_shift-(2*pi-fits_sub[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_sub), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_sub[[i]]$theta_test),rownames(pdata))]
  get.pve(dap[order(fits_sub[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```



---

## PVE of cell times in all samples after removing PC outliers

```{r, eval=F}
rfp_theta_sub <- with(pdata[which(!(rownames(pdata) %in% which_out)),],
                  get.pve(rfp.median.log10sum.adjust[order(theta)]))

gfp_theta_sub <- with(pdata[which(!(rownames(pdata) %in% which_out)),],
                  get.pve(gfp.median.log10sum.adjust[order(theta)]))

dapi_theta_sub <- with(pdata[which(!(rownames(pdata) %in% which_out)),],
                  get.pve(dapi.median.log10sum.adjust[order(theta)]))

save(rfp_theta_sub,
     gfp_theta_sub, 
     dapi_theta_sub, 
     file = "../output/method-labels.Rmd/pve_no_pc_outlier.rda")
```


```{r}
load(file="../output/method-labels.Rmd/pve_no_pc_outlier.rda")
c(rfp_theta_sub,
     gfp_theta_sub, 
     dapi_theta_sub) 
```

---

## Re-fitting after removing PC outliers from the same 5 folds, top 50 genes


```{r, eval=F}
expr_sub <- expr.sig[which(rownames(expr.sig) %in% rownames(sig.genes)[1:50]),]
fits_sub_top50 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train)
  
  Y_train_sub <- Y_train[,which(!(colnames(Y_train) %in% which_out))]
  theta_train_sub <- theta_train[which(!(names(theta_train) %in% which_out))]

  fit.train <- cycle.npreg.insample(Y = Y_train_sub, 
                                    theta = theta_train_sub, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  Y_test_sub <- Y_test[,which(!(colnames(Y_test) %in% which_out))]
  theta_test_sub <- theta_test[which(!(names(theta_test) %in% which_out))]

  fit.test <- cycle.npreg.outsample(Y_test=Y_test_sub,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_sub_top50[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test_sub)
}

for (i in 1:5) {
  fits_sub_top50[[i]]$theta_est_shift <- rotation(fits_sub_top50[[i]]$theta_test, fits_sub_top50[[i]]$fit.test$cell_times_est)$y2shift
}

saveRDS(fits_sub_top50, file = "../output/method-labels.Rmd/fits_sub_top50.rds")
```


```{r}
fits_sub_top50 <- readRDS(file = "../output/method-labels.Rmd/fits_sub_top50.rds")

diff_time <- lapply(1:5, function(i) {
  pmin(abs(fits_sub_top50[[i]]$theta_est_shift-fits_sub_top50[[i]]$theta_test),
    abs(fits_sub_top50[[i]]$theta_est_shift-(2*pi-fits_sub_top50[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_sub_top50), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_sub_top50[[i]]$theta_test),
                                          rownames(pdata))]
  get.pve(dap[order(fits_sub_top50[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```



---

## Re-fitting after removing PC outliers from the same 5 folds, top 10 genes


```{r, eval=F}
expr_sub <- expr.sig[which(rownames(expr.sig) %in% rownames(sig.genes)[1:10]),]
fits_sub_top10 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train)
  
  Y_train_sub <- Y_train[,which(!(colnames(Y_train) %in% which_out))]
  theta_train_sub <- theta_train[which(!(names(theta_train) %in% which_out))]

  fit.train <- cycle.npreg.insample(Y = Y_train_sub, 
                                    theta = theta_train_sub, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  Y_test_sub <- Y_test[,which(!(colnames(Y_test) %in% which_out))]
  theta_test_sub <- theta_test[which(!(names(theta_test) %in% which_out))]

  fit.test <- cycle.npreg.outsample(Y_test=Y_test_sub,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_sub_top10[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test_sub)
}

for (i in 1:5) {
  fits_sub_top10[[i]]$theta_est_shift <- rotation(fits_sub_top10[[i]]$theta_test, fits_sub_top10[[i]]$fit.test$cell_times_est)$y2shift
}

saveRDS(fits_sub_top10, file = "../output/method-labels.Rmd/fits_sub_top10.rds")
```


```{r}
fits_sub_top10 <- readRDS(file = "../output/method-labels.Rmd/fits_sub_top10.rds")

diff_time <- lapply(1:5, function(i) {
  pmin(abs(fits_sub_top10[[i]]$theta_est_shift-fits_sub_top10[[i]]$theta_test),
    abs(fits_sub_top10[[i]]$theta_est_shift-(2*pi-fits_sub_top10[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_sub_top10), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_sub_top10[[i]]$theta_test),
                                          rownames(pdata))]
  get.pve(dap[order(fits_sub_top10[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```



---

## Re-fitting after removing PC outliers, odd number


```{r, eval=F}
expr_sub <- expr.sig[which(rownames(expr.sig) %in% rownames(sig.genes)[seq(1,nrow(expr.sig),by=2)]),]
fits_sub_odd <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train)
  
  Y_train_sub <- Y_train[,which(!(colnames(Y_train) %in% which_out))]
  theta_train_sub <- theta_train[which(!(names(theta_train) %in% which_out))]

  fit.train <- cycle.npreg.insample(Y = Y_train_sub, 
                                    theta = theta_train_sub, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  Y_test_sub <- Y_test[,which(!(colnames(Y_test) %in% which_out))]
  theta_test_sub <- theta_test[which(!(names(theta_test) %in% which_out))]

  fit.test <- cycle.npreg.outsample(Y_test=Y_test_sub,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_sub_odd[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test_sub)
}

for (i in 1:5) {
  fits_sub_odd[[i]]$theta_est_shift <- rotation(fits_sub_odd[[i]]$theta_test, fits_sub_odd[[i]]$fit.test$cell_times_est)$y2shift
}

saveRDS(fits_sub_odd, file = "../output/method-labels.Rmd/fits_sub_odd.rds")
```


```{r}
fits_sub_odd <- readRDS(file = "../output/method-labels.Rmd/fits_sub_odd.rds")

diff_time <- lapply(1:5, function(i) {
  pmin(abs(fits_sub_odd[[i]]$theta_est_shift-fits_sub_odd[[i]]$theta_test),
    abs(fits_sub_odd[[i]]$theta_est_shift-(2*pi-fits_sub_odd[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_sub_odd), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_sub_odd[[i]]$theta_test),
                                          rownames(pdata))]
  get.pve(dap[order(fits_sub_odd[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```




---

## Re-fitting after removing PC outliers, even number


```{r, eval=F}
expr_sub <- expr.sig[which(rownames(expr.sig) %in% rownames(sig.genes)[seq(2,nrow(expr.sig),by=2)]),]
fits_sub_even <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train)
  
  Y_train_sub <- Y_train[,which(!(colnames(Y_train) %in% which_out))]
  theta_train_sub <- theta_train[which(!(names(theta_train) %in% which_out))]

  fit.train <- cycle.npreg.insample(Y = Y_train_sub, 
                                    theta = theta_train_sub, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  Y_test_sub <- Y_test[,which(!(colnames(Y_test) %in% which_out))]
  theta_test_sub <- theta_test[which(!(names(theta_test) %in% which_out))]

  fit.test <- cycle.npreg.outsample(Y_test=Y_test_sub,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_sub_even[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test_sub)
}

for (i in 1:5) {
  fits_sub_even[[i]]$theta_est_shift <- rotation(fits_sub_even[[i]]$theta_test, fits_sub_even[[i]]$fit.test$cell_times_est)$y2shift
}

saveRDS(fits_sub_even, file = "../output/method-labels.Rmd/fits_sub_even.rds")
```


```{r}
fits_sub_even <- readRDS(file = "../output/method-labels.Rmd/fits_sub_even.rds")

diff_time <- lapply(1:5, function(i) {
  pmin(abs(fits_sub_even[[i]]$theta_est_shift-fits_sub_even[[i]]$theta_test),
    abs(fits_sub_even[[i]]$theta_est_shift-(2*pi-fits_sub_even[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_sub_even), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_sub_even[[i]]$theta_test),
                                          rownames(pdata))]
  get.pve(dap[order(fits_sub_even[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```


---

## Fitting using all samples, top 50 genes

```{r, eval =F}
expr_sub <- expr.sig[which(rownames(expr.sig) %in% rownames(sig.genes)[1:50]),]

fits_top50 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 

  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_top50[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test)
}

for (i in 1:5) {
  fits_top50[[i]]$theta_est_shift <- rotation(fits_top50[[i]]$theta_test, fits_top50[[i]]$fit.test$cell_times_est)$y2shift
}
  
saveRDS(fits_top50, file = "../output/method-labels.Rmd/fits_top50.rds")
```


```{r}
fits_top50 <- readRDS(file = "../output/method-labels.Rmd/fits_top50.rds")
source("../peco/R/utility.R")

diff_time <- lapply(1:length(fits_top50), function(i) {
  pmin(abs(fits_top50[[i]]$theta_est_shift-fits_top50[[i]]$theta_test),
    abs(fits_top50[[i]]$theta_est_shift-(2*pi-fits_top50[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_top50), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_top50[[i]]$theta_test),
                                          rownames(pdata))]
  get.pve(dap[order(fits_top50[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```



---

## Fitting using all samples, top 10 genes

```{r, eval =F}
expr_sub <- expr.sig[which(rownames(expr.sig) %in% rownames(sig.genes)[1:10]),]

fits_top10 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 

  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_top10[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test)
}

for (i in 1:5) {
  fits_top10[[i]]$theta_est_shift <- rotation(fits_top10[[i]]$theta_test, fits_top10[[i]]$fit.test$cell_times_est)$y2shift
}
  
saveRDS(fits_top10, file = "../output/method-labels.Rmd/fits_top10.rds")
```


```{r}
fits_top10 <- readRDS(file = "../output/method-labels.Rmd/fits_top10.rds")
source("../peco/R/utility.R")

diff_time <- lapply(1:length(fits_top10), function(i) {
  pmin(abs(fits_top10[[i]]$theta_est_shift-fits_top10[[i]]$theta_test),
    abs(fits_top10[[i]]$theta_est_shift-(2*pi-fits_top10[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_top10), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_top10[[i]]$theta_test),
                                          rownames(pdata))]
  get.pve(dap[order(fits_top10[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```


---

## Fitting using all samples, odd-numbered genes

```{r, eval =F}
expr_sub <- expr.sig[which(rownames(expr.sig) %in% rownames(sig.genes)[seq(1,nrow(expr.sig),by=2)]),]

fits_top50_odd <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 

  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_top50_odd[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test)
}

for (i in 1:5) {
  fits_top50_odd[[i]]$theta_est_shift <- rotation(fits_top50_odd[[i]]$theta_test, fits_top50_odd[[i]]$fit.test$cell_times_est)$y2shift
}
  
saveRDS(fits_top50_odd, file = "../output/method-labels.Rmd/fits_top50_odd.rds")
```


```{r}
fits_top50_odd <- readRDS(file = "../output/method-labels.Rmd/fits_top50_odd.rds")
source("../peco/R/utility.R")

diff_time <- lapply(1:length(fits_top50_odd), function(i) {
  pmin(abs(fits_top50_odd[[i]]$theta_est_shift-fits_top50_odd[[i]]$theta_test),
    abs(fits_top50_odd[[i]]$theta_est_shift-(2*pi-fits_top50_odd[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_top50_odd), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_top50_odd[[i]]$theta_test),
                                          rownames(pdata))]
  get.pve(dap[order(fits_top50_odd[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```


---

## Fitting using all samples, even-numbered genes

```{r, eval =F}
expr_sub <- expr.sig[which(rownames(expr.sig) %in% rownames(sig.genes)[seq(2,nrow(expr.sig),by=2)]),]

fits_top50_even <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 

  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits_top50_even[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test)
}

for (i in 1:5) {
  fits_top50_even[[i]]$theta_est_shift <- rotation(fits_top50_even[[i]]$theta_test, fits_top50_even[[i]]$fit.test$cell_times_est)$y2shift
}
  
saveRDS(fits_top50_even, file = "../output/method-labels.Rmd/fits_top50_even.rds")
```


```{r}
fits_top50_even <- readRDS(file = "../output/method-labels.Rmd/fits_top50_even.rds")
source("../peco/R/utility.R")

diff_time <- lapply(1:length(fits_top50_even), function(i) {
  pmin(abs(fits_top50_even[[i]]$theta_est_shift-fits_top50_even[[i]]$theta_test),
    abs(fits_top50_even[[i]]$theta_est_shift-(2*pi-fits_top50_even[[i]]$theta_test)))
})
mean(sapply(diff_time, mean)/2/pi)

pve <- lapply(1:length(fits_top50_even), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_top50_even[[i]]$theta_test),
                                          rownames(pdata))]
  get.pve(dap[order(fits_top50_even[[i]]$theta_est_shift)])
})
mean(unlist(pve))
```

---


```{r}
sessionInfo()
```



