---
title: "Predicting cell times for individuals"
author: "Joyce Hsiao"
output: 
  html_document:
    toc: TRUE
    toc_float: FALSE
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Pipeline

1. Split data to validation samples and training samples. 

    * The split samples are saved as `data/ind_XX_withheld.rds` and `data/ind_XX_training.rds`.    
    * The code used to generated teh split samples is saved in `code/working/makedata_supervised.R`.

2. In the training sample, repeat the following 5 times. Each time, take samples from 4 individuals, fit trendfilter and identified the top X cyclical genes, and the compute the error in predicting cell time of the held-out individual, as a function of the top X cyclical genes. 

    * `code/run_methods.R`: runs all methods, including supervised, unsupervised trendfilter, unsupervised bspline, unsupervised lowess, seurat) and compute prediction error margin.    
    * `code/working/job_run_methods.train.ind.R` calls `code/run_methods.R`
    * `code/working/job_run_methods.train.ind.sbatch` submits jobs one fold at a time.

3. Based on the training sample results, select the top best X genes and fit to the validation sample. Compute and compare the methods' performance on the validation sample. 


---

## Top X cyclical genes

time based on fucci only

```{r, eval=F}
# observed data
library(gplots)
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
ngenes <- c(5, seq(10, 11040, 10))
for (j in 1:length(inds)) {
  ind <- inds[j]
  gene_names <- rownames(readRDS(paste0("../data/results/ind_",ind,"_data_training_cyclical_genes.fold.",
                   1,".rds")))
  genes_list <- lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    tmp <- do.call(cbind, lapply(1:5, function(fold) {
      fl_name <- paste0("../data/results/ind_",ind,"_data_training_cyclical_genes.fold.",
                   fold,".rds")
      df <- readRDS(fl_name)
      top_list <- rownames(df)[order(df$pve,decreasing = T)[1:ngene]]
      rownames(df) %in% top_list
    }) )
    rownames(tmp) <- gene_names
    return(rownames(tmp)[rowSums(tmp)>=4])
  })
  names(genes_list) <- ngenes
  saveRDS(genes_list, 
          file = paste0("../data/results/ind_",ind,"_results_topgenes.rds"))
}


ind_NA19098_topgenes <- readRDS("../data/results/ind_NA19098_results_topgenes.rds")
ind_NA18511_topgenes <- readRDS("../data/results/ind_NA18511_results_topgenes.rds")
ind_NA18870_topgenes <- readRDS("../data/results/ind_NA18870_results_topgenes.rds")
ind_NA19101_topgenes <- readRDS("../data/results/ind_NA19101_results_topgenes.rds")
ind_NA18855_topgenes <- readRDS("../data/results/ind_NA18855_results_topgenes.rds")
ind_NA19160_topgenes <- readRDS("../data/results/ind_NA19160_results_topgenes.rds")

topgenes_list <- list(ind_NA19098_topgenes, ind_NA18511_topgenes, ind_NA18870_topgenes,
                      ind_NA19101_topgenes, ind_NA18855_topgenes, ind_NA19160_topgenes)
library(dplyr)
topgenes_combined <- lapply(1:length(ind_NA19160_topgenes), function(i) {
  ll <- lapply(topgenes_list, "[[", i)
  ll <- unlist(ll)  
  tab <- table(ll)
  names(tab)[which(tab>=4)]
})
names(topgenes_combined) <- names(ind_NA19160_topgenes)

saveRDS(topgenes_combined, file="../data/results/ind_results_topgenes.rds")
```


time based on fucci and dapi only

```{r, eval=F}
# observed data
library(gplots)
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
ngenes <- c(5, seq(10, 11040, 10))
for (j in 1:length(inds)) {
  ind <- inds[j]
  gene_names <- rownames(readRDS(paste0("../data/results/triple_ind_",ind,
                                        "_data_training_cyclical_genes.fold.",
                   1,".rds")))
  genes_list <- lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    tmp <- do.call(cbind, lapply(1:5, function(fold) {
      fl_name <- paste0("../data/results/triple_ind_",ind,
                        "_data_training_cyclical_genes.fold.",
                   fold,".rds")
      df <- readRDS(fl_name)
      top_list <- rownames(df)[order(df$pve,decreasing = T)[1:ngene]]
      rownames(df) %in% top_list
    }) )
    rownames(tmp) <- gene_names
    return(rownames(tmp)[rowSums(tmp)>=4])
  })
  names(genes_list) <- ngenes
  saveRDS(genes_list, 
          file = paste0("../data/results/ind_",ind,"_results_topgenes_triple.rds"))
}


ind_NA19098_topgenes <- readRDS("../data/results/ind_NA19098_results_topgenes_triple.rds")
ind_NA18511_topgenes <- readRDS("../data/results/ind_NA18511_results_topgenes_triple.rds")
ind_NA18870_topgenes <- readRDS("../data/results/ind_NA18870_results_topgenes_triple.rds")
ind_NA19101_topgenes <- readRDS("../data/results/ind_NA19101_results_topgenes_triple.rds")
ind_NA18855_topgenes <- readRDS("../data/results/ind_NA18855_results_topgenes_triple.rds")
ind_NA19160_topgenes <- readRDS("../data/results/ind_NA19160_results_topgenes_triple.rds")

topgenes_list <- list(ind_NA19098_topgenes, ind_NA18511_topgenes, ind_NA18870_topgenes,
                      ind_NA19101_topgenes, ind_NA18855_topgenes, ind_NA19160_topgenes)
library(dplyr)
topgenes_combined <- lapply(1:length(ind_NA19160_topgenes), function(i) {
  ll <- lapply(topgenes_list, "[[", i)
  ll <- unlist(ll)  
  tab <- table(ll)
  names(tab)[which(tab>=4)]
})
names(topgenes_combined) <- names(ind_NA19160_topgenes)

saveRDS(topgenes_combined, file="../data/results/ind_results_topgenes_triple.rds")
```

Compare number of genes selected at Top X cyclical genes. 

```{r}
topgenes_double <- readRDS(file="../data/results/ind_results_topgenes.rds")
topgenes_triple <- readRDS(file="../data/results/ind_results_topgenes_triple.rds")

topgenes_overlap <- sapply(1:length(topgenes_double), function(i) {
 intersect(topgenes_double[[i]], topgenes_triple[[i]]) 
})
names(topgenes_overlap) <- names(topgenes_double)

plot(names(topgenes_double)[1:21],
     sapply(topgenes_double, length)[1:21],
     xlab="Top X cyclical genes", pch=16, col=1, cex=.7,
     ylim=c(0,60), 
     ylab="Selected genes", 
     main = "Number of selected genes")
points(names(topgenes_double)[1:21],
     sapply(topgenes_triple, length)[1:21], col="blue", pch=16, cex=.7)
points(names(topgenes_double)[1:21],
     sapply(topgenes_overlap, length)[1:21], col="1", pch=1, cex=.7)
legend("topleft", legend=c("double", "triple", "both"),
       col=c("black", "blue", "black"), pch=c(16, 16, 1))
```


---

## Prediction error

summarize results

```{r}
diff_time_wrapper <- function(results_list) {
  
  methods_list <- sapply(names(results_list),
                         function(x) strsplit(x, split=".", fixed=TRUE)[[1]][2])
  
  diff_time_list <- do.call(rbind, lapply(1:length(results_list), function(i) {
    diff_time <- results_list[[i]]$diff_time
    diff_mean <- mean(diff_time/2/pi)
#    diff_se <- sd(diff_time/2/pi)/sqrt(ncol(results_list[[1]]$Y))

    return(data.frame(diff_mean=diff_mean,
#                      diff_se=diff_se,
                      methods=methods_list[i]))
  }) )
  return(diff_time_list)  
}



library(gplots)
ngenes <- c(5, seq(10,200, by=10))
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
train_top <- lapply(1:length(inds), function(j) {
  ind <- inds[j]
  out <- do.call(rbind, lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    train_topX <- do.call(rbind, lapply(1:5, function(fold) {
      # print(ind)
      # print(ngene)
      # print(fold)
      fl_name <- paste0("../data/results/ind_",ind,"_results_train.fold.",fold,
                        ".top",ngene,".rds")
      df <- readRDS(fl_name)
      out <- diff_time_wrapper(df$fit.test)
      out$fold <- fold
      return(out)
    }) )
    train_topX$ngenes <- ngene
    #return(train_topX)  
    agg_mn <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=mean)
    agg_sd <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=sd)
  
    obj <- data.frame(methods=agg_mn$methods, 
                      diff_mean=agg_mn$diff_mean,
                      diff_se=agg_sd$diff_mean/sqrt(5))
    obj$ngenes <- ngene
    return(obj)
  }) )
  out$ind <- ind
  return(out)
})
names(train_top) <- inds
saveRDS(train_top, file = "../output/method-train-ind.Rmd/diff_time_all.rds")



ngenes <- c(5, seq(10,200, by=10))
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
train_triple <- lapply(1:length(inds), function(j) {
  ind <- inds[j]
  out <- do.call(rbind, lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    train_topX <- do.call(rbind, lapply(1:5, function(fold) {
      # print(ind)
      # print(ngene)
      # print(fold)
      fl_name <- paste0("../data/results/triple_ind_",ind,"_results_train.fold.",fold,
                        ".top",ngene,".rds")
      df <- readRDS(fl_name)
      out <- diff_time_wrapper(df$fit.test)
      out$fold <- fold
      return(out)
    }) )
    train_topX$ngenes <- ngene
    #return(train_topX)  
    agg_mn <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=mean)
    agg_sd <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=sd)
  
    obj <- data.frame(methods=agg_mn$methods, 
                      diff_mean=agg_mn$diff_mean,
                      diff_se=agg_sd$diff_mean/sqrt(5))
    obj$ngenes <- ngene
    return(obj)
  }) )
  out$ind <- ind
  return(out)
})
names(train_triple) <- inds
saveRDS(train_triple, file = "../output/method-train-ind.Rmd/diff_time_all_triple.rds")


# ngenes <- c(5, seq(10,200, by=10))
# train_top_permute <- do.call(rbind, lapply(1:length(ngenes), function(i) {
#   ngene <- ngenes[i]
#   train_topX <- do.call(rbind, lapply(1:5, function(fold) {
#     fl_name <- paste0("../data/results/results_train_permute_oncyclical.fold.",
#                       fold,".top",ngene,".rds")
#     df <- readRDS(fl_name)
#     out <- diff_time_wrapper(df$fit.test)
#     out$fold <- fold
#     return(out)
#   }) )
#   train_topX$ngenes <- ngene
#   #return(train_topX)  
#   agg_mn <- aggregate(diff_mean ~ methods,
#                     data=train_topX, FUN=mean)
#   agg_sd <- aggregate(diff_mean ~ methods,
#                     data=train_topX, FUN=sd)
# 
#   obj <- data.frame(methods=agg_mn$methods, 
#                     diff_mean=agg_mn$diff_mean,
#                     diff_se=agg_sd$diff_mean/sqrt(5))
#   obj$ngenes <- ngene
#   return(obj)
# }) )
# 
# saveRDS(train_top_permute, 
#         file = "../output/method-train-classifiers-all.Rmd/diff_time_fucci_on_rando.rds")

```

Results

```{r plot-results}
train_double <- readRDS("../output/method-train-ind.Rmd/diff_time_all.rds")
train_double <- do.call(rbind, train_double)
train_double <- subset(train_double, methods=="supervised")
train_double$ind <- as.factor(train_double$ind)

train_triple <- readRDS("../output/method-train-ind.Rmd/diff_time_all_triple.rds")
train_triple <- do.call(rbind, train_triple)
train_triple <- subset(train_triple, methods=="supervised")
train_triple$ind <- as.factor(train_triple$ind)
```

time based on fucci

```{r}
library(ggplot2)
ggplot(train_double, 
       aes(x=ngenes, y=diff_mean, col=ind)) +
  # geom_errorbar(train_top, 
  #               mapping=aes(ymin=diff_mean-diff_se, ymax=diff_mean+diff_se,
  #                           col=ind)) +
  geom_point() + geom_line(lty=3) +
  ylab("Error margin (% arc length)") + xlab("Top X cyclical genes") +
  ggtitle("Prediction error in the training set") +
  ylim(0,.25)

ggplot(train_double, 
       aes(x=ngenes, y=diff_mean, col=ind)) +
  # geom_errorbar(train_top, 
  #               mapping=aes(ymin=diff_mean-diff_se, ymax=diff_mean+diff_se,
  #                           col=ind)) +
  geom_point() + geom_line(lty=3) +
  ylab("Error margin (% arc length)") + xlab("Top X cyclical genes") +
  ggtitle("Prediction error in the training set") 


ggplot(train_double, 
       aes(x=ngenes, y=diff_mean, col=ind)) +
  geom_errorbar(train_double, 
                mapping=aes(ymin=diff_mean-diff_se, ymax=diff_mean+diff_se,
                            col=ind)) +
  geom_point() + geom_line(lty=3) +
  ylab("Error margin (% arc length)") + xlab("Top X cyclical genes") +
  ggtitle("Prediction error in the training set") 
```


time based on fucci + dapi

```{r}
library(ggplot2)
ggplot(train_triple, 
       aes(x=ngenes, y=diff_mean, col=ind)) +
  # geom_errorbar(train_top, 
  #               mapping=aes(ymin=diff_mean-diff_se, ymax=diff_mean+diff_se,
  #                           col=ind)) +
  geom_point() + geom_line(lty=3) +
  ylab("Error margin (% arc length)") + xlab("Top X cyclical genes") +
  ggtitle("Prediction error in the training set") +
  ylim(0,.25)

library(ggplot2)
ggplot(train_triple, 
       aes(x=ngenes, y=diff_mean, col=ind)) +
  # geom_errorbar(train_top, 
  #               mapping=aes(ymin=diff_mean-diff_se, ymax=diff_mean+diff_se,
  #                           col=ind)) +
  geom_point() + geom_line(lty=3) +
  ylab("Error margin (% arc length)") + xlab("Top X cyclical genes") +
  ggtitle("Prediction error in the training set") 

ggplot(train_triple, 
       aes(x=ngenes, y=diff_mean, col=ind)) +
  geom_errorbar(train_double, 
                mapping=aes(ymin=diff_mean-diff_se, ymax=diff_mean+diff_se,
                            col=ind)) +
  geom_point() + geom_line(lty=3) +
  ylab("Error margin (% arc length)") + xlab("Top X cyclical genes") +
  ggtitle("Prediction error in the training set") 
```




---

## Session information

```{r}
sessionInfo()
```

