---
title: "Evaluate results with/without noisy training labels, equal-sized folds"
author: "Joyce Hsiao"
output: 
  html_document:
    toc: TRUE
    toc_float: FALSE
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Summary

1. Here I remove the noisy labels before partitioning samples. 

2. Compared to results when partitioning the sample before removing the noisy labels, we have a better improvement in prediction margin of error and in DAPI PVE. See those results [here](docs/method-training-labels.html). 




---

## Define PC outliers

```{r}
library(Biobase)
source("../peco/R/cycle.corr.R")
source("../peco/R/cycle.npreg.R")

df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

# select external validation samples
log2cpm.quant <- readRDS("../output/npreg-trendfilter-quantile.Rmd/log2cpm.quant.rds")


# first check the theta in pdata
pca <- prcomp(cbind(pdata$rfp.median.log10sum.adjust,
                    pdata$gfp.median.log10sum.adjust), scale=TRUE)
pca_df <- cbind(pca$x[,1],pca$x[,2])
rownames(pca_df) <- rownames(pdata)

library(circular)
theta_check <- as.numeric(coord2rad(pca_df))
theta_check <- 2*pi-theta_check
#plot(theta_check, pdata$theta)
names(theta_check) <- rownames(pdata)

dist_to_origin <- sqrt(pca_df[,1]^2+pca_df[,2]^2)
which_out <- rownames(pdata)[which(scale(dist_to_origin) < -1)]


log2cpm.quant.filt <- log2cpm.quant[, !(colnames(log2cpm.quant) %in% which_out)]

#intersect(colnames(log2cpm.quant.filt), which_out)

set.seed(99)
nvalid <- round(ncol(log2cpm.quant.filt)*.15)
ii.valid <- sample(1:ncol(log2cpm.quant.filt), nvalid, replace = F)
ii.nonvalid <- setdiff(1:ncol(log2cpm.quant.filt), ii.valid)

log2cpm.quant.filt.nonvalid <- log2cpm.quant.filt[,ii.nonvalid]
log2cpm.quant.filt.valid <- log2cpm.quant.filt[,ii.valid]

sig.genes <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.101.rds")

# get predicted times
# set training samples
source("../peco/R/primes.R")
source("../peco/R/partitionSamples.R")
parts <- partitionSamples(1:ncol(log2cpm.quant.filt.nonvalid), runs=5,
                          nsize.each = rep(133,5))
part_indices <- parts$partitions
```


---

## Fitting excluding PC outliers 

Top 101 genes

```{r, eval=F}
source("../peco/R/fit.trendfilter.generic.R")
source("../peco/R/cycle.npreg.R")
source("../peco/R/utility.R")

expr_sub <- log2cpm.quant.filt.nonvalid[rownames(log2cpm.quant.filt.nonvalid) %in% rownames(sig.genes), ]

fits_top101 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 
  
  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=12,
                                    polyorder=2,
                                    method.trend="trendfilter")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="trendfilter",
                                    ncores=12)
  
  fits_top101[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test)
}

for (i in 1:5) {
  fits_top101[[i]]$theta_est_shift <- rotation(fits_top101[[i]]$theta_test, fits_top101[[i]]$fit.test$cell_times_est)$y2shift
}

saveRDS(fits_top101, file = "../output/method-train-labels-equalfold.Rmd/fits_top101.rds")
#fits_top101 <- readRDS("../output/method-train-labels-equalfold.Rmd/fits_top101.rds")

diff_time <- lapply(1:5, function(i) {
  pmin(abs(fits_top101[[i]]$theta_est_shift-fits_top101[[i]]$theta_test),
    abs(fits_top101[[i]]$theta_est_shift-(2*pi-fits_top101[[i]]$theta_test)))
})

pve <- lapply(1:length(fits_top101), function(i) {
  dap <- pdata$dapi.median.log10sum.adjust[match(names(fits_top101[[i]]$theta_test),
                                                 rownames(pdata))]
  get.pve(dap[order(fits_top101[[i]]$theta_est_shift)])
})

save(diff_time, pve, 
     file = "../output/method-train-labels-equalfold.Rmd/modelresults_top101.rda")
# ```
# 
# 
# 
# Top 10 genes
# 
# ```{r, eval=F}
expr_sub <- log2cpm.quant.filt.nonvalid[rownames(log2cpm.quant.filt.nonvalid) %in% rownames(sig.genes)[1:10], ]

fits_top10 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr_sub[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 
  
  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=12,
                                    polyorder=2,
                                    method.trend="trendfilter")
  # fitting test data
  Y_test <- expr_sub[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="trendfilter",
                                    ncores=12)
  
  fits_top10[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test)
}

for (i in 1:5) {
  fits_top10[[i]]$theta_est_shift <- rotation(fits_top10[[i]]$theta_test, fits_top10[[i]]$fit.test$cell_times_est)$y2shift
}

saveRDS(fits_top10, file = "../output/method-train-labels-equalfold.Rmd/fits_top10.rds")
#fits_top10 <- readRDS("../output/method-train-labels-equalfold.Rmd/fits_top10.rds")


diff_time <- lapply(1:5, function(i) {
  pmin(abs(fits_top10[[i]]$theta_est_shift-fits_top10[[i]]$theta_test),
    abs(fits_top10[[i]]$theta_est_shift-(2*pi-fits_top10[[i]]$theta_test)))
})

pve <- lapply(1:length(fits_top10), function(i) {
  dap <- pdata$dapi.median.log10sum.adjust[match(names(fits_top10[[i]]$theta_test),
                                                 rownames(pdata))]
  get.pve(dap[order(fits_top10[[i]]$theta_est_shift)])
})

save(diff_time, pve, 
     file = "../output/method-train-labels-equalfold.Rmd/modelresults_top10.rda")
```

---

## Results


```{r}
load(file = "../output/method-train-labels-equalfold.Rmd/modelresults_top101.rda")
mean(sapply(diff_time, mean)/2/pi)
mean(sapply(pve, "[[", 1))

load(file = "../output/method-train-labels-equalfold.Rmd/modelresults_top10.rda")
mean(sapply(diff_time, mean)/2/pi)
mean(unlist(pve))
```

---

## fitting top 10 genes without PC outliers

```{r, eval=F}
fits_sub_top10 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.sig[,part_indices[[run]]$train]
  theta_train <- theta_check[match(colnames(Y_train), rownames(pdata))]
  names(theta_train) <- colnames(Y_train) 
  
  Y_train_sub <- Y_train[,which(!(colnames(Y_train) %in% which_out))]
  theta_train_sub <- theta_train[which(!(names(theta_train) %in% which_out))]

  fit.train <- cycle.npreg.insample(Y = Y_train_sub, 
                                    theta = theta_train_sub, 
                                    ncores=12,
                                    polyorder=2,
                                    method.trend="trendfilter")
  # fitting test data
  Y_test <- expr.sig[,part_indices[[run]]$test]
  theta_test <- theta_check[match(colnames(Y_test), rownames(pdata))]
  names(theta_test) <- colnames(Y_test) 

  Y_test_sub <- Y_test[,which(!(colnames(Y_test) %in% which_out))]
  theta_test_sub <- theta_test[which(!(names(theta_test) %in% which_out))]

  fit.test <- cycle.npreg.outsample(Y_test=Y_test_sub,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="trendfilter",
                                    ncores=12)
  
  fits_sub_top10[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test,
                      theta_test=theta_test_sub)
}

for (i in 1:5) {
  fits_sub_top10[[i]]$theta_est_shift <- rotation(fits_sub_top10[[i]]$theta_test, fits_sub_top10[[i]]$fit.test$cell_times_est)$y2shift
}

saveRDS(fits_sub_top10, file = "../output/method-train-labels.Rmd/fits_sub_top10.rds")


diff_time <- lapply(1:5, function(i) {
  pmin(abs(fits_sub_top10[[i]]$theta_est_shift-fits_sub_top10[[i]]$theta_test),
    abs(fits_sub_top10[[i]]$theta_est_shift-(2*pi-fits_sub_top10[[i]]$theta_test)))
})

pve <- lapply(1:length(fits_sub), function(i) {
  dap <- pdata$dapi.median.log10sum[match(names(fits_sub_top10[[i]]$theta_test),rownames(pdata))]
  get.pve(dap[order(fits_sub_top10[[i]]$theta_est_shift)])
})

save(diff_time, pve,
     file="../output/method-train-labels.Rmd/modelresults_excludeoutlier_top10.rda")
```


```{r}
load(file="../output/method-train-labels.Rmd/modelresults_excludeoutlier_top10.rda")
mean(sapply(diff_time, mean)/2/pi)
mean(unlist(pve))
```




---

## Session information

```{r}
sessionInfo()
```



