---
title: "Evaluate the withheld sample"
author: "Joyce Hsiao"
output: 
  html_document:
    toc: TRUE
    toc_float: FALSE
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Goals

1. Compute seurat cell time using seurat S score and G2M score. Take the arctangent transformation of these two scores. 

2. Evaluate the association between Seurat classes and cell times with DAPI, GFP and RFP. 

3. Consider all of the 888 samples included in the analysis. 

---

## Data

```{r}
library(Biobase)
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

#macosko <- readRDS("data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")
counts <- counts[,order(pdata$theta)]
log2cpm.all <- log2cpm.all[,order(pdata$theta)]
pdata <- pdata[order(pdata$theta),]

sig.genes <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.101.rds")
#expr.sig <- log2cpm.quant.valid[rownames(log2cpm.quant.valid) %in% rownames(sig.genes), ]

library(biomaRt)
ensembl <- useMart("ensembl",dataset="hsapiens_gene_ensembl")
symbols <- getBM(attributes=c('ensembl_gene_id','hgnc_symbol'), 
      filter='ensembl_gene_id', 
      values = rownames(log2cpm.all), mart = ensembl)
head(symbols)
```

Get seurat genes 

```{r}
cc.genes <- readLines(con = "../data/cellcycle-genes-previous-studies/seurat_cellcycle/regev_lab_cell_cycle_genes.txt")
s.genes <- cc.genes[1:43]
g2m.genes <- cc.genes[44:97]
```

overlap between seurat genes and top 101 cyclical genes

```{r}
library(VennDiagram)
library(grid)
library(ggplot2)

grid.draw(venn.diagram(
  list(cyclical=symbols$hgnc_symbol[which(symbols$ensembl_gene_id %in% rownames(sig.genes))],
       seurat=cc.genes), file=NULL))

df <- rbind(data.frame(pve=c(sig.genes$pve), method="top 100 cyclical"),
            data.frame(pve=c(sig.genes$pve[rownames(sig.genes) %in% symbols$ensembl_gene_id[which(symbols$hgnc_symbol %in% cc.genes)]]), method="overlap"))

ggplot(df, aes(x=method, y=pve)) +
  geom_violin() +
  geom_boxplot(width=.3) +
  ylab("Proportion of variance explained")
```


---

## Seurat

```{r, eval=F}
source("../peco/R/run_seurat.R")

Y_mat <- log2cpm.all
Y_mat <- Y_mat[rownames(Y_mat) %in% symbols$ensembl_gene_id,]
all.equal(rownames(Y_mat), symbols$ensembl_gene_id)
rownames(Y_mat) <- symbols$hgnc_symbol
  
fit.seurat <- run_seurat(Y=Y_mat, 
             s.genes=s.genes, 
             g2m.genes=g2m.genes, 
             n.bin=25,
             seed.use=1, random.seed=1)
  
table(fit.seurat$assignments)

all.equal(names(theta), rownames(fit.seurat))

seurat.pca <- prcomp(cbind(fit.seurat$G2M, fit.seurat$S), scale=TRUE)
fit.seurat$pred_time <- as.numeric(coord2rad(cbind(seurat.pca$x[,1],seurat.pca$x[,2])))
fit.seurat$ref_time <- theta
fit.seurat$pred_time_shift <- with(fit.seurat, rotation(ref_time, pred_time)$y2shift)
fit.seurat$diff_time <- with(fit.seurat,
          pmin(abs(pred_time_shift-ref_time),
               abs(pred_time_shift-(2*pi-ref_time))))

all.equal(rownames(pdata), rownames(fit.seurat))

fit.seurat$dapi <- pdata$dapi.median.log10sum.adjust
fit.seurat$gfp <- pdata$gfp.median.log10sum.adjust
fit.seurat$rfp <- pdata$rfp.median.log10sum.adjust

saveRDS(fit.seurat,
        file="../output/method-labels-seurat.Rmd/fit.seurat.rds")

source("../peco/R/run_seurat.R")
source("../peco/R/utility.R")
seurat.pve <- rbind(
        data.frame(method="seurat.dapi", 
                   with(fit.seurat, get.pve(dapi[order(pred_time_shift)]))),
        data.frame(method="seurat.gfp", 
                   with(fit.seurat, get.pve(gfp[order(pred_time_shift)]))),
        data.frame(method="seurat.rfp", 
                   with(fit.seurat, get.pve(rfp[order(pred_time_shift)]))),
        data.frame(method="seurat.S", 
                   with(fit.seurat, get.pve(S[order(pred_time_shift)]))),
        data.frame(method="seurat.G2M", 
                   with(fit.seurat, get.pve(G2M[order(pred_time_shift)]))),
        data.frame(method="ref.S", 
                   with(fit.seurat, get.pve(S[order(ref_time)]))),
        data.frame(method="ref.G2M", 
                   with(fit.seurat, get.pve(G2M[order(ref_time)])) ),
        data.frame(method="ref.dapi", 
                   with(fit.seurat, get.pve(dapi[order(ref_time)])) ),
        data.frame(method="ref.gfp", 
                   with(fit.seurat, get.pve(gfp[order(ref_time)])) ),
        data.frame(method="ref.rfp", 
                   with(fit.seurat, get.pve(rfp[order(ref_time)])) ),
        data.frame(method="seurat.class.S", 
                   with(fit.seurat, get.aov(S, assignments))),
        data.frame(method="seurat.class.G2M", 
                   with(fit.seurat, get.aov(G2M, assignments))),
        data.frame(method="seurat.class.dapi", 
                   with(fit.seurat, get.aov(dapi, assignments))),
        data.frame(method="seurat.class.gfp", 
                   with(fit.seurat, get.aov(gfp, assignments))),
        data.frame(method="seurat.class.rfp", 
                   with(fit.seurat, get.aov(rfp, assignments))) )

saveRDS(seurat.pve,
     file="../output/method-labels-seurat.Rmd/seurat.pve.rds")
```


```{r}
fit.seurat <- readRDS("../output/method-labels-seurat.Rmd/fit.seurat.rds")
seurat.pve <- readRDS("../output/method-labels-seurat.Rmd/seurat.pve.rds")


print(seurat.pve)

library(ggplot2)

cols <- c("orange", "red", "brown")

ggplot(fit.seurat, aes(x=assignments,
                       y=pred_time_shift)) +
  geom_violin(aes(fill=assignments), lwd=.7) +
  geom_boxplot(width=.2) +
  ylab("Seurat cell time") + xlab("cell-cycle class") +
  ggtitle("Seurat classification") +
  scale_fill_manual(values=cols)

ggplot(fit.seurat, aes(x=assignments)) +
  geom_bar(aes(fill=assignments)) +
  xlab("cell-cycle class") +
  ggtitle("Seurat classification") +
  ylab("Frequency count") +
  scale_fill_manual(values=cols)

ggplot(fit.seurat, aes(x=S,y=G2M, group=assignments)) +
  geom_point(aes(col=assignments)) +
  geom_abline(intercept=0, slope=1, lty=3) +
  geom_hline(yintercept=0, lty=3) +
  geom_vline(xintercept=0, lty=3) + 
  xlab("S score") + ylab("G2M score") +
  scale_color_manual(values=cols)
```





---

## Session information

```{r}
sessionInfo()
```


