---
title: "Predicting cell times derived under unequal PC variances"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Goals

1. Use cell times derived from fucci and dapi to predict cyclical gene expression patterns

2. Here we consider that the PC1 and PC2 of fucci and dapi are not of equal variance. We fit
 an ellipse to the PC1 and PC2 space and project the PC1 and PC2 of each cell to the ellipse. 
 
3. Results showed that the prediction error is higher than when we predicting cyclical expression patterns just using fucci. However, the top detected cyclical genes are also enriched with cell cycle processes, similarly as the top detected cyclical genes based on cell times derived from fucci only. 

4. These results will be included in the supplemental information. 

---

## Top X cyclical genese

```{r, eval=F}
library(gplots)
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
ngenes <- c(1:9, seq(10, 11040, 10))
gene_names <- rownames(readRDS("../data/results/alge_data_training_cyclical_genes.fold.1.rds"))
genes_list <- lapply(1:length(ngenes), function(i) {
  ngene <- ngenes[i]
  tmp <- do.call(cbind, lapply(1:5, function(fold) {
      fl_name <- paste0("../data/results/alge_data_training_cyclical_genes.fold.",
                   fold,".rds")
      df <- readRDS(fl_name)
      top_list <- rownames(df)[order(df$pve,decreasing = T)[1:ngene]]
      rownames(df) %in% top_list
  }) )
  rownames(tmp) <- gene_names
  return(rownames(tmp)[rowSums(tmp)>=4])
})
names(genes_list) <- ngenes
saveRDS(genes_list, 
        file = paste0("../data/results/alge_results_topgenes.rds"))


# mixed individuals
library(gplots)
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
ngenes <- c(1:9, seq(10, 11040, 10))
for (j in 1:length(inds)) {
  ind <- inds[j]
  gene_names <- rownames(readRDS(paste0("../data/results/alge_ind_",ind,"_data_training_cyclical_genes.fold.", 1,".rds")))
  genes_list <- lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    tmp <- do.call(cbind, lapply(1:5, function(fold) {
      fl_name <- paste0("../data/results/alge_ind_",ind,"_data_training_cyclical_genes.fold.",
                   fold,".rds")
      df <- readRDS(fl_name)
      top_list <- rownames(df)[order(df$pve,decreasing = T)[1:ngene]]
      rownames(df) %in% top_list
    }) )
    rownames(tmp) <- gene_names
    return(rownames(tmp)[rowSums(tmp)>=4])
  })
  names(genes_list) <- ngenes
  saveRDS(genes_list, 
          file = paste0("../data/results/alge_ind_",ind,"_results_topgenes.rds"))
}


ind_NA19098_topgenes <- readRDS("../data/results/alge_ind_NA19098_results_topgenes.rds")
ind_NA18511_topgenes <- readRDS("../data/results/alge_ind_NA18511_results_topgenes.rds")
ind_NA18870_topgenes <- readRDS("../data/results/alge_ind_NA18870_results_topgenes.rds")
ind_NA19101_topgenes <- readRDS("../data/results/alge_ind_NA19101_results_topgenes.rds")
ind_NA18855_topgenes <- readRDS("../data/results/alge_ind_NA18855_results_topgenes.rds")
ind_NA19160_topgenes <- readRDS("../data/results/alge_ind_NA19160_results_topgenes.rds")

topgenes_list <- list(ind_NA19098_topgenes, ind_NA18511_topgenes, ind_NA18870_topgenes,
                      ind_NA19101_topgenes, ind_NA18855_topgenes, ind_NA19160_topgenes)
library(dplyr)
topgenes_combined <- lapply(1:length(ind_NA19160_topgenes), function(i) {
  ll <- lapply(topgenes_list, "[[", i)
  ll <- unlist(ll)  
  tab <- table(ll)
  names(tab)[which(tab>=4)]
})
names(topgenes_combined) <- names(ind_NA19160_topgenes)

saveRDS(topgenes_combined, file="../data/results/alge_ind_results_topgenes.rds")
```


---

## Prediction error

summarize results

```{r}
diff_time_wrapper <- function(results_list) {
  
  methods_list <- sapply(names(results_list),
                         function(x) strsplit(x, split=".", fixed=TRUE)[[1]][2])
  
  diff_time_list <- do.call(rbind, lapply(1:length(results_list), function(i) {
    diff_time <- results_list[[i]]$diff_time
    diff_mean <- mean(diff_time/2/pi)
#    diff_se <- sd(diff_time/2/pi)/sqrt(ncol(results_list[[1]]$Y))

    return(data.frame(diff_mean=diff_mean,
#                      diff_se=diff_se,
                      methods=methods_list[i]))
  }) )
  return(diff_time_list)  
}




library(gplots)
ngenes <- c(2:9, seq(10,200, by=10))
train_randos <- do.call(rbind, lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    train_topX <- do.call(rbind, lapply(1:5, function(fold) {
    # print(ind)
    print(ngene)
    print(fold)
    fl_name <- paste0("../data/results/triple_alge_results_train.fold.",fold,
                      ".top",ngene,".rds")
    df <- readRDS(fl_name)
    out <- diff_time_wrapper(df$fit.test)
    out$fold <- fold
    return(out)
    }) )
    train_topX$ngenes <- ngene
    #return(train_topX)  
    agg_mn <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=mean)
    agg_sd <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=sd)
  
    obj <- data.frame(methods=agg_mn$methods, 
                      diff_mean=agg_mn$diff_mean,
                      diff_se=agg_sd$diff_mean/sqrt(5))
    obj$ngenes <- ngene
    return(obj)
  }) )
saveRDS(train_randos, file = "../output/method-train-alge.Rmd/diff_time_all.rds")




# prediction for one individual
ngenes <- c(2:9, seq(10,200, by=10))
inds <- c("NA19098", "NA18511", "NA18870", "NA19101", "NA18855", "NA19160")
train_top <- lapply(1:length(inds), function(j) {
  ind <- inds[j]
  out <- do.call(rbind, lapply(1:length(ngenes), function(i) {
    ngene <- ngenes[i]
    train_topX <- do.call(rbind, lapply(1:5, function(fold) {
      print(ind)
      print(ngene)
      print(fold)
      fl_name <- paste0("../data/results/triple_alge_ind_",ind,"_results_train.fold.",fold,
                        ".top",ngene,".rds")
      df <- readRDS(fl_name)
      out <- diff_time_wrapper(df$fit.test)
      out$fold <- fold
      return(out)
    }) )
    train_topX$ngenes <- ngene
    #return(train_topX)  
    agg_mn <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=mean)
    agg_sd <- aggregate(diff_mean ~ methods,
                      data=train_topX, FUN=sd)
  
    obj <- data.frame(methods=agg_mn$methods, 
                      diff_mean=agg_mn$diff_mean,
                      diff_se=agg_sd$diff_mean/sqrt(5))
    obj$ngenes <- ngene
    return(obj)
  }) )
  out$ind <- ind
  return(out)
})
names(train_top) <- inds
saveRDS(train_top, file = "../output/method-train-alge.Rmd/diff_time_all_indprediction.rds")
```


print results

```{r}
train_randos <- readRDS("../output/method-train-alge.Rmd/diff_time_all.rds")
train_randos$methods <- as.character(train_randos$methods)
train_randos <- subset(train_randos, methods=="supervised")

train_ind <- readRDS("../output/method-train-alge.Rmd/diff_time_all_indprediction.rds")
train_ind <- do.call(rbind, train_ind)
train_ind$methods <- as.character(train_ind$methods)
train_ind <- subset(train_ind, methods=="supervised")
train_ind$ind <- as.factor(train_ind$ind)


# head(a$fit.test$fit.supervised$cell_times_est)
# head(b$fit.test$fit.supervised$cell_times_est)
# 
# a=readRDS("../data/results/triple_alge_ind_NA18511_data_training_cyclical_genes.fold.1.rds")
# b=readRDS("../data/results/triple_ind_NA18511_data_training_cyclical_genes.fold.1.rds")
# 
# head(a$pve)
# head(b$pve)
# 
# plot(a$pve,b$pve)
# 
# a <- readRDS("../data/results/triple_data_training_alge.rds")
# b <- readRDS("../data/results/triple_data_training.rds")
# plot(a$theta.nonvalid,
#      b$theta.nonvalid)
```



```{r}
library(ggplot2)
ggplot(train_randos, 
       aes(x=ngenes, y=diff_mean)) +
  geom_errorbar(train_randos,
                mapping=aes(ymin=diff_mean-diff_se, ymax=diff_mean+diff_se)) +
  geom_point() + geom_line(lty=3) +
  ylab("Error margin (% arc length)") + xlab("Top X cyclical genes") +
  ggtitle("Predict cell times from mixed individuals") 

ggplot(train_ind, 
       aes(x=ngenes, y=diff_mean, col=ind)) +
  geom_errorbar(train_ind,
                mapping=aes(ymin=diff_mean-diff_se, ymax=diff_mean+diff_se,
                            col=ind)) +
  geom_point() + geom_line(lty=3) +
  ylab("Error margin (% arc length)") + xlab("Top X cyclical genes") +
  ggtitle("Predict cell times for one individual") 
```

---

## Output genes

```{r, eval=F}
topgenes_randos <- readRDS("../data/results/alge_results_topgenes.rds")
topgenes_ind <- readRDS("../data/results/alge_ind_results_topgenes.rds")

write.table(topgenes_randos[[21]], 
            file = "../output/method-train-alge.Rmd/gene_list.txt",
            row.names = F, quote=F, col.names = F)

write.table(topgenes_ind[[21]], 
            file = "../output/method-train-alge.Rmd/gene_list_ind.txt",
            row.names = F, quote=F, col.names = F)
```


---


```{r}
sessionInfo()
```


