---
title: "Model training: number of genes and cyclical function fit"
author: "Joyce Hsiao"
output: 
  html_document:
    toc: TRUE
    toc_float: FALSE
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---


## Compare npcirc.nw vs trendfilter

Extract data from the top 101 genes identified 

```{r}
library(Biobase)
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

#macosko <- readRDS("data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")
counts <- counts[,order(pdata$theta)]
log2cpm.all <- log2cpm.all[,order(pdata$theta)]
pdata <- pdata[order(pdata$theta),]

log2cpm.quant <- readRDS("../output/npreg-trendfilter-quantile.Rmd/log2cpm.quant.rds")


# select external validation samples
set.seed(99)
nvalid <- round(ncol(log2cpm.quant)*.15)
ii.valid <- sample(1:ncol(log2cpm.quant), nvalid, replace = F)
ii.nonvalid <- setdiff(1:ncol(log2cpm.quant), ii.valid)

log2cpm.quant.nonvalid <- log2cpm.quant[,ii.nonvalid]
log2cpm.quant.valid <- log2cpm.quant[,ii.valid]
theta <- pdata$theta
names(theta) <- rownames(pdata)

# theta.nonvalid <- theta_moved[ii.nonvalid]
theta.nonvalid <- theta[ii.nonvalid]
theta.valid <- theta[ii.valid]

sig.genes <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.101.rds")
expr.sig <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% rownames(sig.genes), ]

expr.sig <- expr.sig[1:10,]

# get predicted times
# set training samples
source("../peco/R/primes.R")
source("../peco/R/partitionSamples.R")
parts <- partitionSamples(1:ncol(log2cpm.quant.nonvalid), runs=5,
                          nsize.each = rep(151,5))
part_indices <- parts$partitions
```

Fitting 

```{r}
source("../peco/R/fit.trendfilter.generic.R")
source("../peco/R/cycle.npreg.R")
source("../code/utility.R")
```


```{r, eval = F}
fits.nw <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.sig[,part_indices[[run]]$train]
  theta_train <- theta.nonvalid[part_indices[[run]]$train]
  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr.sig[,part_indices[[run]]$test]
  theta_test <- theta.nonvalid[part_indices[[run]]$test]
  
  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits.nw[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test)
}

saveRDS(fits.nw, file = "../output/method-npreg-prelim-results.Rmd/fits.nw.rds")


fits.trend2 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.sig[,part_indices[[run]]$train]
  theta_train <- theta.nonvalid[part_indices[[run]]$train]
  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    polyorder=2,
                                    ncores=10,
                                    method.trend="trendfilter")
  # fitting test data
  Y_test <- expr.sig[,part_indices[[run]]$test]
  theta_test <- theta.nonvalid[part_indices[[run]]$test]
  
  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="trendfilter",
                                    polyorder=2,
                                    ncores=12)
  
  fits.trend2[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test)
}

saveRDS(fits.trend2, file = "../output/method-npreg-prelim-results.Rmd/fits.trend2.rds")


fits.trend3 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.sig[,part_indices[[run]]$train]
  theta_train <- theta.nonvalid[part_indices[[run]]$train]
  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    polyorder=3,
                                    ncores=10,
                                    method.trend="trendfilter")
  # fitting test data
  Y_test <- expr.sig[,part_indices[[run]]$test]
  theta_test <- theta.nonvalid[part_indices[[run]]$test]
  
  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="trendfilter",
                                    ncores=12)
  
  fits.trend3[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test)
}

saveRDS(fits.trend3, file = "../output/method-npreg-prelim-results.Rmd/fits.trend3.rds")
```

load results

```{r}
fits.nw <- readRDS(file = "../output/method-npreg-prelim-results.Rmd/fits.nw.rds")
fits.trend2 <- readRDS(file = "../output/method-npreg-prelim-results.Rmd/fits.trend2.rds")
fits.trend3 <- readRDS(file = "../output/method-npreg-prelim-results.Rmd/fits.trend3.rds")
```


### Results

```{r}
run=1
Y_test <- expr.sig[,part_indices[[run]]$test]
theta_test <- theta.nonvalid[part_indices[[run]]$test]

time_nw <- fits.nw[[1]]$fit.test$cell_times_est[match(names(theta_test),
                           names(fits.nw[[1]]$fit.test$cell_times_est))]
time_trend2 <- fits.trend2[[1]]$fit.test$cell_times_est[match(names(theta_test),
                           names(fits.trend2[[1]]$fit.test$cell_times_est))]
time_trend3 <- fits.trend3[[1]]$fit.test$cell_times_est[match(names(theta_test),
                           names(fits.trend3[[1]]$fit.test$cell_times_est))]

par(mfrow=c(2,2))
plot(theta_test, time_nw)
plot(theta_test, time_trend2)
plot(theta_test, time_trend3)
```

Misclassification rate

```{r}
xy_time <- lapply(1:5, function(run) {
   xy <- data.frame(
     ref_time=theta.nonvalid[part_indices[[run]]$test],
     pred_time_nw=fits.nw[[run]]$fit.test$cell_times_est[
       match(names(theta.nonvalid[part_indices[[run]]$test]),
             names(fits.nw[[run]]$fit.test$cell_times_est))],
     pred_time_trend2=fits.trend2[[run]]$fit.test$cell_times_est[
       match(names(theta.nonvalid[part_indices[[run]]$test]),
             names(fits.trend2[[run]]$fit.test$cell_times_est))],
     pred_time_trend3=fits.trend3[[run]]$fit.test$cell_times_est[
       match(names(theta.nonvalid[part_indices[[run]]$test]),
             names(fits.trend3[[run]]$fit.test$cell_times_est))])
   return(xy)
})

for (i in 1:5) {
  xy_time[[i]]$diff_time_nw <- pmin(
    abs(xy_time[[i]]$pred_time_nw-xy_time[[i]]$ref_time),
    abs(xy_time[[i]]$pred_time_nw-(2*pi-xy_time[[i]]$ref_time)))
  xy_time[[i]]$diff_time_trend2 <- pmin(
    abs(xy_time[[i]]$pred_time_trend2-xy_time[[i]]$ref_time),
    abs(xy_time[[i]]$pred_time_trend2-(2*pi-xy_time[[i]]$ref_time)))
  xy_time[[i]]$diff_time_trend3 <- pmin(
    abs(xy_time[[i]]$pred_time_trend3-xy_time[[i]]$ref_time),
    abs(xy_time[[i]]$pred_time_trend3-(2*pi-xy_time[[i]]$ref_time)))
}

mean(sapply(xy_time, function(x) mean(x$diff_time_nw))/2/pi)
mean(sapply(xy_time, function(x) mean(x$diff_time_trend2))/2/pi)
mean(sapply(xy_time, function(x) mean(x$diff_time_trend3))/2/pi)
```

---

## npcirc.nw on top 101 genes

Extract data from the top 101 genes identified 

```{r}
library(Biobase)
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

#macosko <- readRDS("data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")
counts <- counts[,order(pdata$theta)]
log2cpm.all <- log2cpm.all[,order(pdata$theta)]
pdata <- pdata[order(pdata$theta),]

log2cpm.quant <- readRDS("../output/npreg-trendfilter-quantile.Rmd/log2cpm.quant.rds")


# select external validation samples
set.seed(99)
nvalid <- round(ncol(log2cpm.quant)*.15)
ii.valid <- sample(1:ncol(log2cpm.quant), nvalid, replace = F)
ii.nonvalid <- setdiff(1:ncol(log2cpm.quant), ii.valid)

log2cpm.quant.nonvalid <- log2cpm.quant[,ii.nonvalid]
log2cpm.quant.valid <- log2cpm.quant[,ii.valid]
theta <- pdata$theta
names(theta) <- rownames(pdata)

# theta.nonvalid <- theta_moved[ii.nonvalid]
theta.nonvalid <- theta[ii.nonvalid]
theta.valid <- theta[ii.valid]

sig.genes.101 <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.101.rds")
expr.sig.101 <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% rownames(sig.genes.101), ]

sig.genes.476 <- readRDS("../output/npreg-trendfilter-quantile.Rmd/out.stats.ordered.sig.476.rds")
expr.sig.476 <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% rownames(sig.genes.476), ]

# get Seurat genes
library(biomaRt)
ensembl <- useMart("ensembl",dataset="hsapiens_gene_ensembl")
symbols <- getBM(attributes=c('ensembl_gene_id','hgnc_symbol'), 
      filter='ensembl_gene_id', 
      values = rownames(log2cpm.all), mart = ensembl)
head(symbols)

cc.genes.symbol <- readLines(con = "../data/cellcycle-genes-previous-studies/seurat_cellcycle/regev_lab_cell_cycle_genes.txt")
# s.genes <- cc.genes[1:43]
# g2m.genes <- cc.genes[44:97]
cc.genes.ensembl <- symbols$ensembl_gene_id[which(symbols$hgnc_symbol %in% cc.genes.symbol)]
expr.seurat <- log2cpm.quant.nonvalid[rownames(log2cpm.quant.nonvalid) %in% cc.genes.ensembl, ]


# get predicted times
# set training samples
source("../peco/R/primes.R")
source("../peco/R/partitionSamples.R")
parts <- partitionSamples(1:ncol(log2cpm.quant.nonvalid), runs=5,
                          nsize.each = rep(151,5))
part_indices <- parts$partitions
```

overlap between our top genes and seurat genes

```{r, eval=F}
library(VennDiagram)
grid.draw(venn.diagram(list(
    "peco top 101 genes" = rownames(sig.genes.101),
    "Seurat 92 genes" = cc.genes.ensembl), filename=NULL))

grid.draw(venn.diagram(list(
    "peco top 476 genes" = rownames(sig.genes.476),
    "Seurat 92 genes" = cc.genes.ensembl), filename=NULL))
```

Fitting 

```{r}
source("../peco/R/fit.trendfilter.generic.R")
source("../peco/R/cycle.npreg.R")
source("../code/utility.R")
```


```{r, eval = F}
fits.101 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.sig.101[,part_indices[[run]]$train]
  theta_train <- theta.nonvalid[part_indices[[run]]$train]
  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr.sig.101[,part_indices[[run]]$test]
  theta_test <- theta.nonvalid[part_indices[[run]]$test]
  
  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits.101[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test)
}

saveRDS(fits.101, file = "../output/method-npreg-prelim-results.Rmd/fits.101.rds")


fits.476 <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.sig.476[,part_indices[[run]]$train]
  theta_train <- theta.nonvalid[part_indices[[run]]$train]
  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr.sig.476[,part_indices[[run]]$test]
  theta_test <- theta.nonvalid[part_indices[[run]]$test]
  
  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits.476[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test)
}

saveRDS(fits.476, file = "../output/method-npreg-prelim-results.Rmd/fits.476.rds")


fits.seurat <- vector("list", 5)
for (run in 1:5) {
  print(run)
  # fitting training data
  Y_train <- expr.seurat[,part_indices[[run]]$train]
  theta_train <- theta.nonvalid[part_indices[[run]]$train]
  fit.train <- cycle.npreg.insample(Y = Y_train, 
                                    theta = theta_train, 
                                    ncores=10,
                                    method.trend="npcirc.nw")
  # fitting test data
  Y_test <- expr.seurat[,part_indices[[run]]$test]
  theta_test <- theta.nonvalid[part_indices[[run]]$test]
  
  fit.test <- cycle.npreg.outsample(Y_test=Y_test,
                                    sigma_est=fit.train$sigma_est,
                                    funs_est=fit.train$funs_est,
                                    method.grid = "uniform",
                                    method.trend="npcirc.nw",
                                    ncores=12)
  
  fits.seurat[[run]] <- list(fit.train=fit.train,
                      fit.test=fit.test)
}

saveRDS(fits.seurat, file = "../output/method-npreg-prelim-results.Rmd/fits.seurat.rds")
```

quick checking on the results

```{r}
fits.101 <- readRDS("../output/method-npreg-prelim-results.Rmd/fits.101.rds")
fits.476 <- readRDS("../output/method-npreg-prelim-results.Rmd/fits.476.rds")
fits.seurat <- readRDS("../output/method-npreg-prelim-results.Rmd/fits.seurat.rds")
```

average log-likelihood

```{r}
mean(sapply(1:5, function(run) {
  fits.101[[run]]$fit.test$loglik_est}))

mean(sapply(1:5, function(run) {
  fits.476[[run]]$fit.test$loglik_est}))

mean(sapply(1:5, function(run) {
  fits.seurat[[run]]$fit.test$loglik_est}))
```


```{r}
run=1
Y_test.101 <- expr.sig.101[,part_indices[[run]]$test]
Y_test.476 <- expr.sig.476[,part_indices[[run]]$test]
Y_test.seurat <- expr.seurat[,part_indices[[run]]$test]
theta_test <- theta.nonvalid[part_indices[[run]]$test]

all.equal(names(theta_test), names(fits.101[[1]]$fit.test$cell_times_est))
all.equal(names(theta_test), names(fits.476[[1]]$fit.test$cell_times_est))
all.equal(names(theta_test), names(fits.seurat[[1]]$fit.test$cell_times_est))


par(mfrow=c(1,2))
plot(theta_test, fits.101[[1]]$fit.test$cell_times_est,
     xlab="reference cell time (from fucci)", 
     ylab="estimated cell time using training data",
     main = "top 101 genes")
plot(theta_test, fits.476[[1]]$fit.test$cell_times_est,
     xlab="reference cell time (from fucci)", 
     ylab="estimated cell time using training data",
     main = "top 476 genes")

par(mfrow=c(1,1))
plot(theta_test, fits.seurat[[1]]$fit.test$cell_times_est,
     xlab="reference cell time (from fucci)", 
     ylab="estimated cell time using training data",
     main = "Seurat 92 genes")
```


Compute differences in predicted time.

```{r}
xy_time <- lapply(1:5, function(run) {
  xy <- data.frame(
    ref_time=theta.nonvalid[part_indices[[run]]$test],
    pred_time.101=fits.101[[run]]$fit.test$cell_times_est,
    pred_time.476=fits.476[[run]]$fit.test$cell_times_est,
    pred_time.seurat=fits.seurat[[run]]$fit.test$cell_times_est,
    dapi=pdata$dapi.median.log10sum.adjust[match(names(theta.nonvalid[part_indices[[run]]$test]), rownames(pdata))],
    gfp=pdata$gfp.median.log10sum.adjust[match(names(theta.nonvalid[part_indices[[run]]$test]), rownames(pdata))],
    rfp=pdata$rfp.median.log10sum.adjust[match(names(theta.nonvalid[part_indices[[run]]$test]), rownames(pdata))])
  return(xy)
})

for (run in 1:5) {
  xy_time[[run]]$diff_time.101 <- pmin(
    abs(xy_time[[run]]$pred_time.101-xy_time[[i]]$ref_time),
    abs(xy_time[[run]]$pred_time.101-(2*pi-xy_time[[i]]$ref_time)))
  xy_time[[run]]$diff_time.476 <- pmin(
    abs(xy_time[[run]]$pred_time.476-xy_time[[i]]$ref_time),
    abs(xy_time[[run]]$pred_time.476-(2*pi-xy_time[[i]]$ref_time)))
  xy_time[[run]]$diff_time.seurat <- pmin(
    abs(xy_time[[run]]$pred_time.seurat-xy_time[[i]]$ref_time),
    abs(xy_time[[run]]$pred_time.seurat-(2*pi-xy_time[[i]]$ref_time)))
}
saveRDS(xy_time, file = "../output/method-npreg-prelim-results.Rmd/xy_time.rds")
```

```{r}
xy_time <- readRDS(file = "../output/method-npreg-prelim-results.Rmd/xy_time.rds")

cbind(mean(sapply(xy_time, function(x) mean(x$diff_time.101))/2/pi),
      sd(sapply(xy_time, function(x) mean(x$diff_time.101))/2/pi))
cbind(mean(sapply(xy_time, function(x) mean(x$diff_time.476))/2/pi),
      sd(sapply(xy_time, function(x) mean(x$diff_time.476))/2/pi))
cbind(mean(sapply(xy_time, function(x) mean(x$diff_time.seurat))/2/pi),
      sd(sapply(xy_time, function(x) mean(x$diff_time.seurat))/2/pi))
```


## print 101 genes reults

```{r}
par(mfrow=c(1,2))
plot(xy_time[[1]]$pred_time.101, xy_time[[1]]$ref_time,
     ylab="reference cell time (from fucci)", 
     xlab="estimated cell time")
abline(0,1, col="blue")
plot(xy_time[[1]]$pred_time.101, xy_time[[1]]$diff_time.101,
     xlab="estimated cell time (from fucci)", 
     ylab="differences in cell times")
title("Top 101 cyclical genes", outer=TRUE, line=-1)
```




---

## Session information

```{r}
sessionInfo()
```

