---
title: "Trying out spml package"
author: "Joyce Hsiao"
output: 
  html_document:
    toc: TRUE
    toc_float: FALSE
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```



---


## Behavior of spml

Use `spml` functions implemented in `Rfast` package. See https://cran.r-project.org/web/packages/Rfast/index.html


### Univariate 

In a simple univariate case, the model converges to the same likelihood over 50 runs in same number of iterations, given a fixed toelerance or between two difference tolerance levels.

```{r}
# simulate a scenario where 
# the response variable y is bivariate normal random variables
# with mean Mu and variance Sigma
# We are interested in the relatsionship Mu = BX
# where X is a n by p matrix representing possibly gene expression profile of n cells

library(Rfast)
library(circular)

N <- 100
set.seed(17)
X <- rnorm(N)
B1 <- cbind(3,1)
B2 <- cbind(2,-3)
Mu <- cbind(3 + 1*X, 2 - 3*X)
#Mu <- X %*% t(B)
Y <- cbind(rnorm(N, Mu[,1],1),
           rnorm(N, Mu[,2],1))
# project Y onto unit circle
U <- Y / sqrt( rowsums(Y^2) )
Theta <- coord2rad(U, contro.circular(modulo = "2pi", units = "radians", zero = 0))

nevals <- 50
res.covariate.one <- lapply(1:nevals, function(i) {
  set.seed(i)
  spml.reg(y=Theta, x=X, seb = TRUE)
})
sapply(res.covariate.one, "[[", "loglik")
sapply(res.covariate.one, "[[", "iters")
```

Do we get very different results if changing tolerance?

```{r}
nevals <- 50
res.covariate.one.2 <- lapply(1:nevals, function(i) {
  set.seed(i)
  spml.reg(y=Theta, x=X, tol = 1e-10, seb = TRUE)
})
sapply(res.covariate.one.2, "[[", "loglik")
sapply(res.covariate.one.2, "[[", "iters")
```

Compare results when tolerance is 1e-07 versus 1e-10. No difference. 

```{r}
res.covariate.one[[1]]$loglik
res.covariate.one.2[[1]]$loglik
```


### P large but P < N

Okay with 10 variable. Converges to the same likelihood in 50 runs.

```{r}
# simulate a scenario where 
# the response variable y is bivariate normal random variables
# with mean Mu and variance Sigma
# We are interested in the relatsionship Mu = BX
# where X is a n by p matrix representing possibly gene expression profile of n cells

library(Rfast)
N <- 100
P <- 10
set.seed(17)
X <- matrix(rnorm(N*P), nrow= N, ncol = P)
B1 <- cbind(3,1, 2, 5, 1, .5, .3, .1, 1, 2)
B2 <- cbind(2,-3, 1, 2, .2, .3, .1, .9, 5, 1)
B <- rbind(B1, B2)
Mu <- X%*%t(B)
Y <- cbind(rnorm(N, Mu[,1],1),
           rnorm(N, Mu[,2],1))
# project Y onto unit circle
U <- Y / sqrt( rowsums(Y^2) )
Theta <- coord2rad(U, contro.circular(modulo = "2pi", units = "radians", zero = 0))

nevals <- 50
res.covariate.many <- lapply(1:nevals, function(i) {
  set.seed(i)
  spml.reg(y=Theta, x=X, seb = TRUE)
})
sapply(res.covariate.many, "[[", "loglik")
sapply(res.covariate.many, "[[", "iters")
```


---

## Session information

```{r}
sessionInfo()
```

