---
title: "Find genes with cyclical patterns"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Data and packages

Packages

```{r}
library(circular)
library(conicfit)
library(Biobase)
library(dplyr)
library(matrixStats)

library(NPCirc)
library(smashr)
```

Load data

```{r}
df <- readRDS(file="../data/eset-filtered.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

# log2cpm.all <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.rds")
# log2cpm.adjust <- readRDS("../output/seqdata-batch-correction.Rmd/log2cpm.adjust.rds")
# log2cpm <- log2cpm.all[grep("ENSG", rownames(log2cpm.all)), ]

# import corrected intensities
pdata.adj <- readRDS("../output/images-normalize-anova.Rmd/pdata.adj.rds")

macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")

# log2cpm.high <- log2cpm.detected[order(rowMeans(log2cpm.detected))[1:100], ]

source("../code/utility.R")

summary(colSums(counts))
summary(rowMeans(log2cpm.all>0))
log2cpm.detected <- log2cpm.all[rowMeans(log2cpm.all>0)>.8,]
```

---

## Projected normal on PCs of Red/Green

```{r, fig.width=7, fig.height=5}
pc.fucci <- prcomp(subset(pdata.adj, 
                          select=c("rfp.median.log10sum.adjust",
                                   "gfp.median.log10sum.adjust")),
                   center = T, scale. = T)
Theta.cart <- pc.fucci$x
library(circular)
Theta.fucci <- coord2rad(Theta.cart)

df <- data.frame(theta=as.numeric(Theta.fucci))
rownames(df) <- rownames(pdata.adj)
Theta.fucci <- 2*pi - Theta.fucci
#saveRDS(as.numeric(Theta.fucci), file = "../output/npreg.Rmd/theta.rds")
```

--


## Compare different methods

* Analysis was done in a batch job. See `/code/npreg.Rmd` for codes for submitting the batch job.

* Three methods are compared: smash, kernel regression using NW-estimate, and kernel regression using LL-estimate. 

* In running smash, I imputed the zero-valued expression with mean of expression for each gene. In running kernel methods, I omitted the zero-valued cells 

* For comparison involving smash, we take 512 randomly selected cells and analyze these for all three methods. MAD is employed to evaluate the fit. Specifically, MAD(data-fitted mean)/MAD(data-data.mean).

* For comparison of the two kernel regression methods, I included all of the ~900 cells. PVE (proportion of variance explained) is computed to evalulate the fit. Specifically, 1-var(data-fitted mean)/var(data). 

* `smashr` package was used to run smash, and `NPCirc` package was used to run the two kernel regression methods.


---

## Results using MAD

Compare MAD ratio of data from the fitted mean versus the grand mean (not fitted). Briefly,

$$
mad.ratio = MAD(expression_g - fitted.g)/MAD(expression_g - mean.g)
$$

```{r}
out.methods <- readRDS("../output/npreg-methods.Rmd/out.methods.rds")

mad.ratio <- data.frame(smash.mad.ratio=sapply(out.methods, "[[", "smash.mad.ratio"),
                        smash.pois.mad.ratio=sapply(out.methods, "[[", "smash.pois.mad.ratio"),
                        npll.mad.ratio=sapply(out.methods, "[[", "npll.mad.ratio"),
                        npnw.mad.ratio=sapply(out.methods, "[[", "npnw.mad.ratio"))

pve <- data.frame(npll.pve=sapply(out.methods, "[[", "npll.pve"),
                  npnw.pve=sapply(out.methods, "[[", "npnw.pve"))

boxplot(mad.ratio)
boxplot(pve)
```

enrichment function choose top ranked genes

```{r}
enrich.order <- function(cutoffs, metrics, cyclegenes, allgenes) {
  #  out <- order(mad.ratio$smash.mad.ratio)
  # cutoffs <- c(100, 200, 300)
  cycle.rich <- sapply(cutoffs, function(x) {
    sig.cycle <- sum(allgenes[metrics[1:x]] %in% cyclegenes)/x
    non.cycle <- sum(allgenes[-metrics[1:x]] %in% cyclegenes)/(length(allgenes)-x)
    cbind(as.numeric(sum(allgenes[metrics[1:x]] %in% cyclegenes)), 
          sig.cycle/non.cycle)
  })
  colnames(cycle.rich) <- cutoffs
  rownames(cycle.rich) <- c("nsig.genes.cycle", "fold.sig.vs.nonsig.cycle")
  cycle.rich 
}
```

smash gaussian enrichment.

```{r}
enrich.order(cutoffs = c(100, 200, 300), 
             metrics = order(mad.ratio$smash.mad.ratio),
             cyclegenes = macosko$ensembl,
             allgenes = rownames(log2cpm.all))
```

smash poisson enrichment.

```{r}
enrich.order(cutoffs = c(100, 200, 300), 
             metrics = order(mad.ratio$smash.pois.mad.ratio),
             cyclegenes = macosko$ensembl,
             allgenes = rownames(log2cpm.all))
```

kernel LL-estimator enrichment

```{r}
enrich.order(cutoffs = c(100, 200, 300), 
             metrics = order(mad.ratio$npll.mad.ratio),
             cyclegenes = macosko$ensembl,
             allgenes = rownames(log2cpm.all))
```

kernel NW-estimator enrichment

```{r}
enrich.order(cutoffs = c(100, 200, 300), 
             metrics = order(mad.ratio$npnw.mad.ratio),
             cyclegenes = macosko$ensembl,
             allgenes = rownames(log2cpm.detected))
```




Check smash genes.

```{r}
ii.check <- order(mad.ratio$smash.mad.ratio)[c(1:16)]

library(mygene)
par(mfrow=c(4,4), mar=c(3,2,2,1))
ensg <- rownames(log2cpm.detected)[ii.check]
symbols <- queryMany(ensg,  scopes="ensembl.gene", 
                     fields=c("symbol"), species="human")@listData$symbol
for (i in 1:length(ii.check)) {
  ii <- ii.check[i]
  with(out.methods[[ii]], {
      plot(y=yy.train,
           x=xx.train,
           xlab = "estimated cell time", ylab = "log2CPM", col = "gray30",
           main=paste(symbols[i], ",", round(mad.ratio$smash.mad.ratio[ii],2)))
      lines(x=smash.xx,
         y=smash.yy, col = "red", pch = 16, cex=.6)
  })
}
```


Check kernel LL genes.

```{r}
ii.check <- order(mad.ratio$npll.mad.ratio)[c(1:20)]

library(mygene)
par(mfrow=c(4,5), mar=c(3,2,2,1))
ensg <- rownames(log2cpm.detected)[ii.check]
symbols <- queryMany(ensg,  scopes="ensembl.gene", 
                     fields=c("symbol"), species="human")@listData$symbol
for (i in 1:length(ii.check)) {
  ii <- ii.check[i]
  with(out.methods[[ii]], {
      plot(y=yy.train,
           x=xx.train,
           xlab = "estimated cell time", ylab = "log2CPM", col = "gray30",
           main=paste(symbols[i], ",", round(mad.ratio$smash.mad.ratio[ii],2)))
      lines(x=smash.xx,
         y=smash.yy, col = "red", pch = 16, cex=.6)
  })
}
```


## Results using PVE for kernel-based method

```{r}
pve <- data.frame(npll.pve=sapply(out.methods, "[[", "npll.pve"),
                  npnw.pve=sapply(out.methods, "[[", "npnw.pve"))
boxplot(pve)
```

enrichment function choose top ranked genes

```{r}
enrich.pve <- function(cutoffs, metrics, cyclegenes, allgenes) {
  #  out <- order(mad.ratio$smash.mad.ratio)
  # cutoffs <- c(100, 200, 300)
  cycle.rich <- sapply(cutoffs, function(x) {
    sig.cycle <- sum(allgenes[metrics>x] %in% cyclegenes)/length(allgenes[metrics>x])
    non.cycle <- sum(allgenes[metrics<x] %in% cyclegenes)/length(allgenes[metrics<x])
    cbind(as.numeric(sum(allgenes[metrics>x] %in% cyclegenes)), 
          sig.cycle/non.cycle)
  })
  colnames(cycle.rich) <- cutoffs
  rownames(cycle.rich) <- c("nsig.genes.cycle", "fold.sig.vs.nonsig.cycle")
  cycle.rich 
}
```

kernel LL-estimator enrichment

```{r}
enrich.pve(cutoffs = c(.05, .1, .2, .3), 
             metrics = pve$npll.pve,
             cyclegenes = macosko$ensembl,
             allgenes = rownames(log2cpm.all))

enrich.order(cutoffs = c(100, 200, 300), 
             metrics = order(pve$npll.pve, decreasing=T),
             cyclegenes = macosko$ensembl,
             allgenes = rownames(log2cpm.all))
```

kernel NW-estimator enrichment

```{r}
enrich.pve(cutoffs = c(.05, .1, .2, .3), 
             metrics = pve$npnw.pve,
             cyclegenes = macosko$ensembl,
             allgenes = rownames(log2cpm.all))

enrich.order(cutoffs = c(100, 200, 300), 
             metrics = order(pve$npnw.pve, decreasing=T),
             cyclegenes = macosko$ensembl,
             allgenes = rownames(log2cpm.all))
```



Print top 500 genes.

```{r}
write.table(rownames(log2cpm.all)[order(pve$npll.pve, decreasing=T)[1:5000]],
            file = "../output/npreg-methods.Rmd/npll.genes.txt",
            quote=F, col.names=F, row.names=F)

write.table(rownames(log2cpm.all)[order(pve$npnw.pve, decreasing=T)[1:5000]],
            file = "../output/npreg-methods.Rmd/npnw.genes.txt",
            quote=F, col.names=F, row.names=F)

write.table(rownames(log2cpm.all),
            file = "../output/npreg-methods.Rmd/allgenes.txt",
            quote=F, col.names=F, row.names=F)
```


