---
title: "npreg: trendfilter"
author: "Joyce Hsiao"
output: 
  html_document:
    toc: TRUE
    toc_float: FALSE
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Data and packages

Packages

```{r}
library(circular)
library(conicfit)
library(Biobase)
library(dplyr)
library(matrixStats)

library(NPCirc)
library(smashr)
library(genlasso)
```

Load data

```{r}
df <- readRDS(file="../data/eset-final.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]

log2cpm.all <- t(log2(1+(10^6)*(t(counts)/pdata$molecules)))

macosko <- readRDS("../data/cellcycle-genes-previous-studies/rds/macosko-2015.rds")

theta <- readRDS("../output/images-time-eval.Rmd/theta.rds")
log2cpm.all.ord <- log2cpm.all[,order(theta)]

source("../code/utility.R")
source("../code/npreg/npreg.methods.R")
```

--

## Emprical data for evaluating methods

Consider 5 genes previously identified to have cyclical patterns and 5 genes previously identified to not have cyclical patterns. Note that there's a discrepany (albeit small) between the samples that was used to identify cyclical patterns and the samples that are in the finalized data.

```{r}
examples <- readRDS("../output/npreg-methods.Rmd/cyclegenes.rds")

library(biomaRt)
ensembl <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")
symbols <- getBM(attributes = c("hgnc_symbol",'ensembl_gene_id'), 
      filters = c('ensembl_gene_id'),
      values = colnames(examples)[-1], 
      mart = ensembl)
symbols <- symbols[match(colnames(examples)[-1], symbols$ensembl_gene_id),]

par(mfrow=c(2,5))
for (i in 1:10) {
  plot(examples$theta,
       examples[,1+i], main = symbols$hgnc_symbol[i])
}
```

Get ENSG IDs.

```{r}
cycles <- symbols[1:5,]
cycles.not <- symbols[6:10,]
```

Get these genes from the updated data.

```{r}
cycles.log2cpm <- log2cpm.all.ord[rownames(log2cpm.all.ord) %in% cycles$ensembl_gene_id,]
cycles.not.log2cpm <- log2cpm.all.ord[rownames(log2cpm.all.ord) %in% cycles.not$ensembl_gene_id,]
```

confirm pattern in the udpated dataset.

```{r, fig.width=7,fig.height=4}
par(mfrow=c(2,5))
for (i in 1:5) {
  plot(cycles.log2cpm[i,], main = cycles$hgnc_symbol[i])
}
for (i in 1:5) {
  plot(cycles.not.log2cpm[i,], main = cycles.not$hgnc_symbol[i])
}
```

## Property of genes with many zeros

1. As expected, genes with many zeros tend to have lower mean non-zero molecule count.

2. Note that at gene mean log2cpm of 4, the average molecule count is about 16 across the 880 samples, which is not many.

```{r}
log2cpm.all.impute <- log2cpm.all
ii.zero <- which(log2cpm.all == 0, arr.ind = TRUE)
log2cpm.all.impute[ii.zero] <- NA

gene_nas <- rowMeans(log2cpm.all==0)
gene_means <- rowMeans(log2cpm.all.impute, na.rm=TRUE)
plot(x=gene_nas, y=gene_means)
```

---

## Simulate from the empirical data

The goal of simulation is to compare the methods in their ability to recover cyclical patterns despite of zero observations. 

1. Scenario 1: original data

2. Scenario 2: subsample to include 20% missing


```{r, fig.width=7,fig.height=4}
N <- ncol(log2cpm.all)
scene.1 <- cycles.log2cpm
scene.2 <- do.call(rbind, lapply(1:nrow(cycles.log2cpm), function(g) {
  yy <- cycles.log2cpm[g,]
  numzeros <- round(N*0.2)
  numzeros <- numzeros - sum(yy==0)
  which.nonzero <- which(yy!=0)
  ii.zeros <- sample(which.nonzero,numzeros, replace = F)
  yy[ii.zeros] <- 0
  return(yy)
}))


par(mfrow=c(2,5))
for (i in 1:5) {
  plot(scene.1[i,], main = cycles$hgnc_symbol[i])
}
for (i in 1:5) {
  plot(scene.2[i,], main = cycles$hgnc_symbol[i])
}
```

## Fitting trendfilter 

out.1: excluding zero

out.2: including zero

```{r, eval =F}
out.1 <- lapply(1:5, function(i) {
  yy <- scene.2[i,]
  theta.sam <- theta
  out.trend <- fit.trendfilter(yy=yy, pos.yy=c(1:length(yy)))

  return(list(yy.sam=yy,
              theta.sam=theta.sam,
              out.trend=out.trend))
})

out.2 <- lapply(1:5, function(i) {
  yy <- scene.2[i,]
  out.trend <- fit.trendfilter.includezero(yy=yy, pos.yy=c(1:length(yy)))

  return(list(yy.sam=yy,
              theta.sam=theta,
              out.trend=out.trend))
})

saveRDS(out.1, file = "../output/npreg-trendfilter-prelim.Rmd/out1.rds")
saveRDS(out.2, file = "../output/npreg-trendfilter-prelim.Rmd/out2.rds")
```


```{r}
out.1 <- readRDS(file = "../output/npreg-trendfilter-prelim.Rmd/out1.rds")
out.2 <- readRDS(file = "../output/npreg-trendfilter-prelim.Rmd/out2.rds")

par(mfrow=c(2,5))
for (i in 1:5) {
  out <- out.1[[i]]
  plot(c(1:length(out$yy.sam)), out$yy.sam, col = "gray50")
  with(out$out.trend, 
       points(trend.pos, trend.yy, col = "royalblue"), cex=.5, pch=16)
}
for (i in 1:5) {
  out <- out.2[[i]]
  plot(c(1:length(out$yy.sam)), out$yy.sam, col = "gray50")
  with(out$out.trend, 
       points(trend.pos, trend.yy, col = "royalblue"), cex=.5, pch=16)
}
```


---

## Eval for selecting polynomial order

```{r, eval =F}
theta <- readRDS("../output/images-time-eval.Rmd/theta.rds")

chip_ids <- c("NA18511","NA18855","NA18870","NA19098","NA19101","NA19160")

id <- which(pdata$chip_id == chip_ids[1])
log2cpm.sub <- log2cpm.all[rownames(log2cpm.all) %in% symbols$ensembl_gene_id[1:5],]
theta.train <- theta[-id]
theta.test <- theta[id]

out.order2 <- lapply(1:5, function(i) {
  yy <- log2cpm.sub[i,]
  yy.train <- yy[-id]
  yy.test <- yy[id]
  pos.yy.train <- c(1:length(yy))[-id]
  pos.yy.test <- c(1:length(yy))[id]
  nonzero.test <- yy.test!=0
    
  out.trend <- fit.trendfilter(yy=yy.train, pos.yy = pos.yy.train, polyorder = 2)
  f <- approxfun(x=out.trend$trend.pos, y=out.trend$trend.yy, rule=2)  
  
  test.mad.pred <- mad(yy.test[nonzero.test]-f(pos.yy.test[nonzero.test]), constant = 1)
  test.mad.constant <- mad(yy.test[nonzero.test]-mean(yy.test[nonzero.test]), constant = 1)
  test.mad.ratio <- test.mad.pred/test.mad.constant

  return(list(yy=yy,
              yy.train=yy.train,
              pos.train=pos.yy.train,
              yy.test=yy.test,
              pos.test=pos.yy.test,
              pred.train.yy=out.trend$trend.yy,
              pred.train.xx=out.trend$trend.pos,
              pred.test.yy=f(pos.yy.test),
              pred.test.xx=pos.yy.test,
              test.mad.ratio=test.mad.ratio,
              train.mad.ratio=out.trend$trend.mad.ratio))
})


out.order3 <- lapply(1:5, function(i) {
  yy <- log2cpm.sub[i,]
  yy.train <- yy[-id]
  yy.test <- yy[id]
  pos.yy.train <- c(1:length(yy))[-id]
  pos.yy.test <- c(1:length(yy))[id]
  nonzero.test <- yy.test!=0
    
  out.trend <- fit.trendfilter(yy=yy.train, pos.yy = pos.yy.train, polyorder = 3)
  f <- approxfun(x=out.trend$trend.pos, y=out.trend$trend.yy, rule=2)  
  
  test.mad.pred <- mad(yy.test[nonzero.test]-f(pos.yy.test[nonzero.test]), constant = 1)
  test.mad.constant <- mad(yy.test[nonzero.test]-mean(yy.test[nonzero.test]), constant = 1)
  test.mad.ratio <- test.mad.pred/test.mad.constant

  return(list(yy=yy,
              yy.train=yy.train,
              pos.train=pos.yy.train,
              yy.test=yy.test,
              pos.test=pos.yy.test,
              pred.train.yy=out.trend$trend.yy,
              pred.train.xx=out.trend$trend.pos,
              pred.test.yy=f(pos.yy.test),
              pred.test.xx=pos.yy.test,
              test.mad.ratio=test.mad.ratio,
              train.mad.ratio=out.trend$trend.mad.ratio))
})

saveRDS(out.order2, file = "../output/npreg-trendfilter-prelim.Rmd/out.order2.rds")
saveRDS(out.order3, file = "../output/npreg-trendfilter-prelim.Rmd/out.order3.rds")
```

```{r}
out.order2 <- readRDS(file = "../output/npreg-trendfilter-prelim.Rmd/out.order2.rds")
out.order3 <- readRDS(file = "../output/npreg-trendfilter-prelim.Rmd/out.order3.rds")

par(mfrow=c(2,5))
for (i in 1:length(out.order2)) {
  with(out.order2[[i]], {
    plot(y=yy.train, x=pos.train, col = "gray50")
    lines(y=pred.train.yy, x=pred.train.xx, col = "black", lwd=1)
    points(y=yy.test,x=pos.test, col = "blue", pch=16, cex=.7)
    lines(y=pred.test.yy, x=pred.test.xx, col = "blue", lwd=2)
  })
}
for (i in 1:length(out.order3)) {
  with(out.order3[[i]], {
    plot(y=yy.train, x=pos.train, col = "gray50")
    lines(y=pred.train.yy, x=pred.train.xx, col = "black", lwd=1)
    points(y=yy.test,x=pos.test, col = "blue", pch=16, cex=.7)
    lines(y=pred.test.yy, x=pred.test.xx, col = "blue", lwd=2)
  })
}

test.order2 <- sapply(out.order2, function(x) x$test.mad.ratio)
test.order3 <- sapply(out.order3, function(x) x$test.mad.ratio)
cbind(test.order2,test.order3)
```


---


## Session information

```{r}
sessionInfo()
```



