---
title: "Select cell cycle genes"
author: "Joyce Hsiao"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Overview

1. Select genes annotated to cell cycle in the GO annotation (GO:0007049).

2. Select genes documented in CycleBase - previously found to relate to cell cycle in microarray studies. 

3. Select genes whose expression variability is above certain threshold. 


This document generated several RDS for later use:

*`genes.go.rds`: 677 human cell cycle genes associated with GO:0007049

*`genes.detect.rds`: 6313 genes in the current study with detection rate > 50% 

*`genes.go.detect.rds`: 450 genes in the current study with detection rate > 50%  and associated with GO:0007049

*`genes.variable.rds`: 359 genes in the current study with detection rate > 50% and dispersion z-score > 1.5 across expression mean bins

*`genes.go.variable.rds`: 24 genes in the current study with detection rate > 50% and dispersion z-score > 1.5 across expression mean bins and associated with GO:0007049



---

## Data and packages

Packages

```{r}
library(biomaRt)
library(Biobase)
```

Load data

```{r}
df <- readRDS(file="../data/eset-filtered.rds")
pdata <- pData(df)
fdata <- fData(df)

# select endogeneous genes
counts <- exprs(df)[grep("ENSG", rownames(df)), ]
```


## GO cell cycle

```{r}
ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl") 
#gets gene symbol, transcript_id and go_id for all genes annotated with GO:0007507
genes.go <- getBM(attributes=c('hgnc_symbol', 'ensembl_gene_id', 'go_id'),
                   filters = 'go', values = 'GO:0007049', mart = ensembl)
genes.go <- unique(genes.go$ensembl_gene_id)
```


## Variable genes

Across individuals, I computed gene mean and coefficient of variation including cells that are detected as expressed for each gene. 


```{r}
gene.means <- apply(counts, 1, function(x) {
  ii_pos <- x >0
  mean(x[ii_pos])
})


gene.cvs <- apply(counts, 1, function(x) {
  ii_pos <- x >0
  mn <- mean(x[ii_pos])
  sd <- sd(x[ii_pos])
  sd/mn
})

gene.drops <- rowMeans(counts>0)
```

Relationship between gene mean and CV is the same as we observed before in the batch paper. The genes with low average expression across cells are those that have high dropout rate. 

```{r, fig.width=7, fig.height=4}
par(mfrow=c(1,2))
plot(x=log10(gene.means), y = gene.cvs,
     xlab = "log10 gene mean count", ylab = "CV")
abline(v=.25, col = "red")
plot(x=gene.drops, y = log10(gene.means),
     xlab = "Proportion of samples detected (count 0)",
     ylab = "log10 gene mean count")
abline(v=.5, col = "blue")
abline(h=.25, col = "red")
```

Two steps here for choosing variable genes:

1. I arbitrary decided on a cutoff for gene mean based on the relationship between CV and log10 gene count. The idea was to filter the genes among which dispersion increases with gene mean (approximately when detection rate 50% or lower).

2. Scan across gene expression bins, identify the ones with high variability in each bin. 

```{r}
genes.detect <- rownames(counts)[which(gene.drops > .5)]

gene.means.detect <- gene.means[which(gene.drops > .5)]
gene.cvs.detect <- gene.cvs[which(gene.drops > .5)]

bins <- cut(gene.means.detect, 
            breaks = quantile(gene.means.detect, prob = seq(0,1, .05)), 
            include.lowest = TRUE)
bins <- as.numeric(bins)

genes.variable <- do.call(c, lapply(1:length(bins), function(i) {
  ii_bin <- bins == i
  cv.z <- scale(gene.cvs.detect[ii_bin])
  rownames(cv.z)[which(cv.z>1.5)]
}) )

plot(x=log10(gene.means.detect), 
     y=gene.cvs.detect,
     xlab = "log10 gene mean count",
     ylab = "CV", main = "Genes detected in > 50% cells",
     col = 1+as.numeric(names(gene.means.detect) %in% genes.variable))
```

---

## Output


```{r, eval = F}
saveRDS(genes.go,
        file = "../output/seqdata-select-cellcyclegenes.Rmd/genes.go.rds")
saveRDS(genes.variable,
        file = "../output/seqdata-select-cellcyclegenes.Rmd/genes.variable.rds")

saveRDS(genes.detect,
        file = "../output/seqdata-select-cellcyclegenes.Rmd/genes.detect.rds")

genes.go.detect <- genes.detect[which(genes.detect %in% genes.go)]
saveRDS(genes.go.detect,
        file = "../output/seqdata-select-cellcyclegenes.Rmd/genes.go.detect.rds")

genes.go.variable <- genes.variable[which(genes.variable %in% genes.go)]
saveRDS(genes.go.variable,
        file = "../output/seqdata-select-cellcyclegenes.Rmd/genes.go.variable.rds")
```


---

## Session information

```{r, echo = FALSE}
sessionInfo()
```





