---
title: "Cyclical trends for expression data of each individual"
author: "Joyce Hsiao"
output: 
  workflowr::wflow_html:
    toc: TRUE
    toc_float: FALSE
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

---

## Summary

* Evaluate cyclical trends of expression levels for each gene in each individual

* Jobs run on cluster using code in `code/trendfilter-individual.Rmd`

* Results saved in `output/trendfilter-individual.Rmd`

* The top 5 cyclical genes are the same across the 6 individuals. When looking at the overlap of the top cyclical genes across individuals, I found that the top 5 cyclical genes are detected in the top 500 cyclial genes across the individuals.

---


## Import results

```{r}
fits_NA18511 <- readRDS("../output/trendfilter-individual.Rmd/fits_NA18511.rds")
fits_NA18855 <- readRDS("../output/trendfilter-individual.Rmd/fits_NA18855.rds")
fits_NA18870 <- readRDS("../output/trendfilter-individual.Rmd/fits_NA18870.rds")
fits_NA19098 <- readRDS("../output/trendfilter-individual.Rmd/fits_NA19098.rds")
fits_NA19101 <- readRDS("../output/trendfilter-individual.Rmd/fits_NA19101.rds")
fits_NA19160 <- readRDS("../output/trendfilter-individual.Rmd/fits_NA19160.rds")
```


---

## Analysis


```{r}
pve_NA18511 <- sapply(fits_NA18511, "[[", 2)
pve_NA18855 <- sapply(fits_NA18855, "[[", 2)
pve_NA18870 <- sapply(fits_NA18870, "[[", 2)
pve_NA19098 <- sapply(fits_NA19098, "[[", 2)
pve_NA19101 <- sapply(fits_NA19101, "[[", 2)
pve_NA19160 <- sapply(fits_NA19160, "[[", 2)

pve_NA18511_ord <- pve_NA18511[order(pve_NA18511, decreasing = T)]
pve_NA18855_ord <- pve_NA18855[order(pve_NA18855, decreasing = T)]
pve_NA18870_ord <- pve_NA18870[order(pve_NA18870, decreasing = T)]
pve_NA19098_ord <- pve_NA19098[order(pve_NA19098, decreasing = T)]
pve_NA19101_ord <- pve_NA19101[order(pve_NA19101, decreasing = T)]
pve_NA19160_ord <- pve_NA19160[order(pve_NA19160, decreasing = T)]

names(pve_NA18511_ord)[1:5]
names(pve_NA18855_ord)[1:5]
names(pve_NA18870_ord)[1:5]
names(pve_NA19098_ord)[1:5]
names(pve_NA19101_ord)[1:5]
names(pve_NA19160_ord)[1:5]

library(venn)
intersect(intersect(
  intersect(intersect(intersect(names(pve_NA18511_ord)[1:500],
          names(pve_NA18855_ord)[1:500]),
          names(pve_NA18870_ord)[1:500]),
          names(pve_NA19098_ord)[1:500]),
          names(pve_NA19101_ord)[1:500]),
          names(pve_NA19160_ord)[1:500])
```


checking that trendfilter was done correctly

```{r}
df_NA19101 <- readRDS("../output/trendfilter-individual.Rmd/df_NA19101.rds")

plot(df_NA19101[which(rownames(df_NA19101)==names(pve_NA19101_ord)[1]),])
points(fits_NA19101[which(names(fits_NA19101) ==names(pve_NA19101_ord)[1])][[1]]$trend.yy,
       col="red", pch=16)
```

getting PVE of CDK1 from all individuals

```{r}
pve_NA18511_ord[names(pve_NA18511_ord)=="ENSG00000170312"]
pve_NA18855_ord[names(pve_NA18855_ord)=="ENSG00000170312"]
pve_NA18870_ord[names(pve_NA18870_ord)=="ENSG00000170312"]
pve_NA19098_ord[names(pve_NA19098_ord)=="ENSG00000170312"]
pve_NA19101_ord[names(pve_NA19101_ord)=="ENSG00000170312"]
pve_NA19160_ord[names(pve_NA19160_ord)=="ENSG00000170312"]
```
distribution of cell times in different individuals

```{r}
eset <- readRDS("../data/eset-final.rds")
library(Biobase)
hist(pData(eset)$theta)

ids <- unique(pData(eset)$chip_id)
par(mfrow=c(2,3))
for (i in 1:6) {
  hist(pData(eset)$theta[pData(eset)$chip_id==ids[i]],
       main=ids[i],
       xlab="Fucci phase")
}

plot(pData(eset)$theta[pData(eset)$chip_id==ids[1]])
points(pData(eset)$theta[pData(eset)$chip_id==ids[2]], col="red")
```



---


```{r}
sessionInfo()
```

